### 2013-10-31 Transient_growth_prl.tex
Create repository

\documentclass[showpacs,preprintnumbers,amsmath,amssymb,superscriptaddress,aip]{revtex4-1}
\usepackage{graphicx}
%\setlength{\parindent}{0in}

% General Latex  --------------------------------------------------
\def\beq{\begin{equation}}
\def\eeq{\end{equation}}
\def\beqar{\begin{eqnarray}}
\def\eeqar{\end{eqnarray}}
\def\nn{\nonumber}
\def\ol{\overline}
\def\para{\parallel}

% Operators  ------------------------------------------------------
\newcommand{\diff}[2]{\frac{d#1}{d#2}}
\newcommand{\diffs}[2]{\frac{d^2#1}{d#2^2}}
\newcommand{\pdiff}[2]{\frac{\partial#1}{\partial#2}}
\newcommand{\pdiffs}[2]{\frac{\partial^2#1}{\partial#2^2}}
\newcommand{\pdiffxy}[3]{\frac{\partial^2#1}{\partial#2 \partial#3}}
\newcommand{\pdt}{\partial_t}
\newcommand{\pdr}{\partial_r}
\newcommand{\pdth}{\partial_\theta}
\newcommand{\pdrr}{\partial^2_r}

\newcommand{\enum}[2]{{#1}\times10^{#2}} % 4.2x10^{3} = \enum{4.2}{3}

\newcommand{\vect}[1]{{\bf #1}}
%\newcommand{\vect}{\overrightarrow}
%\newcommand{\vect}{\vec}
\def\div{\nabla\cdot}
\def\grad{\nabla}
\def\curl{\nabla\times}
\newcommand{\gradpar}{\grad_\parallel}
\newcommand{\gradperp}{\grad_\perp}
\newcommand{\gradr}{\grad_r}
\newcommand{\defeq}{\ensuremath{\stackrel{\text{\tiny def}}{=}}}

\newcommand{\savg}[1]{\left<{#1}\right>}
\newcommand{\vavg}[1]{\left<{#1}\right>_V}
\newcommand{\thavg}[1]{\left<{#1}\right>_\theta}

% Variable names  -------------------------------------------------
\newcommand{\vpar} {v_\parallel}
\newcommand{\Apar} {A_\parallel}
\newcommand{\jpar} {j_\parallel}
\newcommand{\kpar} {k_\parallel}
\newcommand{\kperp} {k_\perp }
\newcommand{\vperp} {v_\perp }
\newcommand{\kthe}{k_\theta}

\newcommand{\Evec}{\ensuremath{\boldsymbol{{\rm E}}}}
\newcommand{\Bvec}{\ensuremath{\boldsymbol{{\rm B}}}}
\newcommand{\Jvec}{\ensuremath{\boldsymbol{{\rm J}}}}
\newcommand{\Fvec}{\ensuremath{\boldsymbol{{\rm F}}}}
\newcommand{\fvec}{\ensuremath{\boldsymbol{{\rm f}}}}
\newcommand{\vE}{\ensuremath{\boldsymbol{{\rm v}_{E}}}}
\newcommand{\bo}{\ensuremath{\boldsymbol{{\rm b}_0}}}
\newcommand{\bvec}{\ensuremath{\boldsymbol{{\rm b}}}}
\newcommand{\xvec}{\ensuremath{\boldsymbol{{\rm x}}}}
\newcommand{\yvec}{\ensuremath{\boldsymbol{{\rm y}}}}
\newcommand{\zvec}{\ensuremath{\boldsymbol{{\rm z}}}}
\newcommand{\vvec}{\ensuremath{\boldsymbol{{\rm v}}}}
\newcommand{\jvec}{\ensuremath{\boldsymbol{{\rm j}}}}

\newcommand{\bxgp}{\bvec\times\gradperp}

\newcommand{\vve}{\ensuremath{\boldsymbol{{\rm v}}_{e}}}
\newcommand{\vvi}{\ensuremath{\boldsymbol{{\rm v}}_{i}}}
\newcommand{\vpe}{v_{\parallel e}}
\newcommand{\vpi}{v_{\parallel i}}
\newcommand{\vvE}{\ensuremath{\boldsymbol{{\rm v}}_{E}}}
\newcommand{\vvD}{\ensuremath{\boldsymbol{{\rm v}}_{D}}}

\newcommand{\nuei}{\nu_{ei}}
\newcommand{\nuii}{\nu_{ii}}
\newcommand{\nue}{\nu_{e}}
\newcommand{\nuen}{\nu_{en}}
\newcommand{\nuin}{\nu_{in}}
\newcommand{\kpe}{\kappa_{\parallel e}}

\newcommand{\rs}{\rho_{s}}
\newcommand{\ri}{\rho_{i}}
\newcommand{\wci}{\Omega_{i}}
\newcommand{\wcix}{\Omega_{ix}}
\newcommand{\wce}{\Omega_{e}}
\newcommand{\tomega}{\tilde\omega}
\newcommand{\Isat}{I_{\rm sat}}
\newcommand{\fmie}{\frac{m_i}{m_e}}
\newcommand{\fmei}{\frac{m_e}{m_i}}

% Often used dimensions
\newcommand{\cm}{\rm cm}
\newcommand{\mm}{\rm mm}
\newcommand{\cmn}{{\rm cm}^{-3}}
\newcommand{\mn}{{\rm m}^{-3}}
\newcommand{\eV}{\rm eV}
\newcommand{\G}{\rm G}
\newcommand{\T}{\rm T}

\begin{document}

\title{Non-modal Growth in LAPD Turbulence}

\author{B. Friedman}
\email{friedman@physics.ucla.edu}

\author{T.A. Carter}

\affiliation{Department of Physics and Astronomy, University of California, Los Angeles, California 90095-1547, USA}

\begin{abstract}
Large Plasma Device (LAPD) [W. Gekelman \emph{et al.}, Rev. Sci. Inst. {\bf 62}, 2875 (1991)] 
\end{abstract}

\maketitle

\section{Introduction}

\section{The Simulation Model}
\label{dw_model}

A Braginskii-based fluid model~\cite{Braginskii1965} is used to simulate global drift wave turbulence in LAPD using the BOUT++ code~\cite{dudson2009}. 
The evolved variables in the model are the plasma density, $N$, the electron fluid parallel velocity $\vpe$, the potential vorticity $\varpi \equiv \gradperp \cdot (N_0 \gradperp \phi)$,
and the electron temperature $T_e$. The ions are assumed cold in the
model ($T_i = 0$), and sound wave effects are neglected ($v_i = 0$). Details of the simulation code, derivations of the model, grid convergence studies, and analyses of simplified models
may be found in previously published LAPD simulation studies~\cite{Popovich2010a,Popovich2010b,Umansky2011,friedman2012,friedman2012b}.

The equations are developed with Bohm normalizations: lengths are
normalized to the ion sound gyroradius, times to the ion
cyclotron time, velocities to the sound speed, densities to the equilibrium peak density, and electron
temperatures and potentials to the equilibrium peak electron temperature. These normalizations are constants (not functions of radius) and are calculated from these reference values:
the magnetic field is $1$ kG, the ion unit mass is $4$, the peak density is $2.86 \times 10^{12}$ cm$^{-3}$, and the peak electron temperature
is $6$ eV. The equations are:

\beqar
\label{ni_eq}
\pdt N = - {\mathbf v_E} \cdot \grad N_0 - N_0 \gradpar \vpe + \mu_N \gradperp^2 N + S_N + \{\phi,N\}, \\
\label{ve_eq}
\pdt \vpe = - \fmie \frac{T_{e0}}{N_0} \gradpar N - 1.71 \fmie \gradpar T_e  + \fmie \gradpar \phi - \nue \vpe + \{\phi,\vpe \}, \\
\label{rho_eq}
\pdt \varpi = - N_0 \gradpar \vpe  - \nuin \varpi + \mu_\phi \gradperp^2 \varpi + \{\phi,\varpi \}, \\
\label{te_eq}
\pdt T_e = - {\mathbf v_E} \cdot \grad T_{e0} - 1.71 \frac{2}{3} T_{e0} \gradpar \vpe + \frac{2}{3 N_0} \kpe \gradpar^2 T_e  \nonumber \\
- \frac{2 m_e}{m_i} \nue T_e  + \mu_T \gradperp^2 T_e +  S_T + \{\phi,T_e\}.
\eeqar

In these equations, $\mu_N$, $\mu_T$, and $\mu_\phi$ are artificial diffusion and viscosity coefficients used for subgrid dissipation. They are large enough to allow saturation
and grid convergence~\cite{friedman2012}, but small enough to allow for turbulence to develop. In the simulations, they are all given the same value of $1.25 \times 10^{-3}$ in Bohm-normalized units. 
This is the only free parameter in the simulations. All other parameters such as the electron collisionality $\nue$, ion-neutral
collisionality $\nuin$, parallel electron thermal conductivity $\kpe$, and mass ratio $\fmie$ are calculated from experimental quantities.
There are two sources of free energy: the density gradient due to the equilibrium density profile $N_0$, and the equilibrium electron temperature gradient in $T_{e0}$, both of which are
taken from experimental fits. $N_0$ and $T_{e0}$ are functions of only the radial cylindrical coordinate $r$, and they are shown in Fig.~\ref{eq_profiles}. 
The mean potential profile $\phi_0$ is set to zero in the model, and terms involving $\phi_0$ are not included in Eqs.~\ref{ni_eq}-\ref{te_eq}. 
The justification for this is that biasable azimuthal limiters in LAPD allow for the mean $E \times B$ flow and flow shear to be varied with high precision, even allowing the flow to be
nulled out~\cite{schaffner2012}. 
The simulations in this paper use the $N_0$, $T_{e0}$, and $\phi_0$ profiles from the nulled out flow experiment, justifying setting the mean potential profile to zero in the simulations.

Simulations also use density and temperature sources ($S_n$ and $S_T$) in order to keep the equilibrium profiles from relaxing away from their experimental shapes. 
These sources suppress the azimuthal averages ($m=0$ component of the density and temperature fluctuations) at each time step. 
The azimuthal average of the potential $\phi$ is allowed to evolve in
the simulation, allowing zonal flows to form, although they are relatively unimportant to the turbulent dynamics~\cite{friedman2012b}.

The terms in Poisson brackets are the $E \times B$ advective nonlinearities, which are the only nonlinearities used in the simulations.
The numerical simulations use finite differences in all three dimensions and use cylindrical annular geometry ($12<r<40$ cm). The radial extent used in the simulation
encompasses the region where fluctuations are above a few percent in the experiment. Therefore, the radial boundaries are fixed to zero value.

\section{Conclusion}
\label{conclusion}

\begin{acknowledgments}

\end{acknowledgments}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{refs}
%\bibliographystyle{unsrt}

\end{document}

### 2013-11-08 Transient_growth_prl.tex
Nov 8

@@
 
+Since the early 1990's, it has become clear that in dynamical systems with energetically conservative nonlinearities, 
+the linear operator must be non-normal for subcritical turbulence to occur~\cite{trefethen1993}.
+Non-normality is the condition that an operator or matrix does not commute with its adjoint. One consequence of non-normality is that the linear eigenvectors of the operator are nonorthogonal.
+In non-normal systems, even when all of the eigenvalues lie in the stable domain (where all linear eigenvectors decay exponential), 
+certain types of fluctuations in a system can access the free energy in the equilibrium gradients and transiently amplify the fluctuation energy in the system. 
+A paradigmatic illustration of this transient growth process may be seen if Figure 2 of a review by Schmid~\cite{schmid2007}.
+Such behavior is obscured by traditional normal mode analysis, which in non-normal systems, only effectively describes the long time asymptotic behavior of fluctuations under the 
+action of the linear operator. Transient events, such as those that can dominate nonlinear turbulent evolution, require initial-value (non-modal) calculations rather than normal mode calculations.
+Non-modal analysis, in fact, has provided the tool that can explain and predict subcritical turbulence.
+
+While the hydrodynamics community has embraced non-modal analysis and transient growth over the past two decades, the plasma community
+still heavily relies on normal mode analysis to inform turbulent predictions and observations, often without checking for non-modal effects. 
+While this has been successful in certain facets of plasma physics, it has failed in others.
+
+Direct numerical simulations of plasma turbulence have become increasingly common with the increase in computing power. These simulations produce so much data that the results can be difficult to
+comprehend. It is often easiest to compare simulation results with simple models to achieve some understanding.
+
 \section{The Simulation Model}
@@
 collisionality $\nuin$, parallel electron thermal conductivity $\kpe$, and mass ratio $\fmie$ are calculated from experimental quantities.
-There are two sources of free energy: the density gradient due to the equilibrium density profile $N_0$, and the equilibrium electron temperature gradient in $T_{e0}$, both of which are
-taken from experimental fits. $N_0$ and $T_{e0}$ are functions of only the radial cylindrical coordinate $r$, and they are shown in Fig.~\ref{eq_profiles}. 
+There are two sources of free energy: the radial density gradient due to the equilibrium density profile $N_0$, and the equilibrium radial electron temperature gradient in $T_{e0}$, both of which are
+taken from experimental fits. 
 The mean potential profile $\phi_0$ is set to zero in the model, and terms involving $\phi_0$ are not included in Eqs.~\ref{ni_eq}-\ref{te_eq}. 

### 2013-11-26 Transient_growth_prl.tex
Nov 25

--- 2013-10-31 Transient_growth_prl.tex
@@
 
-\section{The Simulation Model}
-\label{dw_model}
+Normal mode analysis -- dealing with the eigenvalues and eigenvectors of a dynamical system -- has been used for the last two centuries to solve a number of problems ranging from
+heat conduction along a solid bar to quantum mechanical energy states of atoms. Despite its wide-ranging success, normal mode analysis has seemingly failed in particular instances,
+particularly in predicting the onset of turbulence in many hydrodynamic flows, where the turbulence is called subcritical. 
+The reason for this failure was not clearly explained until the early 1990's when Trefethen and others attributed the pitfalls of normal mode analysis to the non-normality of
+systems~\cite{trefethen1993,trefethen2005} -- a non-normal system is one whose linear matrix or operator does not commute with its adjoint. 
+Non-normal systems have eigenvectors that are nonorthogonal to one another. One consequence of non-normality is that even when all of the eigenvalues lie in the stable domain 
+(where all linear eigenvectors decay exponential), some fluctuations can transiently grow by accessing the free energy in the equilibrium gradients, allowing for sustained turbulence. 
+A paradigmatic illustration of this transient growth process for two nonorthogonal eigenvectors may be seen if Figure 2 of a review paper by Schmid~\cite{schmid2007}.
+Such behavior is obscured by traditional normal mode analysis, which only effectively describes the long time asymptotic behavior of fluctuations under the 
+action of the linear operator. Transient events, such as those that can dominate nonlinear turbulent evolution, require initial-value (non-modal) calculations rather than normal mode calculations.
 
-A Braginskii-based fluid model~\cite{Braginskii1965} is used to simulate global drift wave turbulence in LAPD using the BOUT++ code~\cite{dudson2009}. 
-The evolved variables in the model are the plasma density, $N$, the electron fluid parallel velocity $\vpe$, the potential vorticity $\varpi \equiv \gradperp \cdot (N_0 \gradperp \phi)$,
-and the electron temperature $T_e$. The ions are assumed cold in the
-model ($T_i = 0$), and sound wave effects are neglected ($v_i = 0$). Details of the simulation code, derivations of the model, grid convergence studies, and analyses of simplified models
-may be found in previously published LAPD simulation studies~\cite{Popovich2010a,Popovich2010b,Umansky2011,friedman2012,friedman2012b}.
+Non-modal analysis has been embraced by the hydrodynamics community over the past two decades in the attempt to explain and predict subcritical turbulence, but the plasma community
+still heavily relies on normal mode analysis to inform turbulent predictions and observations, with a few notable exceptions~\cite{camargo1998,camporeale2010,schekochihin2012}. 
+Additionally, non-modal treatments thus far have generally been used to explain rather than predict turbulent characteristics, 
+while the few predictions that have been made are often qualitative rather than quantitative. Perhaps this is why non-modal analysis has not received greater attention.
+Therefore, this paper takes up the task of developing a quantitative approach to predicting turbulent properties using only non-modal linear calculations. 
+The basic approach we take is to solve the linear initial value problem using an ensemble of random initial conditions, 
+and then calculate the average growth rate of the solutions up to one nonlinear decorrelation time, which we take to be the characteristic time of the linear process. 
+The effective growth rate spectrum produced by this technique can then be used in a quasilinear fashion to gain more information about the turbulence.
+While this technique is general and can be applied to many nonlinear dynamical systems, we restrict our treatment to one particular model that relates to one particular experiment. Not only does
+this allow us to thoroughly explain and validate this technique, but it also allows us to attribute transient growth effects to physically realizable turbulence.
 
-The equations are developed with Bohm normalizations: lengths are
-normalized to the ion sound gyroradius, times to the ion
-cyclotron time, velocities to the sound speed, densities to the equilibrium peak density, and electron
-temperatures and potentials to the equilibrium peak electron temperature. These normalizations are constants (not functions of radius) and are calculated from these reference values:
-the magnetic field is $1$ kG, the ion unit mass is $4$, the peak density is $2.86 \times 10^{12}$ cm$^{-3}$, and the peak electron temperature
-is $6$ eV. The equations are:
+The experiment considered is a drift wave turbulence experiment conducted in the Large Plasma Device (LAPD)~\cite{gekelman1991}. LAPD is a linear machine with a straight magnetic field $\mathbf{B}$.
+Due to its large dimensions and high collisionality, LAPD is suitable for fluid modelling. We use a reduced Braginskii~\cite{Braginskii1965} 2-fluid model:
 
@@
 \pdt T_e = - {\mathbf v_E} \cdot \grad T_{e0} - 1.71 \frac{2}{3} T_{e0} \gradpar \vpe + \frac{2}{3 N_0} \kpe \gradpar^2 T_e  \nonumber \\
-- \frac{2 m_e}{m_i} \nue T_e  + \mu_T \gradperp^2 T_e +  S_T + \{\phi,T_e\}.
+- \frac{2 m_e}{m_i} \nue T_e  + \mu_T \gradperp^2 T_e +  S_T + \{\phi,T_e\},
 \eeqar
+where $N_0$ and $N$ are the equilibrium and fluctuating density, $\vpe$ is the fluctuating parallel electron velocity, $\varpi \equiv \gradperp \cdot (N_0 \gradperp \phi)$ is the potential vorticity
+of the fluctuating potential $\phi$, and $T_{e0}$ and $T_e$ are the equilibrium and fluctuating electron temperature. The equations are developed with Bohm normalizations: lengths are
+normalized to the ion sound gyroradius $\rho_s$, times to the ion cyclotron time $\omega_{ci}^{-1}$, velocities to the sound speed $c_s$, densities to the equilibrium peak density, and electron
+temperatures and potentials to the equilibrium peak electron temperature. The profiles $N_0$ and $T_{e0}$ and other parameters are all taken from experimental measurements. Note that $\phi_0$ is
+neglected because we model only one experiment in which the equilibrium radial electric field was nulled out by boundary biasing~\cite{schaffner2012}.
 
-In these equations, $\mu_N$, $\mu_T$, and $\mu_\phi$ are artificial diffusion and viscosity coefficients used for subgrid dissipation. They are large enough to allow saturation
-and grid convergence~\cite{friedman2012}, but small enough to allow for turbulence to develop. In the simulations, they are all given the same value of $1.25 \times 10^{-3}$ in Bohm-normalized units. 
-This is the only free parameter in the simulations. All other parameters such as the electron collisionality $\nue$, ion-neutral
-collisionality $\nuin$, parallel electron thermal conductivity $\kpe$, and mass ratio $\fmie$ are calculated from experimental quantities.
-There are two sources of free energy: the density gradient due to the equilibrium density profile $N_0$, and the equilibrium electron temperature gradient in $T_{e0}$, both of which are
-taken from experimental fits. $N_0$ and $T_{e0}$ are functions of only the radial cylindrical coordinate $r$, and they are shown in Fig.~\ref{eq_profiles}. 
-The mean potential profile $\phi_0$ is set to zero in the model, and terms involving $\phi_0$ are not included in Eqs.~\ref{ni_eq}-\ref{te_eq}. 
-The justification for this is that biasable azimuthal limiters in LAPD allow for the mean $E \times B$ flow and flow shear to be varied with high precision, even allowing the flow to be
-nulled out~\cite{schaffner2012}. 
-The simulations in this paper use the $N_0$, $T_{e0}$, and $\phi_0$ profiles from the nulled out flow experiment, justifying setting the mean potential profile to zero in the simulations.
+The equations are global and partially linearized -- the only nonlinearities we retain are the advective nonlinearities in the Poisson brackets. 
+We also add artificial diffusion and viscosity terms with small numerical coefficients ($\mu_N, \mu_\phi, \mu_T = 10^{-3}$), 
+used for numerical stability in nonlinear simulations, which are performed with the BOUT++ code~\cite{dudson2009}. We use periodic axial and zero value radial 
+boundary conditions. Further details of the model, including validation studies, may be found in the references~\cite{Popovich2010a,Popovich2010b,Umansky2011,friedman2012,friedman2012b},
+though we mention here that the model reproduces all statistical properties of the experimental turbulence within a factor of two, and in most cases, much better than that.
+Of particular interest, was the finding that the sustainment of the turbulence in the nonlinear simulation is dominated by a nonlinear instability process~\cite{friedman2012,friedman2012b}.
+This nonlinear instability process -- discovered by Drake et al.~\cite{drake1995} in simulations using a similar plasma model to ours -- works as follows: long $k_\para=0$ convective filaments transport
+density (and temperature to a lesser extent) across the equilibrium density profile, setting up $k_\para=0$ density filaments. These filaments are unstable to drift waves causing a secondary drift
+wave instability to grow on the filaments. These drift waves, which have finite $k_\para$ and propagate primarily radially, nonlinearly couple to one another and reinforce the convective filaments.
 
-Simulations also use density and temperature sources ($S_n$ and $S_T$) in order to keep the equilibrium profiles from relaxing away from their experimental shapes. 
-These sources suppress the azimuthal averages ($m=0$ component of the density and temperature fluctuations) at each time step. 
-The azimuthal average of the potential $\phi$ is allowed to evolve in
-the simulation, allowing zonal flows to form, although they are relatively unimportant to the turbulent dynamics~\cite{friedman2012b}.
+\begin{figure}
+\centerline{\includegraphics[]{spec_nl_gamma}}
+\caption{The linear and turbulent growth rate $m$ spectra for $n=0$ (solid lines) and $n=1$ (dashed lines) Fourier components. The linear growth rates are those for the least stable eigenmodes at each
+$n$ and $m$, while the turbulent growth rates represent $\pdiff{E}{t}|_{\rm{lin}}/2 E$ from the nonlinear simulation.
+The shaded region marks the $1 \sigma$ spread in the turbulent growth rate spectrum,
+obtained from the distribution of growth rates calculated from the nonlinear simulation over a long time range.}
+\label{spec_nl_gamma}
+\end{figure}
 
-The terms in Poisson brackets are the $E \times B$ advective nonlinearities, which are the only nonlinearities used in the simulations.
-The numerical simulations use finite differences in all three dimensions and use cylindrical annular geometry ($12<r<40$ cm). The radial extent used in the simulation
-encompasses the region where fluctuations are above a few percent in the experiment. Therefore, the radial boundaries are fixed to zero value.
+Although the instability mechanism is nonlinear, the first part of the mechanism -- the transport of density by the convective filaments -- is a linear one. 
+In fact, it is surprisingly similar to the so-called "lift-up'' mechanism often seen in subcritical hydrodynamic shear flows, for which transient growth arguments were first applied~\cite{trefethen1993},
+making it particularly apt to explore in a non-modal framework. Furthermore, since the other steps of the nonlinear instability
+are controlled by nonlinear interactions, the convective transport is the only step responsible for energy injection and dissipation from the laminar state. This fact may be seen from simple
+energetic principles. Starting from Eqs.~\ref{ni_eq}-\ref{te_eq},
+an equation for the evolution of the energy can be obtained~\cite{friedman2012,friedman2012b}. The equation takes the form:
 
-\section{Conclusion}
-\label{conclusion}
+\beq
+\label{dEdt_def}
+\diff{E(m,n,t)}{t} = Q(m,n,t) + D(m,n,t) + \sum_{m',n'} T(m,m',n,n',t).
+\eeq
+Here, $m$ and $n$ are the azimuthal and axial Fourier mode numbers, respectively. Note that since Eqs.~\ref{ni_eq}-\ref{te_eq} are not explicitly dependent on $\theta$ and $z$, the energy equation
+is separable for every $m$ and $n$. Furthermore, $Q$ represents the linear energy exchange with the equilibrium, $D$ is the dissipation from irreversible processes
+like collisions, and $T$ accounts for the nonlinear triad wave transfer. Because $T$ results from the Poisson bracketed advective nonlinearities in the model and $\int f \{g,f\} dV = 0$, 
+$\sum_{m,m',n,n'} T(m,m',n,n',t)=0$, meaning the nonlinearities cause no net energy injection or dissipation. In other words, the fluctuation energy originates completely from the linear terms.
+
+The expression $\diff{E(m,n,t)}{t}\big|_{\rm{lin}} = Q(m,n,t) + D(m,n,t)$ accounts for the total net energy injection at each $m,n$ into the fluctuations. 
+An effective instantaneous growth rate $\gamma_{\rm{eff}}(m,n,t) = \diff{E(m,n,t)}{t}\bigg|_{\rm{lin}}/ 2 E(m,n,t)$ can then be defined. 
+This effective growth rate is calculated from the spatial structure of the fluctuations, so it has meaning for turbulent fluctuations as well as those obtained from a linear simulation. If calculated
+from a linear simulation run for a sufficiently long time, $\gamma_{\rm{eff}}(m,n)$ is the same as the eigenmode growth rate of the least stable linear eigenmode for each $m,n$.
+We plot this effective growth rate in Fig.~\ref{spec_nl_gamma} for $n=0$ and $n=1$ components. The turbulent growth rate curves are calculated from this formula, 
+averaged over a time of about $3 \times 10^{4} \omega_{ci}^{-1}$ during the saturated turbulent phase of the nonlinear simulation --$\gamma_{\rm{Turb}}(m,n) = \int \gamma_{\rm{eff}}(m,n,t) dt$. 
+We show this turbulent growth rate along with the $1 \sigma$ contour, which indicates how greatly $\gamma_{\rm{eff}}(m,n,t)$ varies over time.
+Along with these, the growth rates of the least stable (or most unstable) linear eigenmodes are shown.
+
+From Fig.~\ref{spec_nl_gamma}, one sees that the linear eigenmodes are unstable for $n=1$ and stable for $n=0$ as expected for drift waves. 
+However, the turbulent growth rates are positive for $n=0$ and low $m$, but negative for $n=1$ at all $m$. 
+The net injection of the $n=0$ components is reminiscent of subcritical turbulence, where there is turbulent sustainment in the absense of linear instability.
+Moreover, the $n=0$ energy injection is a manifestation of the convective filaments drawing energy from the equilibrium density gradient and depositing this energy into the density fluctuations,
+which again, is similar to the hydrodynamic ``lift-up'' mechanism associated with subcritical hydrodynamic shear-flow turbulence.
+It it prudent then to call the turbulence in the LAPD simulation subcritical or at least subcritical-like.
+
+Like subcritical hydrodynamic turbulence, the cause of subcritical LAPD turbulence (the positive growth rate of the $n=0$ structures) 
+is a transient growth mechanism due to nonorthogonal linear eigenmodes. The transient growth can be seen clearly in Fig.~\ref{transient_n0_n1_growth},
+where we follow the evolution of the energy of several Fourier $m,n$ modes after turning off the nonlinearities in the turbulent simulation. The $n=0$ modes grow for some time before decaying
+exponentially at the rate of their least stable eigenmode. Conversely, the $n=1,m=20$ mode decays transiently before growing exponentially. This is not an attribute of subcritical turbulence
+because there are no unstable eigenmodes in truly subcritical turbulence, but it is nevertheless interesting and is associated with a transient event.
+The growth (or decay) rate of these curves at $t=0$ is necessarily equal to the growth rate that is used in the calculation of Fig.~\ref{spec_nl_gamma}. 
+This $t=0$ growth rate, however, is just one number that goes into that calculation. The calculation is based on the turbulence over a long time range; this is just one instant of time.
+In any case, the $t=0$ behavior of the curves in Fig.~\ref{transient_n0_n1_growth} generally corroborates with the effective growth rates in Fig.~\ref{spec_nl_gamma}.
+
+\begin{figure}
+\centerline{\includegraphics[]{transient_n0_n1_growth}}
+\caption{The linear evolution of the energy of several Fourier modes starting from a turbulent initial state. The $n=0$ curves have an initial period of transient growth before exponentially decaying,
+while the $n=1,m=20$ curve transiently decays before growing exponentially.}
+\label{transient_n0_n1_growth}
+\end{figure}
+
+We stress again that Fig.~\ref{transient_n0_n1_growth} is obtained by \emph{linear} evolution from an initial turbulent state.
+Since the transient growth (and decay) is a purely linear phenomenon, it appears possible to predict turbulent growth rates like those of Fig.~\ref{spec_nl_gamma} using only linear calculations.
+Such calculations would be much faster and easier to interpret than direct simulations of nonlinear equations. However, the difficulty in such calculations lies in the dependence of the transient
+evolution on the initial conditions and the change in the growth rate as a function of time as the linear evolution progresses. To overcome this difficulty, we propose a statistical procedure
+that linearly evolves an ensemble of random initial conditions and calculates the average growth rate of the evolution over a characteristic linear time.
+
+To elucidate this procedure, we take Eqs.~\ref{ni_eq}-\ref{te_eq}, remove the nonlinearities and sources and Fourier decompose in the azimuthal and axial directions.
+Then, we discritize in the radial direction ($r \rightarrow r_0, r_1, \ldots, r_N $), and approximate radial derivatives with finite differences. 
+The resulting system of equations may be written in matrix form:
+
+\beq
+\label{lin_eq}
+\mathbf{B}_{m,n} \diff{\mathbf{v}_{m,n}(t)}{t} = \mathbf{C}_{m,n} \mathbf{v}_{m,n}(t),
+\eeq
+where $\mathbf{v}_{m,n} = \left( N(r_0), N(r_1), \ldots, \vpe(r_0), \vpe(r_1), \ldots, \phi(r_0), \phi(r_1), \ldots, T_e(r_0), T_e(r_1), \ldots \right)_{m,n}^{T}$,
+and $\mathbf{B}_{m,n}$ and $\mathbf{C}_{m,n}$ are coefficient matrices that include the equilibrium information and finite difference coefficients. 
+The equations for different $m$ and $n$ are separable because the equations are linear and the eigenfunctions have azimuthal and axial Fourier dependence resulting from the lack of explicit
+dependence on $\theta$ and $z$ in the equations. Thus, eigenmodes with different $m$ and $n$ are orthogonal and linearly non-interacting. Note that for each $m,n$, there exist $4 \times N_r$ linearly
+independent, but nonorthogonal eigenvectors. Hence forth, we drop the $m,n$ subscripts.
+
+In order to use non-modal analysis to calculate growth rates and other measures, one must choose a norm and inner product with which to work. While any choice of
+inner product is possible, including that which orthogonalizes all eigenvectors, a physically relevant one is generally preferred~\cite{camargo1998,schmid2007,camporeale2010}. The most commonly used
+is an energy inner product. Recall that the inner product of two vectors may be written $\left< \mathbf{x},\mathbf{y} \right> = \mathbf{y}^{*} \mathbf{M} \mathbf{x}$,
+where $^*$ stands for the conjugate transpose, and $\mathbf{M}$ is a positive-definite matrix. $\mathbf{M}$ should be chosen such that 
+$||\mathbf{u}|| = \sqrt{\left< \mathbf{u},\mathbf{u} \right>} = \sqrt{E}$ -- that is, the norm of the state vector is the square root of the energy. Furthermore, it is often convenient in
+computations to use the $L_2$-norm, $||\mathbf{u}||_2 = \sqrt{\sum_i |u_i|^2}$.
+These requirements can be accomplished through the change of variables $\mathbf{u} = \mathbf{M}^{\frac{1}{2}} \mathbf{v}$.
+Then Eq.~\ref{lin_eq} becomes
+
+\beq
+\label{lin_eq_A}
+\diff{\mathbf{u}}{t} = \mathbf{A} \mathbf{u},  \quad \rm{where} \ \mathbf{A} = \mathbf{M}^{-\frac{1}{2}} \mathbf{B}^{-1} \mathbf{C} \mathbf{M}^{\frac{1}{2}}.
+\eeq
+The eigenvalues of $\mathbf{A}$ are the same as those of $\mathbf{B}^{-1} \mathbf{C}$ because $\mathbf{M}^{-\frac{1}{2}} \mathbf{B}^{-1} \mathbf{C} \mathbf{M}^{\frac{1}{2}}$ is a similarity transformation.
+
+The solution of Eq.~\ref{lin_eq_A} is
+
+\beq
+\label{lin_soln}
+\mathbf{u}(t) = e^{\mathbf{A} t} \mathbf{u}(0).
+\eeq
+This solution depends on the initial condition $\mathbf{u}(0)$ in addition to the spectral properties of $\mathbf{A}$. For purposes of turbulent growth rate prediction, we are interested in
+the behavior of $||\mathbf{u}(t)||$. Therefore, we introduce the growth ratio $G(t)$, which measures the amplification or reduction in the square root of the energy from an initial state:
+
+\beq
+\label{g_def}
+G(t) = \frac{||\mathbf{u}(t)||}{||\mathbf{u}(0)||} = \frac{||e^{\mathbf{A} t} \mathbf{u}(0)||}{||\mathbf{u}(0)||}.
+\eeq
+It can be shown that this growth ratio is bounded from above by $G_{\rm{max}}(t) = ||e^{\mathbf{A} t}||$. Furthermore, $G_{\rm{max}}(t) = e^{\gamma_s t}$, where $\gamma_s$ is the spectral growth rate
+(the real part of the eigenvalue with largest real part), if and only if $\mathbf{A}$ is normal~\cite{schmid2007}. Otherwise, $G_{\rm{max}}(t) > e^{\gamma_s t}$.
+
+It is common practice in normal mode analysis to look for the fastest growing eigenmode, the idea being that eventually the fastest growing structure will overcome the others.
+For the non-modal case, it is common to study the properties of $G_{\rm{max}}(t)$, because if $G_{\rm{max}}(t) > 1$ at any time, there is a possibility of amplification, even if it is only transient.
+This allows for the possibility of exciting subcritical turbulence, which is where non-modal analysis is most often applied. 
+However, it can be misleading to study only $G_{\rm{max}}(t)$ in instances where turbulence is already known to exist and one is interested in predicting specific properties of turbulence. 
+The main reason is that $G_{\rm{max}}(t)$ is actually the upper envelope of all possible $G(t)$ curves.
+In other words, no one particular initial condition $\mathbf{u}(0)$ evolves along $G_{\rm{max}}(t)$.
+One particular $\mathbf{u}_\tau(0)$ will evolve such that  $G(\tau)= G_{\rm{max}}(\tau)$. Another $\mathbf{u}_\rho(0)$ will evolve such that $G(\rho)= G_{\rm{max}}(\rho)$. 
+Furthermore, it isn't obvious what kind of spatial structure or structures will come to dominate a turbulent system. 
+Unlike in the normal case, in the non-normal case, optimal structures don't amplify themselves. They amplify the energy, but evolve into different structures, at least until late times.
+
+We illustrate $G_{\rm{max}}(t)$ along with $G(t)$ for several random initial conditions in Fig.~\ref{m20n0_time_evs} a), all at $n=0$ and $m=20$. First, despite the fact that all eigenmodes at this
+wavenumber are stable, there is the possibility for transient amplification by a factor of $20$. Second, a sample of randomly initialized vectors are all transiently amplified, but none display
+optimal growth as compared to $G_{\rm{max}}(t)$. We show a zoomed-in view of these randomly initialized curves in Fig.~\ref{m20n0_time_evs} b).
+From the figure, it appears extremely unlikely to choose a random initial vector that either optimally amplifies the energy at any time or one that monotonically decays, which would be the case
+if the initial vector were one of the linear eigenvectors.
+
+\begin{figure}
+\centerline{\includegraphics[]{m20n0_time_evs}}
+\caption{a) The maximum growth ratio curve $G_{\rm{max}}(t)$ (dashed line) for $n=0,m=20$, and an ensemble of growth ratio curves (solid gray lines, with the solid black line the ensemble average)
+that start with a random initial condition $u(0)$ and evolve under the linear operator. b) A zoomed in version of the ensemble of growth ratio curves. c) A further zoomed in version of the ensemble of
+curves with a time range of $2 \pi / \omega_*$, which is the characteristic linear time scale for drift waves. d) The instantaneous growth rates as a function of time of the maximum and the ensemble 
+average growth ratios.}
+\label{m20n0_time_evs}
+\end{figure}
+
+It is not surprising that non-modal analysis has remained a qualitative endeavor due to its intrinsic dependence on initial conditions and the time-dependent growth properties of initial-value analysis.
+The key to quantifying non-modal analysis and making it predictive is to successfully model the effect that the nonlinearities have on the transient linear processes.
+In this regard, we model the nonlinearities as a periodic randomizing force. That is, the nonlinearities randomize the turbulent spatial structures on a characteristic nonlinear time scale. 
+In practice, we completely randomize the system (initial condition) and then let it evolve linearly for a time $t_{nl}$, at which point, the system is re-randomized.
+While the nonlinearities obviously do not affect the system in discrete time steps, this simplification provides a way to use linear non-modal analysis in a predictive manner. 
+Since we do not know, \emph{a priori} at what level the turbulence will saturate, 
+we cannot calculate the relative magnitude of the nonlinear terms in the equations to get the nonlinear time scale $t_{nl}$.
+Thus, we invoke the conjecture of \emph{Critical Balance}, which posits that the nonlinear time scale equals the linear time scale at all spatial scales~\cite{schekochihin2012}. 
+We can then estimate the nonlinear time scale as $t_{nl} = \omega_s^{-1}$, the inverse of the linear frequency of the fastest growing linear eigenmode (at each $m,n$).
+
+In Fig.~\ref{m20n0_time_evs} c), we show the randomly initialized linear evolution curves from start up until a time of $\omega_s^{-1}$. 
+Then, in Fig.~\ref{m20n0_time_evs} d), we calculate the instantaneous growth rate as a function of time of the ensemble averaged curve over that time period. 
+For comparison, the instantaneous growth rate of the maximum growth curve is also shown.
+In theory, this method should produce ensembled-averaged growth rate curves (like that in Fig.~\ref{m20n0_time_evs} d)) that have instantaneous growth rates consistent with the instantaneous
+turbulent growth rates. Both the instantaneous transient growth rates and the turbulent growth rates have an average and a range associated with them. We show a comparison of their averages
+in Fig.~\ref{spec_nl_transient_gamma}, which we call the Transient and Turbulent Growth Rate Spectra. They are fairly similar, especially when compared to the eigenmode spectra, also
+shown in Fig.~\ref{spec_nl_transient_gamma}. Thus, this randomizing transient method, which requires only linear calculations, somewhat reproduces the energetic properties of the turbulence.
+
+\begin{figure}
+\centerline{\includegraphics[]{spec_nl_transient_gamma}}
+\caption{The linear, turbulent, and transient growth rate spectra. The transient growth rate spectra is calculated by the average growth rate over a time $2 \pi / \omega_*$ for the ensemble average
+growth ratio curves.}
+\label{spec_nl_transient_gamma}
+\end{figure}
+
+These transient growth rates that comprise the curves in Fig.~\ref{spec_nl_transient_gamma} are more physically relevant to turbulence than the linear eigenmode growth rates.
+When the linear system under
+consideration is highly non-normal, transient growth rate analysis should be performed in place of (or in addition to) linear eigenmode analysis. Furthermore, the transient growth rates should
+be used in the kinds of predictions that are currently made with eigenmode growth rates. For instance, quasilinear theory uses the most unstable eigenmode growth rate $\gamma_{s,max}$ 
+to predict the turbulent saturation levels through the mixing length formula $\gamma_s/k_\perp^2$.
+Using this formula with $\gamma_{s,\rm{max}} = 0.002$ and $m=60$, one would predict a turbulent saturation
+level on the order of $0.1 \%$. On the other hand, using this formula with the transient values -- $\gamma_{t,\rm{max}} = 0.003, m=10$ -- we predict the saturation level on the order of $10 \%$,
+which matches the saturation level in the nonlinear simulation.
 

### 2013-12-03 Transient_growth_prl.tex
Dec 3

@@
 used for numerical stability in nonlinear simulations, which are performed with the BOUT++ code~\cite{dudson2009}. We use periodic axial and zero value radial 
-boundary conditions. Further details of the model, including validation studies, may be found in the references~\cite{Popovich2010a,Popovich2010b,Umansky2011,friedman2012,friedman2012b},
+boundary conditions. Further details of the model, including validation studies, may be found in the references~\cite{Popovich2010a,Popovich2010b,Umansky2011,friedman2012b,friedman2013},
 though we mention here that the model reproduces all statistical properties of the experimental turbulence within a factor of two, and in most cases, much better than that.
-Of particular interest, was the finding that the sustainment of the turbulence in the nonlinear simulation is dominated by a nonlinear instability process~\cite{friedman2012,friedman2012b}.
+Of particular interest, was the finding that the sustainment of the turbulence in the nonlinear simulation is dominated by a nonlinear instability process~\cite{friedman2012b,friedman2013}.
 This nonlinear instability process -- discovered by Drake et al.~\cite{drake1995} in simulations using a similar plasma model to ours -- works as follows: long $k_\para=0$ convective filaments transport
@@
 energetic principles. Starting from Eqs.~\ref{ni_eq}-\ref{te_eq},
-an equation for the evolution of the energy can be obtained~\cite{friedman2012,friedman2012b}. The equation takes the form:
+an equation for the evolution of the energy can be obtained~\cite{friedman2012b,friedman2013}. The equation takes the form:
 
@@
 that start with a random initial condition $u(0)$ and evolve under the linear operator. b) A zoomed in version of the ensemble of growth ratio curves. c) A further zoomed in version of the ensemble of
-curves with a time range of $2 \pi / \omega_*$, which is the characteristic linear time scale for drift waves. d) The instantaneous growth rates as a function of time of the maximum and the ensemble 
-average growth ratios.}
+curves with a time range of $\omega_s^{-1}$, which is the characteristic linear time scale for drift waves. 
+d) The instantaneous growth rate as a function of time of the the ensemble averaged growth ratio along with the average value over the time spanning $\omega_s^{-1}$.}
 \label{m20n0_time_evs}
@@
 we cannot calculate the relative magnitude of the nonlinear terms in the equations to get the nonlinear time scale $t_{nl}$.
-Thus, we invoke the conjecture of \emph{Critical Balance}, which posits that the nonlinear time scale equals the linear time scale at all spatial scales~\cite{schekochihin2012}. 
+Thus, we invoke the conjecture of \emph{critical balance}, which posits that the nonlinear time scale equals the linear time scale at all spatial scales~\cite{schekochihin2012}. 
 We can then estimate the nonlinear time scale as $t_{nl} = \omega_s^{-1}$, the inverse of the linear frequency of the fastest growing linear eigenmode (at each $m,n$).
@@
 turbulent growth rates. Both the instantaneous transient growth rates and the turbulent growth rates have an average and a range associated with them. We show a comparison of their averages
-in Fig.~\ref{spec_nl_transient_gamma}, which we call the Transient and Turbulent Growth Rate Spectra. They are fairly similar, especially when compared to the eigenmode spectra, also
-shown in Fig.~\ref{spec_nl_transient_gamma}. Thus, this randomizing transient method, which requires only linear calculations, somewhat reproduces the energetic properties of the turbulence.
+in Fig.~\ref{spec_nl_transient_gamma}, which we call the Transient and Turbulent Growth Rate Spectra. The $n=0$ curves are fairly similar, especially when compared to the eigenmode spectra, also
+shown in Fig.~\ref{spec_nl_transient_gamma}, although the transient curve is slightly down-shifted in $m$ compared to the turbulent curve. 
+The $n=1$ curves, on the other hand, are not very similar, although they trend with $m$ in the same way, which is very different from the linear eigenmode trend.
+The difference between the $n=1$ transient and turbulent growth rate spectra is due to nonlinear effects not captured by the simple transient calculation procedure. In particular, the turbulence
+exhibits strong energy transfer from $n=0$ to $n=1$ Fourier components which is then dissipated by electron-ion collisions as the electrons stream along the field lines~\cite{friedman2012b}. 
+This extra $n=1$ energy dissipation in the turbulent simulation is not a transient effect, and the $n=0$ to $n=1$ transfer is not a simple nonlinear mixing effect.
 
@@
 \centerline{\includegraphics[]{spec_nl_transient_gamma}}
-\caption{The linear, turbulent, and transient growth rate spectra. The transient growth rate spectra is calculated by the average growth rate over a time $2 \pi / \omega_*$ for the ensemble average
+\caption{The linear, turbulent, and transient growth rate spectra. The transient growth rate spectra is calculated by the average growth rate over a time $\omega_s^{-1}$ for the ensemble average
 growth ratio curves.}
@@
 
-These transient growth rates that comprise the curves in Fig.~\ref{spec_nl_transient_gamma} are more physically relevant to turbulence than the linear eigenmode growth rates.
-When the linear system under
-consideration is highly non-normal, transient growth rate analysis should be performed in place of (or in addition to) linear eigenmode analysis. Furthermore, the transient growth rates should
-be used in the kinds of predictions that are currently made with eigenmode growth rates. For instance, quasilinear theory uses the most unstable eigenmode growth rate $\gamma_{s,max}$ 
-to predict the turbulent saturation levels through the mixing length formula $\gamma_s/k_\perp^2$.
-Using this formula with $\gamma_{s,\rm{max}} = 0.002$ and $m=60$, one would predict a turbulent saturation
-level on the order of $0.1 \%$. On the other hand, using this formula with the transient values -- $\gamma_{t,\rm{max}} = 0.003, m=10$ -- we predict the saturation level on the order of $10 \%$,
+Nevertheless, our randomizing transient method, which requires only linear calculations, reproduces energetic growth properties of the turbulence that are not captured by eigenmode analysis.
+Clearly, when the linear systems are highly non-normal, transient growth rate analyses should be performed in place of (or in addition to) linear eigenmode analyses. 
+Furthermore, the transient growth rates should be used in the kinds of predictions that are currently made with eigenmode growth rates. 
+For instance, quasilinear theory uses the most unstable eigenmode growth rate $\gamma_{s,max}$ to predict the turbulent saturation levels through the mixing length formula $\gamma/k_\perp^2$. 
+Using this formula with $\gamma_{s,\rm{max}} = 0.002$ and $m=60$, one would predict a turbulent saturation level on the order of $0.1 \%$. 
+On the other hand, using this formula with the transient values -- $\gamma_{t,\rm{max}} = 0.003, m=10$ -- we predict the saturation level on the order of $10 \%$,
 which matches the saturation level in the nonlinear simulation.
+
+In summary, non-modal linear calculations are more informative than normal mode calculations when the linear system under consideration is highly non-normal. In the specific case of a low-flow
+LAPD experiment, a nonlinear instability with linear $k_\para=0$ drive dominates the energy dynamics of the turbulence. While this is a surprising result in light of normal mode analysis,
+it is not at all surprising given the system's non-modal behavior. In general, non-modal initial-value analysis is difficult to quantify and make predictive, but using some simple nonlinear modelling,
+we have shown that it is possible.
 

### 2013-12-11 Transient_growth_prl.tex
Dec 11

@@
 
-Normal mode analysis -- dealing with the eigenvalues and eigenvectors of a dynamical system -- has been used for the last two centuries to solve a number of problems ranging from
+Normal mode analysis -- the calculation of eigenvalues and eigenvectors of a linearized dynamical system -- has been used for the last two centuries to solve a number of problems ranging from
 heat conduction along a solid bar to quantum mechanical energy states of atoms. Despite its wide-ranging success, normal mode analysis has seemingly failed in particular instances,
 particularly in predicting the onset of turbulence in many hydrodynamic flows, where the turbulence is called subcritical. 
-The reason for this failure was not clearly explained until the early 1990's when Trefethen and others attributed the pitfalls of normal mode analysis to the non-normality of
-systems~\cite{trefethen1993,trefethen2005} -- a non-normal system is one whose linear matrix or operator does not commute with its adjoint. 
-Non-normal systems have eigenvectors that are nonorthogonal to one another. One consequence of non-normality is that even when all of the eigenvalues lie in the stable domain 
-(where all linear eigenvectors decay exponential), some fluctuations can transiently grow by accessing the free energy in the equilibrium gradients, allowing for sustained turbulence. 
+The reason for this failure was not clearly explained until the early 1990's when Trefethen and others attributed the pitfalls of normal mode analysis to the non-normality of linear operators of
+dynamical systems~\cite{trefethen1993,trefethen2005} -- a non-normal operator or matrix is one that does not commute with its adjoint and consequently has 
+eigenvectors that are nonorthogonal to one another. One consequence of eigenvector nonorthogonality is that even when all eigenvectors are stable and decay exponentially under linear evolution, 
+superpositions of eigenvectors can grow, albeit transiently.
 A paradigmatic illustration of this transient growth process for two nonorthogonal eigenvectors may be seen if Figure 2 of a review paper by Schmid~\cite{schmid2007}.
+Physically, this means that certain types of fluctuations of the laminar state can access free energy from background gradients even though normal mode fluctuations cannot.
+When combined with nonlinear effects, this allows for sustained subcritical turbulence.
 Such behavior is obscured by traditional normal mode analysis, which only effectively describes the long time asymptotic behavior of fluctuations under the 
-action of the linear operator. Transient events, such as those that can dominate nonlinear turbulent evolution, require initial-value (non-modal) calculations rather than normal mode calculations.
+action of the linear operator. Transient growth events, which can dominate nonlinear turbulent evolution, require initial-value (non-modal) calculations rather than normal mode calculations.
 
-Non-modal analysis has been embraced by the hydrodynamics community over the past two decades in the attempt to explain and predict subcritical turbulence, but the plasma community
+Non-modal analysis has been embraced by the hydrodynamics community over the past two decades in the attempt to explain and predict subcritical turbulence. But the plasma community
 still heavily relies on normal mode analysis to inform turbulent predictions and observations, with a few notable exceptions~\cite{camargo1998,camporeale2010,schekochihin2012}. 
-Additionally, non-modal treatments thus far have generally been used to explain rather than predict turbulent characteristics, 
-while the few predictions that have been made are often qualitative rather than quantitative. Perhaps this is why non-modal analysis has not received greater attention.
-Therefore, this paper takes up the task of developing a quantitative approach to predicting turbulent properties using only non-modal linear calculations. 
+Furthermore, non-modal treatments thus far have generally been explanatory rather than predictive and have centered around the transition to turbulence in subcritical systems rather 
+than on properties of fully-developed turbulence.
+This paper, in contrast, takes up the task of developing a quantitative approach to predicting turbulent properties using only non-modal linear calculations. 
 The basic approach we take is to solve the linear initial value problem using an ensemble of random initial conditions, 
-and then calculate the average growth rate of the solutions up to one nonlinear decorrelation time, which we take to be the characteristic time of the linear process. 
-The effective growth rate spectrum produced by this technique can then be used in a quasilinear fashion to gain more information about the turbulence.
-While this technique is general and can be applied to many nonlinear dynamical systems, we restrict our treatment to one particular model that relates to one particular experiment. Not only does
-this allow us to thoroughly explain and validate this technique, but it also allows us to attribute transient growth effects to physically realizable turbulence.
+and then calculate the average growth rate of the solutions up to one nonlinear decorrelation time. We take this nonlinear decorrelation time to be a characteristic \emph{linear} time.
+This procedure produces an effective growth rate spectrum that can be used to predict turbulent properties such as saturation levels.
+While the concepts behind this technique are general enough to be applied to many nonlinear dynamical systems, the details will vary from system to system, 
+so we restrict our treatment to one particular turbulence model.
 
-The experiment considered is a drift wave turbulence experiment conducted in the Large Plasma Device (LAPD)~\cite{gekelman1991}. LAPD is a linear machine with a straight magnetic field $\mathbf{B}$.
-Due to its large dimensions and high collisionality, LAPD is suitable for fluid modelling. We use a reduced Braginskii~\cite{Braginskii1965} 2-fluid model:
+The model is designed to describe a drift wave turbulence experiment conducted in the Large Plasma Device (LAPD)~\cite{gekelman1991}.
+LAPD is a linear machine with a straight magnetic $\mathbf{B}$ field.
+Due to its large dimensions and high collisionality, LAPD is apt for fluid modelling. Thus, we use a reduced Braginskii~\cite{Braginskii1965} 2-fluid model:
 
@@
 where $N_0$ and $N$ are the equilibrium and fluctuating density, $\vpe$ is the fluctuating parallel electron velocity, $\varpi \equiv \gradperp \cdot (N_0 \gradperp \phi)$ is the potential vorticity
-of the fluctuating potential $\phi$, and $T_{e0}$ and $T_e$ are the equilibrium and fluctuating electron temperature. The equations are developed with Bohm normalizations: lengths are
+of the fluctuating potential $\phi$, and $T_{e0}$ and $T_e$ are the equilibrium and fluctuating electron temperature. The equations employ Bohm normalizations: lengths are
 normalized to the ion sound gyroradius $\rho_s$, times to the ion cyclotron time $\omega_{ci}^{-1}$, velocities to the sound speed $c_s$, densities to the equilibrium peak density, and electron
-temperatures and potentials to the equilibrium peak electron temperature. The profiles $N_0$ and $T_{e0}$ and other parameters are all taken from experimental measurements. Note that $\phi_0$ is
-neglected because we model only one experiment in which the equilibrium radial electric field was nulled out by boundary biasing~\cite{schaffner2012}.
+temperatures and potentials to the equilibrium peak electron temperature. The profiles $N_0$ and $T_{e0}$ and other parameters are all taken from experimental measurements. $\phi_0$ 
+does not appear in the equations because we model one particular experiment in which the equilibrium radial electric field was nulled by boundary biasing~\cite{schaffner2012}.
 
-The equations are global and partially linearized -- the only nonlinearities we retain are the advective nonlinearities in the Poisson brackets. 
-We also add artificial diffusion and viscosity terms with small numerical coefficients ($\mu_N, \mu_\phi, \mu_T = 10^{-3}$), 
+The equations are global and partially linearized; the only nonlinearities retained are the advective nonlinearities in the Poisson brackets. 
+We add artificial diffusion and viscosity terms with small numerical coefficients ($\mu_N, \mu_\phi, \mu_T = 10^{-3}$), 
 used for numerical stability in nonlinear simulations, which are performed with the BOUT++ code~\cite{dudson2009}. We use periodic axial and zero value radial 
@@
 though we mention here that the model reproduces all statistical properties of the experimental turbulence within a factor of two, and in most cases, much better than that.
-Of particular interest, was the finding that the sustainment of the turbulence in the nonlinear simulation is dominated by a nonlinear instability process~\cite{friedman2012b,friedman2013}.
-This nonlinear instability process -- discovered by Drake et al.~\cite{drake1995} in simulations using a similar plasma model to ours -- works as follows: long $k_\para=0$ convective filaments transport
-density (and temperature to a lesser extent) across the equilibrium density profile, setting up $k_\para=0$ density filaments. These filaments are unstable to drift waves causing a secondary drift
-wave instability to grow on the filaments. These drift waves, which have finite $k_\para$ and propagate primarily radially, nonlinearly couple to one another and reinforce the convective filaments.
+
+The nonlinear simulation reveals a fascinating property of the turbulence -- it is dominated by a nonlinear instability process despite being linearly unstable to drift waves
+~\cite{friedman2012b,friedman2013}.
+This nonlinear instability process, which was discovered by Drake et al.~\cite{drake1995} using a similar plasma model to ours, works as follows: 
+long ($k_\para=0$) convective filaments transport density across the equilibrium density profile, setting up $k_\para=0$ density filaments (the same applies to the temperature). 
+These filaments are unstable to drift waves causing a secondary drift wave instability to grow on the filaments. 
+These drift waves, which have finite $k_\para$ and propagate primarily radially, nonlinearly couple to one another and reinforce the convective filaments.
+
+Although the instability is called a nonlinear instability, the first part of the mechanism -- the transport of background density by the convective filaments -- is a linear one.
+In fact, the other parts of the mechanism are driven by energetically conservative nonlinear interactions, 
+meaning that the convective transport is the only step responsible for energy injection into the fluctuations from the laminar state.
+Deriving an equation for the evolution of the energy from Eqs.~\ref{ni_eq}-\ref{te_eq}~\cite{friedman2012b,friedman2013}, we may symbolically write
+
+\beq
+\label{dEdt_def}
+\diff{E(m,n,t)}{t} = Q(m,n,t) + D(m,n,t) + \sum_{m',n'} T(m,m',n,n',t).
+\eeq
+In this equation, $m$ and $n$ represent the azimuthal and axial Fourier mode numbers, respectively. 
+Since Eqs.~\ref{ni_eq}-\ref{te_eq} are not explicitly dependent on $\theta$ and $z$, the energy equation is separable for every $m$ and $n$. 
+Futher, $Q$ represents the linear energy exchange with the equilibrium (convective transport), $D$ the dissipation from irreversible processes
+like collisions, and $T$ the nonlinear triad wave transfer. Since $\int f \{g,f\} dV = 0$, or equivalently $\sum_{m,m',n,n'} T(m,m',n,n',t)=0$, 
+the Poisson bracketed advective nonlinearities in the model are energetically conservative and therefore do no inject or dissipate energy on the whole.
+Consequently, the fluctuation energy originates completely from $Q$, the linear convective transport mechanism.
+
+\begin{figure}
+\centerline{\includegraphics[]{transient_n0_n1_growth}}
+\caption{The linear evolution of the energy of several Fourier modes starting from a turbulent initial state. The $n=0$ curves have an initial period of transient growth before exponentially decaying,
+while the $n=1,m=20$ curve transiently decays before growing exponentially.}
+\label{transient_n0_n1_growth}
+\end{figure}
+
+This convective transport of density filaments is akin to the paradigmatic ``lift-up'' mechanism in hydrodynamic shear flows whereby streamwise vortices drive streamwise streaks~\cite{krommes1999}.
+Both are transient growth processes. We see this in our simulations by following the evolution of the energy of several Fourier $m,n$ modes after turning off the nonlinearities in an already
+turbulent simulation. A few representative $m,n$ modes are shown in Fig.~\ref{transient_n0_n1_growth}. The transient growth of the filamentary structures is seen in the $n=0$ modes, 
+which grow transiently before decaying exponentially at the rate of their least stable eigenmode. This kind of behavior is rather unituitive to some because all $n=0$ linear eigenmodes are stable, 
+and the normal mode paradigm has conditioned us to believe that a linear superposition of such eigenmodes should decay. But as the hydrodynamics community discovered, linear
+superpositions of nonorthogonal vectors may behave unintuitively.
+Notice now that the $n=1,m=20$ mode decays transiently before growing exponentially with growth rate of the most unstable eigenmode at $n=1,m=20$. This isn't quite as interesting as the $n=0$
+behavior, and it can occur in normal systems. Futhermore, it has no parallel in truly subcritical turbulence, but it can be helpful for interpreting later results.
+
+We stress that Fig.~\ref{transient_n0_n1_growth} is obtained by \emph{linear} evolution from an initial turbulent state.
+Since the transient growth, which is behind subcritical turbulence, is a purely linear phenomenon, it has long been a goal of researchers to predict the onset of subcritical turbulence using
+only linear, non-modal calculations. It is our goal, however, to go a step further and use linear calculations to understand and even predict turbulent properties.
+Such calculations would be much faster and easier to interpret than direct simulations of nonlinear equations. Perhaps one of the most useful statistical properties that we can use to describe
+the turbulence -- and one that is conveniently ammenable to linear analysis -- is the turbulent growth rate spectrum.
 
@@
 
-Although the instability mechanism is nonlinear, the first part of the mechanism -- the transport of density by the convective filaments -- is a linear one. 
-In fact, it is surprisingly similar to the so-called "lift-up'' mechanism often seen in subcritical hydrodynamic shear flows, for which transient growth arguments were first applied~\cite{trefethen1993},
-making it particularly apt to explore in a non-modal framework. Furthermore, since the other steps of the nonlinear instability
-are controlled by nonlinear interactions, the convective transport is the only step responsible for energy injection and dissipation from the laminar state. This fact may be seen from simple
-energetic principles. Starting from Eqs.~\ref{ni_eq}-\ref{te_eq},
-an equation for the evolution of the energy can be obtained~\cite{friedman2012b,friedman2013}. The equation takes the form:
-
-\beq
-\label{dEdt_def}
-\diff{E(m,n,t)}{t} = Q(m,n,t) + D(m,n,t) + \sum_{m',n'} T(m,m',n,n',t).
-\eeq
-Here, $m$ and $n$ are the azimuthal and axial Fourier mode numbers, respectively. Note that since Eqs.~\ref{ni_eq}-\ref{te_eq} are not explicitly dependent on $\theta$ and $z$, the energy equation
-is separable for every $m$ and $n$. Furthermore, $Q$ represents the linear energy exchange with the equilibrium, $D$ is the dissipation from irreversible processes
-like collisions, and $T$ accounts for the nonlinear triad wave transfer. Because $T$ results from the Poisson bracketed advective nonlinearities in the model and $\int f \{g,f\} dV = 0$, 
-$\sum_{m,m',n,n'} T(m,m',n,n',t)=0$, meaning the nonlinearities cause no net energy injection or dissipation. In other words, the fluctuation energy originates completely from the linear terms.
-
-The expression $\diff{E(m,n,t)}{t}\big|_{\rm{lin}} = Q(m,n,t) + D(m,n,t)$ accounts for the total net energy injection at each $m,n$ into the fluctuations. 
-An effective instantaneous growth rate $\gamma_{\rm{eff}}(m,n,t) = \diff{E(m,n,t)}{t}\bigg|_{\rm{lin}}/ 2 E(m,n,t)$ can then be defined. 
-This effective growth rate is calculated from the spatial structure of the fluctuations, so it has meaning for turbulent fluctuations as well as those obtained from a linear simulation. If calculated
+To obtain the turbulent growth rate spectrum from the nonlinear simulation, we begin with the expression $\diff{E(m,n,t)}{t}\big|_{\rm{lin}} = Q(m,n,t) + D(m,n,t)$, 
+which accounts for the total net energy injection at each $m,n$ into the fluctuations. 
+An effective instantaneous growth rate $\gamma_{\rm{eff}}(m,n,t) = \diff{E(m,n,t)}{t}\big|_{\rm{lin}}/ 2 E(m,n,t)$ can then be defined. 
+This effective growth rate is calculated from the spatial structure of the fluctuations, so it is well-defined for both turbulent fluctuations and eigenmode fluctuations. If calculated
 from a linear simulation run for a sufficiently long time, $\gamma_{\rm{eff}}(m,n)$ is the same as the eigenmode growth rate of the least stable linear eigenmode for each $m,n$.
-We plot this effective growth rate in Fig.~\ref{spec_nl_gamma} for $n=0$ and $n=1$ components. The turbulent growth rate curves are calculated from this formula, 
-averaged over a time of about $3 \times 10^{4} \omega_{ci}^{-1}$ during the saturated turbulent phase of the nonlinear simulation --$\gamma_{\rm{Turb}}(m,n) = \int \gamma_{\rm{eff}}(m,n,t) dt$. 
+For turbulence, $\gamma_{\rm{eff}}(m,n)$ is generally not equal to an eigenmode growth rate, especially in the case where the system is non-normal.
+We plot this growth rate in Fig.~\ref{spec_nl_gamma} for $n=0$ and $n=1$ components. The turbulent growth rate curves are calculated from this formula, 
+averaged over a time of about $3 \times 10^{4} \ \omega_{ci}^{-1}$ during the saturated turbulent phase of the nonlinear simulation: $\gamma_{\rm{Turb}}(m,n) = \int \gamma_{\rm{eff}}(m,n,t) dt$. 
 We show this turbulent growth rate along with the $1 \sigma$ contour, which indicates how greatly $\gamma_{\rm{eff}}(m,n,t)$ varies over time.
@@
 
-From Fig.~\ref{spec_nl_gamma}, one sees that the linear eigenmodes are unstable for $n=1$ and stable for $n=0$ as expected for drift waves. 
-However, the turbulent growth rates are positive for $n=0$ and low $m$, but negative for $n=1$ at all $m$. 
-The net injection of the $n=0$ components is reminiscent of subcritical turbulence, where there is turbulent sustainment in the absense of linear instability.
-Moreover, the $n=0$ energy injection is a manifestation of the convective filaments drawing energy from the equilibrium density gradient and depositing this energy into the density fluctuations,
-which again, is similar to the hydrodynamic ``lift-up'' mechanism associated with subcritical hydrodynamic shear-flow turbulence.
-It it prudent then to call the turbulence in the LAPD simulation subcritical or at least subcritical-like.
+As already stated, one sees from Fig.~\ref{spec_nl_gamma} that the linear eigenmodes are unstable for $n=1$ and stable for $n=0$ as expected for drift waves. 
+However, the turbulent growth rates are positive for $n=0$ and low $m$ (which is subcritical-like), but negative for $n=1$ at all $m$. 
+There was already a hint of this from the initial behavior of the curves in Fig.~\ref{transient_n0_n1_growth}, but this is a quantitative and more thorough picture.
+Again, the $n=0$ energy injection is a manifestation of the convective filaments drawing energy from the equilibrium density gradient and depositing this energy into the density fluctuations.
 
-Like subcritical hydrodynamic turbulence, the cause of subcritical LAPD turbulence (the positive growth rate of the $n=0$ structures) 
-is a transient growth mechanism due to nonorthogonal linear eigenmodes. The transient growth can be seen clearly in Fig.~\ref{transient_n0_n1_growth},
-where we follow the evolution of the energy of several Fourier $m,n$ modes after turning off the nonlinearities in the turbulent simulation. The $n=0$ modes grow for some time before decaying
-exponentially at the rate of their least stable eigenmode. Conversely, the $n=1,m=20$ mode decays transiently before growing exponentially. This is not an attribute of subcritical turbulence
-because there are no unstable eigenmodes in truly subcritical turbulence, but it is nevertheless interesting and is associated with a transient event.
-The growth (or decay) rate of these curves at $t=0$ is necessarily equal to the growth rate that is used in the calculation of Fig.~\ref{spec_nl_gamma}. 
-This $t=0$ growth rate, however, is just one number that goes into that calculation. The calculation is based on the turbulence over a long time range; this is just one instant of time.
-In any case, the $t=0$ behavior of the curves in Fig.~\ref{transient_n0_n1_growth} generally corroborates with the effective growth rates in Fig.~\ref{spec_nl_gamma}.
+Our goal now is to reproduce the turbulent growth rate curves in Fig.~\ref{spec_nl_gamma} without performing nonlinear simulations. This is difficult because the growth rates depend on turbulent
+structures, which for non-normal systems, may be very different from eigenmode structures, and which change in time due to transient evolution.
+It is clear, however, that the turbulent structures at any given time are a result of both linear evolution and nonlinear energy transfer. 
+If we imagine that the nonlinearities randomize the turbulent structures at each wavenumber while the linearities evolve the structures
+deterministically, we can develop a procedure for choosing proper spatial structures that may be used in calculating effective growth rates.
 
-\begin{figure}
-\centerline{\includegraphics[]{transient_n0_n1_growth}}
-\caption{The linear evolution of the energy of several Fourier modes starting from a turbulent initial state. The $n=0$ curves have an initial period of transient growth before exponentially decaying,
-while the $n=1,m=20$ curve transiently decays before growing exponentially.}
-\label{transient_n0_n1_growth}
-\end{figure}
-
-We stress again that Fig.~\ref{transient_n0_n1_growth} is obtained by \emph{linear} evolution from an initial turbulent state.
-Since the transient growth (and decay) is a purely linear phenomenon, it appears possible to predict turbulent growth rates like those of Fig.~\ref{spec_nl_gamma} using only linear calculations.
-Such calculations would be much faster and easier to interpret than direct simulations of nonlinear equations. However, the difficulty in such calculations lies in the dependence of the transient
-evolution on the initial conditions and the change in the growth rate as a function of time as the linear evolution progresses. To overcome this difficulty, we propose a statistical procedure
-that linearly evolves an ensemble of random initial conditions and calculates the average growth rate of the evolution over a characteristic linear time.
-
-To elucidate this procedure, we take Eqs.~\ref{ni_eq}-\ref{te_eq}, remove the nonlinearities and sources and Fourier decompose in the azimuthal and axial directions.
+To elucidate this procedure, we begin by taking Eqs.~\ref{ni_eq}-\ref{te_eq} and Fourier decomposing in the azimuthal and axial directions.
 Then, we discritize in the radial direction ($r \rightarrow r_0, r_1, \ldots, r_N $), and approximate radial derivatives with finite differences. 
@@
 \beq
-\label{lin_eq}
-\mathbf{B}_{m,n} \diff{\mathbf{v}_{m,n}(t)}{t} = \mathbf{C}_{m,n} \mathbf{v}_{m,n}(t),
+\label{mat_eq}
+\mathbf{B}_{m,n} \diff{\mathbf{v}_{m,n}(t)}{t} = \mathbf{C}_{m,n} \mathbf{v}_{m,n}(t) - \sum_{m',n'}  \mathbf{v}_{E,m-m',n-n'} \cdot \gradperp \left( \mathbf{B}_{m',n'} \mathbf{v}_{m',n'}(t) \right),
 \eeq
-where $\mathbf{v}_{m,n} = \left( N(r_0), N(r_1), \ldots, \vpe(r_0), \vpe(r_1), \ldots, \phi(r_0), \phi(r_1), \ldots, T_e(r_0), T_e(r_1), \ldots \right)_{m,n}^{T}$,
+where $\mathbf{v}_{m,n} = \left( N(r_0), N(r_1), \ldots, \vpe(r_0), \vpe(r_1), \ldots, \phi(r_0), \phi(r_1), \ldots, T_e(r_0), T_e(r_1), \ldots \right)_{m,n}^{T}$ is the state of the system,
 and $\mathbf{B}_{m,n}$ and $\mathbf{C}_{m,n}$ are coefficient matrices that include the equilibrium information and finite difference coefficients.
-The equations for different $m$ and $n$ are separable because the equations are linear and the eigenfunctions have azimuthal and axial Fourier dependence resulting from the lack of explicit
-dependence on $\theta$ and $z$ in the equations. Thus, eigenmodes with different $m$ and $n$ are orthogonal and linearly non-interacting. Note that for each $m,n$, there exist $4 \times N_r$ linearly
+The first term on the RHS represents the linearities,
+and the second term the nonlinearities.
+The equations for different $m$ and $n$ are separated, although the nonlinearity couples the different $m$'s and $n$'s. 
+In other words, eigenmodes with different $m$ and $n$ are orthogonal and linearly non-interacting. Note that for each $m,n$, there exist $4 \times N_r$ linearly
 independent, but nonorthogonal eigenvectors. Hence forth, we drop the $m,n$ subscripts.
@@
 where $^*$ stands for the conjugate transpose, and $\mathbf{M}$ is a positive-definite matrix. $\mathbf{M}$ should be chosen such that 
-$||\mathbf{u}|| = \sqrt{\left< \mathbf{u},\mathbf{u} \right>} = \sqrt{E}$ -- that is, the norm of the state vector is the square root of the energy. Furthermore, it is often convenient in
+$||\mathbf{v}|| = \sqrt{\left< \mathbf{v},\mathbf{v} \right>} = \sqrt{E}$. That is, the norm of the state vector is the square root of the energy. Furthermore, it is often convenient in
 computations to use the $L_2$-norm, $||\mathbf{u}||_2 = \sqrt{\sum_i |u_i|^2}$.
 These requirements can be accomplished through the change of variables $\mathbf{u} = \mathbf{M}^{\frac{1}{2}} \mathbf{v}$.
-Then Eq.~\ref{lin_eq} becomes
+Then the linear portion of Eq.~\ref{mat_eq} becomes
 
@@
 \label{lin_eq_A}
-\diff{\mathbf{u}}{t} = \mathbf{A} \mathbf{u},  \quad \rm{where} \ \mathbf{A} = \mathbf{M}^{-\frac{1}{2}} \mathbf{B}^{-1} \mathbf{C} \mathbf{M}^{\frac{1}{2}}.
+\diff{\mathbf{u}}{t} = \mathbf{A} \mathbf{u},  \quad \rm{where} \ \mathbf{A} = \mathbf{M}^{\frac{1}{2}} \mathbf{B}^{-1} \mathbf{C} \mathbf{M}^{-\frac{1}{2}}.
 \eeq
-The eigenvalues of $\mathbf{A}$ are the same as those of $\mathbf{B}^{-1} \mathbf{C}$ because $\mathbf{M}^{-\frac{1}{2}} \mathbf{B}^{-1} \mathbf{C} \mathbf{M}^{\frac{1}{2}}$ is a similarity transformation.
+The eigenvalues of $\mathbf{A}$ are the same as those of $\mathbf{B}^{-1} \mathbf{C}$ because $\mathbf{M}^{\frac{1}{2}} \mathbf{B}^{-1} \mathbf{C} \mathbf{M}^{-\frac{1}{2}}$ is a similarity transformation.
 
@@
 This allows for the possibility of exciting subcritical turbulence, which is where non-modal analysis is most often applied. 
-However, it can be misleading to study only $G_{\rm{max}}(t)$ in instances where turbulence is already known to exist and one is interested in predicting specific properties of turbulence. 
-The main reason is that $G_{\rm{max}}(t)$ is actually the upper envelope of all possible $G(t)$ curves.
-In other words, no one particular initial condition $\mathbf{u}(0)$ evolves along $G_{\rm{max}}(t)$.
-One particular $\mathbf{u}_\tau(0)$ will evolve such that  $G(\tau)= G_{\rm{max}}(\tau)$. Another $\mathbf{u}_\rho(0)$ will evolve such that $G(\rho)= G_{\rm{max}}(\rho)$. 
+However, it can be misleading to study only $G_{\rm{max}}(t)$ when predicting specific properties of turbulence because
+$G_{\rm{max}}(t)$ is only the upper envelope of all possible $G(t)$ curves. No one particular initial condition $\mathbf{u}(0)$ evolves along $G_{\rm{max}}(t)$. 
 Furthermore, it isn't obvious what kind of spatial structure or structures will come to dominate a turbulent system. 
-Unlike in the normal case, in the non-normal case, optimal structures don't amplify themselves. They amplify the energy, but evolve into different structures, at least until late times.
+In non-normal systems, unlike in the normal systems, optimal structures don't amplify themselves. Energy increases, but structures evolve in a complicated manner.
 
@@
 The key to quantifying non-modal analysis and making it predictive is to successfully model the effect that the nonlinearities have on the transient linear processes. 
-In this regard, we model the nonlinearities as a periodic randomizing force. That is, the nonlinearities randomize the turbulent spatial structures on a characteristic nonlinear time scale. 
-In practice, we completely randomize the system (initial condition) and then let it evolve linearly for a time $t_{nl}$, at which point, the system is re-randomized.
+To this effect, we first note that the advective nonlinearity in Eq.~\ref{mat_eq} presents a characteristic time scale $\tau_{\rm{nl}} \sim (v_E k_\perp)^{-1}$. This nonlinear
+time scale is generally associated with the eddy turnover time or the time on which eddies become decorrelated. We therefore heuristically model the nonlinearities 
+as a randomizing force that acts on this characteristic nonlinear time scale.
+Specifically, we start with a random initial condition, let it evolve linearly for a time $\tau_{\rm{nl}}$, and then re-randomized the system.
+In other words, we take the curves in Fig.~\ref{m20n0_time_evs} b) and cut them off after a time $\tau_{\rm{nl}}$.
 While the nonlinearities obviously do not affect the system in discrete time steps, this simplification provides a way to use linear non-modal analysis in a predictive manner.
-Since we do not know, \emph{a priori} at what level the turbulence will saturate, 
-we cannot calculate the relative magnitude of the nonlinear terms in the equations to get the nonlinear time scale $t_{nl}$.
+
+Now, since we do not know, \emph{a priori} at what level the turbulence will saturate, we cannot calculate the relative magnitude of $\tau_{\rm{nl}}$.
 Thus, we invoke the conjecture of \emph{critical balance}, which posits that the nonlinear time scale equals the linear time scale at all spatial scales~\cite{schekochihin2012}. 
-We can then estimate the nonlinear time scale as $t_{nl} = \omega_s^{-1}$, the inverse of the linear frequency of the fastest growing linear eigenmode (at each $m,n$).
+We can then estimate the nonlinear time scale as $\tau_{\rm{nl}} = \omega_{\rm{lin}}^{-1}$, the inverse of the linear frequency (at each $m,n$).
+There are several choices of linear frequencies that we can use. Since we are dealing with a system of drift waves, $\omega_* = \frac{T_e m}{L_n r m_i \omega_{ci}^2}$ seems suitable, however,
+$\omega_*$ is radially dependent, so it is not obvious which radius to choose. Alternatively, the linear frequency of the fastest growing eigenmode $\omega_s$ may be used. This is not a function
+of radius, but $\omega_s = 0$ for $n=0$ eigenmodes, so one is forced to use $\omega_s$ for the fastest growing $n=1$ eigenmode at each $m$ to get a meaningful time scale. This is not wholly
+unjustified, however, due to the inherent nonlinear coupling of different modes, which is evident in Eq.~\ref{mat_eq}.
 
-In Fig.~\ref{m20n0_time_evs} c), we show the randomly initialized linear evolution curves from start up until a time of $\omega_s^{-1}$. 
+To illustrate our procedure further, we show in Fig.~\ref{m20n0_time_evs} c) the randomly initialized linear evolution curves from start up until a time of $\omega_s^{-1}$. 
 Then, in Fig.~\ref{m20n0_time_evs} d), we calculate the instantaneous growth rate as a function of time of the ensemble averaged curve over that time period.
-For comparison, the instantaneous growth rate of the maximum growth curve is also shown.
-In theory, this method should produce ensembled-averaged growth rate curves (like that in Fig.~\ref{m20n0_time_evs} d)) that have instantaneous growth rates consistent with the instantaneous
-turbulent growth rates. Both the instantaneous transient growth rates and the turbulent growth rates have an average and a range associated with them. We show a comparison of their averages
-in Fig.~\ref{spec_nl_transient_gamma}, which we call the Transient and Turbulent Growth Rate Spectra. The $n=0$ curves are fairly similar, especially when compared to the eigenmode spectra, also
-shown in Fig.~\ref{spec_nl_transient_gamma}, although the transient curve is slightly down-shifted in $m$ compared to the turbulent curve. 
+The average of this instantaneous growth rate curve (the dashed line) gives the average ``Transient Growth Rate.''
+In Fig.~\ref{spec_nl_transient_gamma}, we compare the Transient Growth Rate Spectra to the Turbulent Growth Rate Spectra, where we use both $\omega_*^{-1}$ and $\omega_s^{-1}$ for $\tau_{\rm{nl}}$.
+The $n=0$ curves in Fig.~\ref{spec_nl_transient_gamma} a) are fairly similar, especially when compared to the eigenmode spectra.
 The $n=1$ curves, on the other hand, are not very similar, although they trend with $m$ in the same way, which is very different from the linear eigenmode trend.
@@
 This extra $n=1$ energy dissipation in the turbulent simulation is not a transient effect, and the $n=0$ to $n=1$ transfer is not a simple nonlinear mixing effect.
+
+There is one more growth rate curve in Fig.~\ref{spec_nl_transient_gamma}, labelled ``Trans $\omega_s^{-1}$ cascade.'' We include this to address the apparent downshift (in $m$) of
+the transient $\omega_s^{-1}$ curve in comparison to the turbulent curve. The downshift is a consequence of our oversimplification of the nonlinearity. Specifically, we
+used $\omega_s^{-1}$ at each $m$ for the nonlinear time scale at that same $m$. As seen in Eq.~\ref{mat_eq}, $v_{m,n}$ is affected by every $\mathbf{v}_{E,m-m',n-n'} k^{'}_\perp$, not by only
+$\mathbf{v}_{E,m,n} k_\perp$. Since we do not want to overcomplicate matters, we address this by posing a simple forward cascade model for the nonlinearity. That is, $v_{m,n}$ should be
+most affected by the term with $m' = m/2 \rightarrow m-m' = m/2$, corresponding to eddy mitosis.
+Then, $\mathbf{v}_{E,m-m',n-n'} k^{'}_\perp \sim \mathbf{v}_{E,m/2} k^{'}_\perp/2 \sim \omega_{s,m/2}^{-1}$,
+so we use $\omega_s^{-1}$ at $m/2$ for the randomizing time scale at $m$ in deriving the ``Trans $\omega_s^{-1}$ cascade'' curves in Fig.~\ref{spec_nl_transient_gamma}. 
+The $n=0$ curve matches well with the $n=0$ Turbulent Growth Rate curve at $m>20$, but poorly for $m<20$, which is expected because the cascade applies only at high $m$. \emph{A priori}, we
+cannot determine the minimum $m$ for which the cascade should apply, so this method is simply explanatory.
 

### 2013-12-15 Transient_growth_prl.tex
Dec 13

@@
 
-\title{Non-modal Growth in LAPD Turbulence}
+\title{A Linear Technique to Predict Non-Normal LAPD Turbulence}
 
@@
 \begin{abstract}
-Large Plasma Device (LAPD) [W. Gekelman \emph{et al.}, Rev. Sci. Inst. {\bf 62}, 2875 (1991)] 
+In dynamical systems with highly nonorthogonal linear eigenvectors, non-modal linear analysis can be much more useful than normal-mode analysis in predicting turbulent properties. 
+However, the non-trivial time evolution of non-modal structures makes quantitative prediction difficult. We present a technique using only linear calculations to overcome
+this difficulty by modelling the effect that the advective nonlinearities have on spatial turbulent structures. Specifically, we model the nonlinearities as a periodic randomizing force with
+period consistent with critical balance arguments. We apply this technique to a model of drift wave turbulence in the Large Plasma Device (LAPD) 
+[W. Gekelman \emph{et al.}, Rev. Sci. Inst. {\bf 62}, 2875 (1991)], where non-modal effects dominate the turbulent system. 
+We compare the resulting growth rate spectrum to that obtained from a nonlinear simulation, showing good quantitative agreement.
 \end{abstract}
@@
 
-It is common practice in normal mode analysis to look for the fastest growing eigenmode, the idea being that eventually the fastest growing structure will overcome the others.
+It is common practice in normal mode analysis to look for the least stable eigenmode, the idea being that eventually the fastest growing structure will overcome the others.
 For the non-modal case, it is common to study the properties of $G_{\rm{max}}(t)$, because if $G_{\rm{max}}(t) > 1$ at any time, there is a possibility of amplification, even if it is only transient.
@@
 Furthermore, it isn't obvious what kind of spatial structure or structures will come to dominate a turbulent system. 
-In non-normal systems, unlike in the normal systems, optimal structures don't amplify themselves. Energy increases, but structures evolve in a complicated manner.
+In non-normal systems, unlike in the normal systems, optimal structures don't amplify themselves, rather, they evolve while increasing the total fluctuating energy.
 
-We illustrate $G_{\rm{max}}(t)$ along with $G(t)$ for several random initial conditions in Fig.~\ref{m20n0_time_evs} a), all at $n=0$ and $m=20$. First, despite the fact that all eigenmodes at this
-wavenumber are stable, there is the possibility for transient amplification by a factor of $20$. Second, a sample of randomly initialized vectors are all transiently amplified, but none display
+We illustrate $G_{\rm{max}}(t)$ along with $G(t)$ for several random initial conditions in Fig.~\ref{m20n0_time_evs} a), all at $n=0$ and $m=20$. Despite the fact that all eigenmodes at this
+wavenumber are stable, there is the possibility for transient amplification by a factor of $20$. Further, a sample of randomly initialized vectors are all transiently amplified, but none display
 optimal growth as compared to $G_{\rm{max}}(t)$. We show a zoomed-in view of these randomly initialized curves in Fig.~\ref{m20n0_time_evs} b).
-From the figure, it appears extremely unlikely to choose a random initial vector that either optimally amplifies the energy at any time or one that monotonically decays, which would be the case
+It appears extremely unlikely to choose a random initial vector that either optimally amplifies the energy at any time or one that monotonically decays, which would be the case
 if the initial vector were one of the linear eigenvectors.
@@
 
-It is not surprising that non-modal analysis has remained a qualitative endeavor due to its intrinsic dependence on initial conditions and the time-dependent growth properties of initial-value analysis.
+Non-modal analysis has likely remained a qualitative endeavor due to the time-dependent growth properties of initial-value analysis as well as its intrinsic dependence on initial conditions.
 The key to quantifying non-modal analysis and making it predictive is to successfully model the effect that the nonlinearities have on the transient linear processes. 
-To this effect, we first note that the advective nonlinearity in Eq.~\ref{mat_eq} presents a characteristic time scale $\tau_{\rm{nl}} \sim (v_E k_\perp)^{-1}$. This nonlinear
-time scale is generally associated with the eddy turnover time or the time on which eddies become decorrelated. We therefore heuristically model the nonlinearities 
+To this effect, we first note that the advective nonlinearity in Eq.~\ref{mat_eq} has the form of the state vector divided by a time $\tau_{\rm{nl}} \sim (v_E k_\perp)^{-1}$. This nonlinear
+time scale is generally associated with the eddy turnover time or the time in which eddies become decorrelated. In light of this, we present a heuristical model of the nonlinearities 
 as a randomizing force that acts on this characteristic nonlinear time scale.
 Specifically, we start with a random initial condition, let it evolve linearly for a time $\tau_{\rm{nl}}$, and then re-randomized the system.
-In other words, we take the curves in Fig.~\ref{m20n0_time_evs} b) and cut them off after a time $\tau_{\rm{nl}}$.
 While the nonlinearities obviously do not affect the system in discrete time steps, this simplification provides a way to use linear non-modal analysis in a predictive manner.
+In other words, there are certain spatial structures associated with the curves in Fig.~\ref{m20n0_time_evs} b) that change as a function of time. We consider only those structures from $t=0$, where
+the structures are completely random, to $t = \tau_{\rm{nl}}$, and associate those structures with turbulent structures. These provide us with all we need to characterize the turbulence.
 
-Now, since we do not know, \emph{a priori} at what level the turbulence will saturate, we cannot calculate the relative magnitude of $\tau_{\rm{nl}}$.
+Now, since we do not know, \emph{a priori} at what level the turbulence will saturate, we cannot calculate the relative magnitude of $\tau_{\rm{nl}}$ from $(v_E k_\perp)^{-1}$ directly.
 Thus, we invoke the conjecture of \emph{critical balance}, which posits that the nonlinear time scale equals the linear time scale at all spatial scales~\cite{schekochihin2012}. 
@@
 The average of this instantaneous growth rate curve (the dashed line) gives the average ``Transient Growth Rate.''
-In Fig.~\ref{spec_nl_transient_gamma}, we compare the Transient Growth Rate Spectra to the Turbulent Growth Rate Spectra, where we use both $\omega_*^{-1}$ and $\omega_s^{-1}$ for $\tau_{\rm{nl}}$.
+In Fig.~\ref{spec_nl_transient_gamma}, we compare the Transient Growth Rate Spectra to the Turbulent Growth Rate Spectra, where we use $\omega_s^{-1}$ for $\tau_{\rm{nl}}$.
 The $n=0$ curves in Fig.~\ref{spec_nl_transient_gamma} a) are fairly similar, especially when compared to the eigenmode spectra.
@@
 \centerline{\includegraphics[]{spec_nl_transient_gamma}}
-\caption{The linear, turbulent, and transient growth rate spectra. The transient growth rate spectra is calculated by the average growth rate over a time $\omega_s^{-1}$ for the ensemble average
-growth ratio curves.}
+\caption{The linear, turbulent, and transient growth rate spectra.
+The transient growth rate spectra are calculated by the average growth rate over a time $\omega_s^{-1}(m,n=1)$ for the ensemble average
+growth ratio curves with the cascade curve using $\omega_s^{-1}$ at $m/2$ rather than $m$.}
 \label{spec_nl_transient_gamma}

### 2013-12-16 Transient_growth_prl.tex
Dec 15

@@
 \begin{abstract}
-In dynamical systems with highly nonorthogonal linear eigenvectors, non-modal linear analysis can be much more useful than normal-mode analysis in predicting turbulent properties. 
-However, the non-trivial time evolution of non-modal structures makes quantitative prediction difficult. We present a technique using only linear calculations to overcome
-this difficulty by modelling the effect that the advective nonlinearities have on spatial turbulent structures. Specifically, we model the nonlinearities as a periodic randomizing force with
+In nonlinear dynamical systems with highly nonorthogonal linear eigenvectors, linear non-modal analysis is more useful than normal mode analysis in predicting turbulent properties. 
+However, the non-trivial time evolution of non-modal structures makes quantitative prediction difficult. Here, we present a technique to overcome this difficulty, which includes
+modelling the effect that the advective nonlinearities have on spatial turbulent structures. The nonlinearities are taken as a periodic randomizing force with
 period consistent with critical balance arguments. We apply this technique to a model of drift wave turbulence in the Large Plasma Device (LAPD) 
@@
 The reason for this failure was not clearly explained until the early 1990's when Trefethen and others attributed the pitfalls of normal mode analysis to the non-normality of linear operators of
-dynamical systems~\cite{trefethen1993,trefethen2005} -- a non-normal operator or matrix is one that does not commute with its adjoint and consequently has 
+dynamical systems~\cite{trefethen1993,trefethen2005}. A non-normal operator or matrix is one that does not commute with its adjoint, and it has 
 eigenvectors that are nonorthogonal to one another. One consequence of eigenvector nonorthogonality is that even when all eigenvectors are stable and decay exponentially under linear evolution, 
@@
 than on properties of fully-developed turbulence.
-This paper, in contrast, takes up the task of developing a quantitative approach to predicting turbulent properties using only non-modal linear calculations. 
+This paper takes up the task of developing a quantitative approach to predicting turbulent properties using only non-modal linear calculations. 
 The basic approach we take is to solve the linear initial value problem using an ensemble of random initial conditions, 
@@
 
-The equations are global and partially linearized; the only nonlinearities retained are the advective nonlinearities in the Poisson brackets. 
-We add artificial diffusion and viscosity terms with small numerical coefficients ($\mu_N, \mu_\phi, \mu_T = 10^{-3}$), 
-used for numerical stability in nonlinear simulations, which are performed with the BOUT++ code~\cite{dudson2009}. We use periodic axial and zero value radial 
+The equations are global and partially linearized. The only nonlinearities retained are the advective nonlinearities, which are written with Poisson brackets. 
+We add artificial diffusion and viscosity terms with small numerical coefficients ($\mu_N, \mu_\phi, \mu_T = 10^{-3}$)
+for numerical stability in nonlinear simulations, which are performed with the BOUT++ code~\cite{dudson2009}. We use periodic axial and zero value radial 
 boundary conditions. Further details of the model, including validation studies, may be found in the references~\cite{Popovich2010a,Popovich2010b,Umansky2011,friedman2012b,friedman2013},
@@
 ~\cite{friedman2012b,friedman2013}.
-This nonlinear instability process, which was discovered by Drake et al.~\cite{drake1995} using a similar plasma model to ours, works as follows: 
+This nonlinear instability process, which was discovered by Drake et al.~\cite{drake1995} in a similar plasma model to ours, works as follows: 
 long ($k_\para=0$) convective filaments transport density across the equilibrium density profile, setting up $k_\para=0$ density filaments (the same applies to the temperature). 
@@
 As already stated, one sees from Fig.~\ref{spec_nl_gamma} that the linear eigenmodes are unstable for $n=1$ and stable for $n=0$ as expected for drift waves. 
-However, the turbulent growth rates are positive for $n=0$ and low $m$ (which is subcritical-like), but negative for $n=1$ at all $m$. 
+However, the turbulent growth rates are positive for $n=0, m<50$ (subcritical-like behavior), but negative for $n=1$ at all $m$. 
 There was already a hint of this from the initial behavior of the curves in Fig.~\ref{transient_n0_n1_growth}, but this is a quantitative and more thorough picture.
@@
 While the nonlinearities obviously do not affect the system in discrete time steps, this simplification provides a way to use linear non-modal analysis in a predictive manner.
-In other words, there are certain spatial structures associated with the curves in Fig.~\ref{m20n0_time_evs} b) that change as a function of time. We consider only those structures from $t=0$, where
-the structures are completely random, to $t = \tau_{\rm{nl}}$, and associate those structures with turbulent structures. These provide us with all we need to characterize the turbulence.
+In other words, there are certain spatial structures associated with the curves in Fig.~\ref{m20n0_time_evs} b) that change as a function of time. We consider only those structures from $t=0$ (where
+the structures are completely random) to $t = \tau_{\rm{nl}}$, and associate those structures with turbulent structures. These provide us with all we need to characterize the turbulence.
 
@@
 Thus, we invoke the conjecture of \emph{critical balance}, which posits that the nonlinear time scale equals the linear time scale at all spatial scales~\cite{schekochihin2012}. 
-We can then estimate the nonlinear time scale as $\tau_{\rm{nl}} = \omega_{\rm{lin}}^{-1}$, the inverse of the linear frequency (at each $m,n$).
-There are several choices of linear frequencies that we can use. Since we are dealing with a system of drift waves, $\omega_* = \frac{T_e m}{L_n r m_i \omega_{ci}^2}$ seems suitable, however,
-$\omega_*$ is radially dependent, so it is not obvious which radius to choose. Alternatively, the linear frequency of the fastest growing eigenmode $\omega_s$ may be used. This is not a function
-of radius, but $\omega_s = 0$ for $n=0$ eigenmodes, so one is forced to use $\omega_s$ for the fastest growing $n=1$ eigenmode at each $m$ to get a meaningful time scale. This is not wholly
+We can then estimate the nonlinear time scale as $\tau_{\rm{nl}} = \omega_s^{-1}$ -- the inverse of the linear frequency (at each $m,n$).
+However, $\omega_s = 0$ for $n=0$ eigenmodes, so we are forced to use $\omega_s$ for the fastest growing $n=1$ eigenmode at each $m$ to get a meaningful time scale. This is not wholly
 unjustified, however, due to the inherent nonlinear coupling of different modes, which is evident in Eq.~\ref{mat_eq}.
@@
 Then, in Fig.~\ref{m20n0_time_evs} d), we calculate the instantaneous growth rate as a function of time of the ensemble averaged curve over that time period.
-The average of this instantaneous growth rate curve (the dashed line) gives the average ``Transient Growth Rate.''
-In Fig.~\ref{spec_nl_transient_gamma}, we compare the Transient Growth Rate Spectra to the Turbulent Growth Rate Spectra, where we use $\omega_s^{-1}$ for $\tau_{\rm{nl}}$.
+The average of this instantaneous growth rate curve (the dashed line) gives a ``Transient Growth Rate.''
+In Fig.~\ref{spec_nl_transient_gamma}, we compare the Transient Growth Rate Spectra (labeled ``Trans. $\omega_s^{-1}$'') to the Turbulent Growth Rate Spectra.
 The $n=0$ curves in Fig.~\ref{spec_nl_transient_gamma} a) are fairly similar, especially when compared to the eigenmode spectra.
-The $n=1$ curves, on the other hand, are not very similar, although they trend with $m$ in the same way, which is very different from the linear eigenmode trend.
-The difference between the $n=1$ transient and turbulent growth rate spectra is due to nonlinear effects not captured by the simple transient calculation procedure. In particular, the turbulence
-exhibits strong energy transfer from $n=0$ to $n=1$ Fourier components which is then dissipated by electron-ion collisions as the electrons stream along the field lines~\cite{friedman2012b}. 
-This extra $n=1$ energy dissipation in the turbulent simulation is not a transient effect, and the $n=0$ to $n=1$ transfer is not a simple nonlinear mixing effect.
+The $n=1$ curves in Fig.~\ref{spec_nl_transient_gamma} b), on the other hand, are not very similar, although they trend with $m$ in the same way, which is very different from the linear eigenmode trend.
+The difference between the $n=1$ transient and turbulent growth rate spectra is due to nonlinear effects not captured by the simple transient calculation procedure. In particular, the nonlinear
+simulation has a strong energy transfer from $n=0$ to $n=1$ Fourier components which is then dissipated by electron-ion collisions as the electrons stream along the field lines~\cite{friedman2012b}.
 
 There is one more growth rate curve in Fig.~\ref{spec_nl_transient_gamma}, labelled ``Trans $\omega_s^{-1}$ cascade.'' We include this to address the apparent downshift (in $m$) of
-the transient $\omega_s^{-1}$ curve in comparison to the turbulent curve. The downshift is a consequence of our oversimplification of the nonlinearity. Specifically, we
+the ``Trans $\omega_s^{-1}$'' curve in comparison to the turbulent curve. The downshift is another consequence of our oversimplification of the nonlinearity. Specifically, we
 used $\omega_s^{-1}$ at each $m$ for the nonlinear time scale at that same $m$. As seen in Eq.~\ref{mat_eq}, $v_{m,n}$ is affected by every $\mathbf{v}_{E,m-m',n-n'} k^{'}_\perp$, not by only
@@
 Using this formula with $\gamma_{s,\rm{max}} = 0.002$ and $m=60$, one would predict a turbulent saturation level on the order of $0.1 \%$. 
-On the other hand, using this formula with the transient values -- $\gamma_{t,\rm{max}} = 0.003, m=10$ -- we predict the saturation level on the order of $10 \%$,
+On the other hand, using this formula with the transient values $\gamma_{t,\rm{max}} = 0.003, \ m=10$, we predict the saturation level on the order of $10 \%$,
 which matches the saturation level in the nonlinear simulation.

### 2013-12-19 Transient_growth_prl.tex
Dec 19

@@
-\documentclass[showpacs,preprintnumbers,amsmath,amssymb,superscriptaddress,aip]{revtex4-1}
+%\documentclass[showpacs,preprintnumbers,amsmath,amssymb,superscriptaddress,aip]{revtex4-1}
+\documentclass[letter,scriptaddress,twocolumn, prl,showkeys]{revtex4}
 \usepackage{graphicx}
@@
 
-\title{A Linear Technique to Predict Non-Normal LAPD Turbulence}
+\title{A Linear Technique to Predict Non-Normal Turbulence Applied to a Magnetized Plasma}
 
@@
 In nonlinear dynamical systems with highly nonorthogonal linear eigenvectors, linear non-modal analysis is more useful than normal mode analysis in predicting turbulent properties. 
-However, the non-trivial time evolution of non-modal structures makes quantitative prediction difficult. Here, we present a technique to overcome this difficulty, which includes
-modelling the effect that the advective nonlinearities have on spatial turbulent structures. The nonlinearities are taken as a periodic randomizing force with
+However, the non-trivial time evolution of non-modal structures makes quantitative prediction difficult.
+We present a technique to overcome this difficulty by modelling the effect that the advective nonlinearities have on spatial turbulent structures.
+The nonlinearities are taken as a periodic randomizing force with
 period consistent with critical balance arguments. We apply this technique to a model of drift wave turbulence in the Large Plasma Device (LAPD) 
-[W. Gekelman \emph{et al.}, Rev. Sci. Inst. {\bf 62}, 2875 (1991)], where non-modal effects dominate the turbulent system.
+[W. Gekelman \emph{et al.}, Rev. Sci. Inst. {\bf 62}, 2875 (1991)], where non-modal effects dominate the turbulence.
 We compare the resulting growth rate spectrum to that obtained from a nonlinear simulation, showing good quantitative agreement.
@@
 
-\section{Introduction}
-
-Normal mode analysis -- the calculation of eigenvalues and eigenvectors of a linearized dynamical system -- has been used for the last two centuries to solve a number of problems ranging from
-heat conduction along a solid bar to quantum mechanical energy states of atoms. Despite its wide-ranging success, normal mode analysis has seemingly failed in particular instances,
-particularly in predicting the onset of turbulence in many hydrodynamic flows, where the turbulence is called subcritical. 
-The reason for this failure was not clearly explained until the early 1990's when Trefethen and others attributed the pitfalls of normal mode analysis to the non-normality of linear operators of
-dynamical systems~\cite{trefethen1993,trefethen2005}. A non-normal operator or matrix is one that does not commute with its adjoint, and it has 
+Normal mode analysis -- the calculation of eigenvalues and eigenvectors of a linearized dynamical system -- has long been used to solve a number of problems ranging from
+heat conduction along a solid bar to quantum mechanical energy states of atoms. Despite its wide-ranging success, normal mode analysis has
+failed in important instances, particularly in predicting the onset of subcritical turbulence in hydrodynamic flows. 
+The reason for this failure was not explained until the early 1990's when Trefethen and others attributed the pitfalls of normal mode analysis to the non-normality of linear operators of
+dynamical systems~\cite{trefethen1993,trefethen2005, schmid2007}. A non-normal operator has 
 eigenvectors that are nonorthogonal to one another. One consequence of eigenvector nonorthogonality is that even when all eigenvectors are stable and decay exponentially under linear evolution, 
 superpositions of eigenvectors can grow, albeit transiently.
-A paradigmatic illustration of this transient growth process for two nonorthogonal eigenvectors may be seen if Figure 2 of a review paper by Schmid~\cite{schmid2007}.
 Physically, this means that certain types of fluctuations of the laminar state can access free energy from background gradients even though normal mode fluctuations cannot.
@@
 Such behavior is obscured by traditional normal mode analysis, which only effectively describes the long time asymptotic behavior of fluctuations under the 
-action of the linear operator. Transient growth events, which can dominate nonlinear turbulent evolution, require initial-value (non-modal) calculations rather than normal mode calculations.
+action of the linear operator. Transient growth events, which can dominate nonlinear evolution, require initial-value (non-modal) calculations rather than normal mode calculations.
 
@@
 still heavily relies on normal mode analysis to inform turbulent predictions and observations, with a few notable exceptions~\cite{camargo1998,camporeale2010,schekochihin2012}. 
-Furthermore, non-modal treatments thus far have generally been explanatory rather than predictive and have centered around the transition to turbulence in subcritical systems rather 
+Furthermore, non-modal treatments have generally been explanatory rather than predictive and have centered around the transition to turbulence in subcritical systems rather 
 than on properties of fully-developed turbulence.
 This paper takes up the task of developing a quantitative approach to predicting turbulent properties using only non-modal linear calculations. 
-The basic approach we take is to solve the linear initial value problem using an ensemble of random initial conditions, 
-and then calculate the average growth rate of the solutions up to one nonlinear decorrelation time. We take this nonlinear decorrelation time to be a characteristic \emph{linear} time.
+Our approach is to evolve the linear solution from an ensemble of random initial conditions 
+and calculate the average growth rate of the solutions over a specified timescale.  Optimally this time scale would be the
+nonlinear decorrelation time of the turbulent system, but in order to enable predictive capability, we employ critical balance
+arguments to use a characteristic \emph{linear} time.
 This procedure produces an effective growth rate spectrum that can be used to predict turbulent properties such as saturation levels.
-While the concepts behind this technique are general enough to be applied to many nonlinear dynamical systems, the details will vary from system to system, 
-so we restrict our treatment to one particular turbulence model.
+While the concepts behind this technique are general enough to be applied to many nonlinear dynamical systems, the details vary for each case, 
+so we restrict our treatment to one particular turbulence model. The technique reproduces the turbulent growth rate spectrum of the direct nonlinear simulation fairly well, especially
+in comparison to the linear spectrum, which has little bearing on the turbulence.
 
-The model is designed to describe a drift wave turbulence experiment conducted in the Large Plasma Device (LAPD)~\cite{gekelman1991}.
-LAPD is a linear machine with a straight magnetic $\mathbf{B}$ field.
-Due to its large dimensions and high collisionality, LAPD is apt for fluid modelling. Thus, we use a reduced Braginskii~\cite{Braginskii1965} 2-fluid model:
+The model describes pressure-gradient-driven turbulence in the uniformly magnetized, cylindrical plasma
+produced by the Large Plasma Device (LAPD)~\cite{gekelman1991}. 
+Due to LAPD's large size and high collisionality, a fluid model is
+appropriate to describe the turbulence~\cite{Popovich2010a,Popovich2010b,Umansky2011}. We use a reduced Braginskii~\cite{Braginskii1965} 2-fluid model:
 
@@
 \label{ve_eq}
-\pdt \vpe = - \fmie \frac{T_{e0}}{N_0} \gradpar N - 1.71 \fmie \gradpar T_e  + \fmie \gradpar \phi - \nue \vpe + \{\phi,\vpe \}, \\
+\pdt \vpe = - \fmie \frac{T_{e0}}{N_0} \gradpar N - 1.71 \fmie \gradpar T_e  \nonumber \\
++ \fmie \gradpar \phi - \nue \vpe + \{\phi,\vpe \}, \\
 \label{rho_eq}
@@
 \eeqar
-where $N_0$ and $N$ are the equilibrium and fluctuating density, $\vpe$ is the fluctuating parallel electron velocity, $\varpi \equiv \gradperp \cdot (N_0 \gradperp \phi)$ is the potential vorticity
-of the fluctuating potential $\phi$, and $T_{e0}$ and $T_e$ are the equilibrium and fluctuating electron temperature. The equations employ Bohm normalizations: lengths are
+where $N$ is the density, $\vpe$ the parallel electron velocity, $\varpi \equiv \gradperp \cdot (N_0 \gradperp \phi)$ is potential vorticity,
+and $T_e$ the electron temperature. The equations employ Bohm normalizations: lengths are
 normalized to the ion sound gyroradius $\rho_s$, times to the ion cyclotron time $\omega_{ci}^{-1}$, velocities to the sound speed $c_s$, densities to the equilibrium peak density, and electron
 temperatures and potentials to the equilibrium peak electron temperature. The profiles $N_0$ and $T_{e0}$ and other parameters are all taken from experimental measurements. $\phi_0$ 
-does not appear in the equations because we model one particular experiment in which the equilibrium radial electric field was nulled by boundary biasing~\cite{schaffner2012}.
+does not appear in the equations because we model one particular
+experiment in which the equilibrium radial electric field and
+therefore azimuthal flow was reduced to near zero value using a
+limiter biasing technique~\cite{schaffner2012}.
 
@@
 We add artificial diffusion and viscosity terms with small numerical coefficients ($\mu_N, \mu_\phi, \mu_T = 10^{-3}$)
-for numerical stability in nonlinear simulations, which are performed with the BOUT++ code~\cite{dudson2009}. We use periodic axial and zero value radial 
+to ensure numerical stability in nonlinear simulations, which are performed with the BOUT++ code~\cite{dudson2009}. We use periodic axial and zero value radial 
 boundary conditions. Further details of the model, including validation studies, may be found in the references~\cite{Popovich2010a,Popovich2010b,Umansky2011,friedman2012b,friedman2013},
-though we mention here that the model reproduces all statistical properties of the experimental turbulence within a factor of two, and in most cases, much better than that.
+though we mention here that the model does very well in reproducing
+the statistical properties of the experimentally-observed turbulence.
 
-The nonlinear simulation reveals a fascinating property of the turbulence -- it is dominated by a nonlinear instability process despite being linearly unstable to drift waves
-~\cite{friedman2012b,friedman2013}.
-This nonlinear instability process, which was discovered by Drake et al.~\cite{drake1995} in a similar plasma model to ours, works as follows: 
-long ($k_\para=0$) convective filaments transport density across the equilibrium density profile, setting up $k_\para=0$ density filaments (the same applies to the temperature). 
-These filaments are unstable to drift waves causing a secondary drift wave instability to grow on the filaments. 
-These drift waves, which have finite $k_\para$ and propagate primarily radially, nonlinearly couple to one another and reinforce the convective filaments.
+The nonlinear simulation reveals a fascinating property of the turbulence -- it is dominated by a nonlinear instability process despite being linearly unstable to drift waves~\cite{friedman2012b,friedman2013}.
+This nonlinear instability process, which was discovered by Drake et al.~\cite{drake1995} in a similar plasma model, works as follows: 
+magnetic-field-aligned ($k_\para=0$) convective filaments transport density across the equilibrium density gradient, setting up $k_\para=0$ density filaments (the same applies to the temperature). 
+These filaments are unstable to a secondary drift wave instability,
+which grows on the periphery of the filaments. 
+These drift waves, which have finite $k_\para$ and propagate primarily
+radially, nonlinearly couple to one another and reinforce the original
+convective filaments.
 
@@
 \label{dEdt_def}
-\diff{E(m,n,t)}{t} = Q(m,n,t) + D(m,n,t) + \sum_{m',n'} T(m,m',n,n',t).
+\diff{E(m,n)}{t} = Q(m,n) + D(m,n) + \sum_{m',n'} T(m,m',n,n').
 \eeq
@@
 Futher, $Q$ represents the linear energy exchange with the equilibrium (convective transport), $D$ the dissipation from irreversible processes
-like collisions, and $T$ the nonlinear triad wave transfer. Since $\int f \{g,f\} dV = 0$, or equivalently $\sum_{m,m',n,n'} T(m,m',n,n',t)=0$, 
+like collisions, and $T$ the nonlinear triad wave transfer. Since $\int f \{g,f\} dV = 0$, or equivalently $\sum_{m,m',n,n'} T(m,m',n,n')=0$, 
 the Poisson bracketed advective nonlinearities in the model are energetically conservative and therefore do no inject or dissipate energy on the whole.
@@
 \begin{figure}
-\centerline{\includegraphics[]{transient_n0_n1_growth}}
-\caption{The linear evolution of the energy of several Fourier modes starting from a turbulent initial state. The $n=0$ curves have an initial period of transient growth before exponentially decaying,
-while the $n=1,m=20$ curve transiently decays before growing exponentially.}
+\centerline{\includegraphics[width=0.4\textwidth]{transient_n0_n1_growth}}
+\caption{Linear evolution of energy starting from a turbulent initial state. The $n=0$ curves have an initial period of transient growth before exponentially decaying.}
 \label{transient_n0_n1_growth}
@@
 Both are transient growth processes. We see this in our simulations by following the evolution of the energy of several Fourier $m,n$ modes after turning off the nonlinearities in an already
-turbulent simulation. A few representative $m,n$ modes are shown in Fig.~\ref{transient_n0_n1_growth}. The transient growth of the filamentary structures is seen in the $n=0$ modes, 
+turbulent simulation. A few representative $m,n$ modes are shown in
+Fig.~\ref{transient_n0_n1_growth}. The linear transient growth of the filamentary structures is seen in the $n=0$ modes, 
 which grow transiently before decaying exponentially at the rate of their least stable eigenmode. This kind of behavior is rather unituitive to some because all $n=0$ linear eigenmodes are stable, 
@@
 \begin{figure}
-\centerline{\includegraphics[]{spec_nl_gamma}}
-\caption{The linear and turbulent growth rate $m$ spectra for $n=0$ (solid lines) and $n=1$ (dashed lines) Fourier components. The linear growth rates are those for the least stable eigenmodes at each
-$n$ and $m$, while the turbulent growth rates represent $\pdiff{E}{t}|_{\rm{lin}}/2 E$ from the nonlinear simulation.
-The shaded region marks the $1 \sigma$ spread in the turbulent growth rate spectrum,
+\centerline{\includegraphics[width=0.45\textwidth]{spec_nl_gamma}}
+\caption{Linear and turbulent growth rate $m$ spectra for $n=0$ (solid lines) and $n=1$ (dashed lines) Fourier components. The linear growth rates are those of the least stable eigenmodes at each
+$m,n$, while the turbulent growth rates represent $\pdiff{E}{t}|_{\rm{lin}}/2 E$ from the nonlinear simulation. The shaded region marks the $1 \sigma$ spread in the turbulent growth rate spectrum,
 obtained from the distribution of growth rates calculated from the nonlinear simulation over a long time range.}
@@
 
-To obtain the turbulent growth rate spectrum from the nonlinear simulation, we begin with the expression $\diff{E(m,n,t)}{t}\big|_{\rm{lin}} = Q(m,n,t) + D(m,n,t)$, 
+To obtain the turbulent growth rate spectrum from the nonlinear simulation, we begin with the expression $\diff{E(m,n)}{t}\big|_{\rm{lin}} = Q(m,n) + D(m,n)$, 
 which accounts for the total net energy injection at each $m,n$ into the fluctuations. 
@@
 However, the turbulent growth rates are positive for $n=0, m<50$ (subcritical-like behavior), but negative for $n=1$ at all $m$. 
-There was already a hint of this from the initial behavior of the curves in Fig.~\ref{transient_n0_n1_growth}, but this is a quantitative and more thorough picture.
+There was already a hint of this from the initial behavior of the
+curves in Fig.~\ref{transient_n0_n1_growth}, but this figure represents
+quantitative confirmation.
 Again, the $n=0$ energy injection is a manifestation of the convective filaments drawing energy from the equilibrium density gradient and depositing this energy into the density fluctuations.
@@
 
-\beq
+\beqar
 \label{mat_eq}
-\mathbf{B}_{m,n} \diff{\mathbf{v}_{m,n}(t)}{t} = \mathbf{C}_{m,n} \mathbf{v}_{m,n}(t) - \sum_{m',n'}  \mathbf{v}_{E,m-m',n-n'} \cdot \gradperp \left( \mathbf{B}_{m',n'} \mathbf{v}_{m',n'}(t) \right),
-\eeq
-where $\mathbf{v}_{m,n} = \left( N(r_0), N(r_1), \ldots, \vpe(r_0), \vpe(r_1), \ldots, \phi(r_0), \phi(r_1), \ldots, T_e(r_0), T_e(r_1), \ldots \right)_{m,n}^{T}$ is the state of the system,
+\mathbf{B}_{m,n} \diff{\mathbf{v}_{m,n}(t)}{t} = \mathbf{C}_{m,n} \mathbf{v}_{m,n}(t) \nonumber \\
+- \sum_{m',n'}  \mathbf{v}_{E,m-m',n-n'} \cdot \gradperp \left( \mathbf{B}_{m',n'} \mathbf{v}_{m',n'}(t) \right),
+\eeqar
+where $\mathbf{v}_{m,n} = \left( N(r), \vpe(r), \phi(r), T_e(r) \right)_{m,n}^{T}$ is the state of the system,
 and $\mathbf{B}_{m,n}$ and $\mathbf{C}_{m,n}$ are coefficient matrices that include the equilibrium information and finite difference coefficients.
@@
 We illustrate $G_{\rm{max}}(t)$ along with $G(t)$ for several random initial conditions in Fig.~\ref{m20n0_time_evs} a), all at $n=0$ and $m=20$. Despite the fact that all eigenmodes at this
-wavenumber are stable, there is the possibility for transient amplification by a factor of $20$. Further, a sample of randomly initialized vectors are all transiently amplified, but none display
+wavenumber are stable, there is the possibility for transient amplification by a factor of $20$. Furthermore, a sample of randomly initialized vectors are all transiently amplified, but none display
 optimal growth as compared to $G_{\rm{max}}(t)$. We show a zoomed-in view of these randomly initialized curves in Fig.~\ref{m20n0_time_evs} b).
@@
 \begin{figure}
-\centerline{\includegraphics[]{m20n0_time_evs}}
+\centerline{\includegraphics[width=0.5\textwidth]{m20n0_time_evs}}
 \caption{a) The maximum growth ratio curve $G_{\rm{max}}(t)$ (dashed line) for $n=0,m=20$, and an ensemble of growth ratio curves (solid gray lines, with the solid black line the ensemble average)
-that start with a random initial condition $u(0)$ and evolve under the linear operator. b) A zoomed in version of the ensemble of growth ratio curves. c) A further zoomed in version of the ensemble of
-curves with a time range of $\omega_s^{-1}$, which is the characteristic linear time scale for drift waves. 
-d) The instantaneous growth rate as a function of time of the the ensemble averaged growth ratio along with the average value over the time spanning $\omega_s^{-1}$.}
+that start with a random initial condition $u(0)$ and evolve under the linear operator. b) The ensemble of growth ratio curves alone, and c) cut off at $t=\omega_s^{-1}$, the drift wave eigenmode
+time scale. d) The instantaneous growth rate of the ensemble averaged growth ratio along with the average value.}
 \label{m20n0_time_evs}
@@
 
-Now, since we do not know, \emph{a priori} at what level the turbulence will saturate, we cannot calculate the relative magnitude of $\tau_{\rm{nl}}$ from $(v_E k_\perp)^{-1}$ directly.
+Now, since we do not know \emph{a priori} at what level the turbulence will saturate, we cannot calculate the relative magnitude of $\tau_{\rm{nl}}$ from $(v_E k_\perp)^{-1}$ directly.
 Thus, we invoke the conjecture of \emph{critical balance}, which posits that the nonlinear time scale equals the linear time scale at all spatial scales~\cite{schekochihin2012}. 
@@
 To illustrate our procedure further, we show in Fig.~\ref{m20n0_time_evs} c) the randomly initialized linear evolution curves from start up until a time of $\omega_s^{-1}$. 
+The transient growth is cut off well before optimal amplification occurs, but as reflected by our procedure, there is no guarantee that the nonlinearities will allow complete non-modal amplification.
 Then, in Fig.~\ref{m20n0_time_evs} d), we calculate the instantaneous growth rate as a function of time of the ensemble averaged curve over that time period.
@@
 \begin{figure}
-\centerline{\includegraphics[]{spec_nl_transient_gamma}}
+\centerline{\includegraphics[width=0.5\textwidth]{spec_nl_transient_gamma}}
 \caption{The linear, turbulent, and transient growth rate spectra.
@@
 LAPD experiment, a nonlinear instability with linear $k_\para=0$ drive dominates the energy dynamics of the turbulence. While this is a surprising result in light of normal mode analysis,
-it is not at all surprising given the system's non-modal behavior. In general, non-modal initial-value analysis is difficult to quantify and make predictive, but using some simple nonlinear modelling,
-we have shown that it is possible.
+it is not at all surprising given the system's non-modal behavior, which we show with a technique that takes into account the effect of the nonlinearities. 
+In general, non-modal initial-value analysis is difficult to quantify and make predictive, but using some simple nonlinear modelling, we have shown that it is possible. Future studies will
+attempt to improve on the nonlinear modelling and expand the use of this non-modal technique to other non-normal turbulence models.
 
@@
 
+%\bibliographystyle{unsrt}
 \bibliography{refs}
-%\bibliographystyle{unsrt}
 

### 2013-12-20 Transient_growth_prl.tex
Dec 20

@@
 The reason for this failure was not explained until the early 1990's when Trefethen and others attributed the pitfalls of normal mode analysis to the non-normality of linear operators of
-dynamical systems~\cite{trefethen1993,trefethen2005, schmid2007}. A non-normal operator has 
-eigenvectors that are nonorthogonal to one another. One consequence of eigenvector nonorthogonality is that even when all eigenvectors are stable and decay exponentially under linear evolution, 
+dynamical systems~\cite{trefethen1993,schmid2007}. A non-normal operator has 
+eigenvectors that are nonorthogonal to one another. One consequence of eigenvector nonorthogonality is that even when all eigenvectors decay exponentially under linear evolution, 
 superpositions of eigenvectors can grow, albeit transiently.
-Physically, this means that certain types of fluctuations of the laminar state can access free energy from background gradients even though normal mode fluctuations cannot.
+In other words, certain fluctuations of the laminar state can access free energy from background gradients even though normal mode fluctuations cannot.
 When combined with nonlinear effects, this allows for sustained subcritical turbulence.
@@
 Due to LAPD's large size and high collisionality, a fluid model is
-appropriate to describe the turbulence~\cite{Popovich2010a,Popovich2010b,Umansky2011}. We use a reduced Braginskii~\cite{Braginskii1965} 2-fluid model:
+appropriate to describe the turbulence~\cite{Popovich2010a,Popovich2010b,Umansky2011}. We use a reduced Braginskii 2-fluid model:
 
@@
 \label{ni_eq}
-\pdt N = - {\mathbf v_E} \cdot \grad N_0 - N_0 \gradpar \vpe + \mu_N \gradperp^2 N + S_N + \{\phi,N\}, \\
+\pdt N = - {\mathbf v_E} \cdot \grad N_0 - N_0 \gradpar \vpe + S_N + \{\phi,N\}, \\
 \label{ve_eq}
@@
 \label{rho_eq}
-\pdt \varpi = - N_0 \gradpar \vpe  - \nuin \varpi + \mu_\phi \gradperp^2 \varpi + \{\phi,\varpi \}, \\
+\pdt \varpi = - N_0 \gradpar \vpe  - \nuin \varpi + \{\phi,\varpi \}, \\
 \label{te_eq}
-\pdt T_e = - {\mathbf v_E} \cdot \grad T_{e0} - 1.71 \frac{2}{3} T_{e0} \gradpar \vpe + \frac{2}{3 N_0} \kpe \gradpar^2 T_e  \nonumber \\
-- \frac{2 m_e}{m_i} \nue T_e  + \mu_T \gradperp^2 T_e +  S_T + \{\phi,T_e\},
+\pdt T_e = - {\mathbf v_E} \cdot \grad T_{e0} - 1.71 \frac{2}{3} T_{e0} \gradpar \vpe \nonumber \\
++ \frac{2}{3 N_0} \kpe \gradpar^2 T_e  - \frac{2 m_e}{m_i} \nue T_e  +  S_T + \{\phi,T_e\},
 \eeqar
-where $N$ is the density, $\vpe$ the parallel electron velocity, $\varpi \equiv \gradperp \cdot (N_0 \gradperp \phi)$ is potential vorticity,
-and $T_e$ the electron temperature. The equations employ Bohm normalizations: lengths are
+where $N$ is the density, $\vpe$ the parallel electron velocity, $\varpi \equiv \gradperp \cdot (N_0 \gradperp \phi)$ the potential vorticity,
+and $T_e$ the electron temperature. All lengths are
 normalized to the ion sound gyroradius $\rho_s$, times to the ion cyclotron time $\omega_{ci}^{-1}$, velocities to the sound speed $c_s$, densities to the equilibrium peak density, and electron
-temperatures and potentials to the equilibrium peak electron temperature. The profiles $N_0$ and $T_{e0}$ and other parameters are all taken from experimental measurements. $\phi_0$ 
-does not appear in the equations because we model one particular
-experiment in which the equilibrium radial electric field and
-therefore azimuthal flow was reduced to near zero value using a
+temperatures and potentials to the equilibrium peak electron temperature. The profiles $N_0$ and $T_{e0}$ and other parameters are taken from experimental measurements. $\phi_0$ 
+is absent in the equations because we model one particular experiment in which the equilibrium radial electric field was reduced to near zero value using a
 limiter biasing technique~\cite{schaffner2012}.
@@
 The equations are global and partially linearized. The only nonlinearities retained are the advective nonlinearities, which are written with Poisson brackets. 
-We add artificial diffusion and viscosity terms with small numerical coefficients ($\mu_N, \mu_\phi, \mu_T = 10^{-3}$)
+We add artificial diffusion and viscosity terms with small numerical coefficients ($10^{-3}$)
 to ensure numerical stability in nonlinear simulations, which are performed with the BOUT++ code~\cite{dudson2009}. We use periodic axial and zero value radial 
@@
 In fact, the other parts of the mechanism are driven by energetically conservative nonlinear interactions, 
-meaning that the convective transport is the only step responsible for energy injection into the fluctuations from the laminar state.
+meaning that the convective transport is the only step responsible for energy injection into the fluctuations.
 Deriving an equation for the evolution of the energy from Eqs.~\ref{ni_eq}-\ref{te_eq}~\cite{friedman2012b,friedman2013}, we may symbolically write
@@
 \label{dEdt_def}
-\diff{E(m,n)}{t} = Q(m,n) + D(m,n) + \sum_{m',n'} T(m,m',n,n').
+\diff{E(m,n)}{t} = Q(m,n) + D(m,n) + \sum_{m',n'} T(m,m',n,n')
 \eeq
-In this equation, $m$ and $n$ represent the azimuthal and axial Fourier mode numbers, respectively. 
+where $m$ and $n$ represent the azimuthal and axial Fourier mode numbers. 
 Since Eqs.~\ref{ni_eq}-\ref{te_eq} are not explicitly dependent on $\theta$ and $z$, the energy equation is separable for every $m$ and $n$. 
-Futher, $Q$ represents the linear energy exchange with the equilibrium (convective transport), $D$ the dissipation from irreversible processes
+$Q$ represents the linear energy exchange with the equilibrium (convective transport), $D$ the dissipation from irreversible processes
 like collisions, and $T$ the nonlinear triad wave transfer. Since $\int f \{g,f\} dV = 0$, or equivalently $\sum_{m,m',n,n'} T(m,m',n,n')=0$, 
@@
 This convective transport of density filaments is akin to the paradigmatic ``lift-up'' mechanism in hydrodynamic shear flows whereby streamwise vortices drive streamwise streaks~\cite{krommes1999}.
-Both are transient growth processes. We see this in our simulations by following the evolution of the energy of several Fourier $m,n$ modes after turning off the nonlinearities in an already
-turbulent simulation. A few representative $m,n$ modes are shown in
-Fig.~\ref{transient_n0_n1_growth}. The linear transient growth of the filamentary structures is seen in the $n=0$ modes, 
-which grow transiently before decaying exponentially at the rate of their least stable eigenmode. This kind of behavior is rather unituitive to some because all $n=0$ linear eigenmodes are stable, 
+Both are transient growth processes. We see this in our simulations by following the evolution of the energy of several $m,n$ modes after turning off the nonlinearities in an already
+turbulent simulation. A few representative modes are shown in Fig.~\ref{transient_n0_n1_growth}. The linear transient growth of the filamentary $n=0$ structures is evident as their modes 
+grow transiently before decaying exponentially at the rate of their least stable eigenmode. Such behavior is rather unituitive to some because all $n=0$ linear eigenmodes are stable, 
 and the normal mode paradigm has conditioned us to believe that a linear superposition of such eigenmodes should decay. But as the hydrodynamics community discovered, linear
 superpositions of nonorthogonal vectors may behave unintuitively.
-Notice now that the $n=1,m=20$ mode decays transiently before growing exponentially with growth rate of the most unstable eigenmode at $n=1,m=20$. This isn't quite as interesting as the $n=0$
-behavior, and it can occur in normal systems. Futhermore, it has no parallel in truly subcritical turbulence, but it can be helpful for interpreting later results.
-
-We stress that Fig.~\ref{transient_n0_n1_growth} is obtained by \emph{linear} evolution from an initial turbulent state.
-Since the transient growth, which is behind subcritical turbulence, is a purely linear phenomenon, it has long been a goal of researchers to predict the onset of subcritical turbulence using
-only linear, non-modal calculations. It is our goal, however, to go a step further and use linear calculations to understand and even predict turbulent properties.
-Such calculations would be much faster and easier to interpret than direct simulations of nonlinear equations. Perhaps one of the most useful statistical properties that we can use to describe
-the turbulence -- and one that is conveniently ammenable to linear analysis -- is the turbulent growth rate spectrum.
+Notice now that the $n=1,m=20$ mode decays transiently before growing exponentially with growth rate of the most unstable eigenmode. This isn't that interesting since this behavior
+can occur in normal systems, but it nevertheless explains later results.
 
@@
 \centerline{\includegraphics[width=0.45\textwidth]{spec_nl_gamma}}
-\caption{Linear and turbulent growth rate $m$ spectra for $n=0$ (solid lines) and $n=1$ (dashed lines) Fourier components. The linear growth rates are those of the least stable eigenmodes at each
-$m,n$, while the turbulent growth rates represent $\pdiff{E}{t}|_{\rm{lin}}/2 E$ from the nonlinear simulation. The shaded region marks the $1 \sigma$ spread in the turbulent growth rate spectrum,
-obtained from the distribution of growth rates calculated from the nonlinear simulation over a long time range.}
+\caption{Linear and turbulent growth rate spectra for $n=0$ (solid lines) and $n=1$ (dashed lines) Fourier components. The linear growth rates are those of the least stable eigenmodes,
+while the turbulent growth rates represent $\pdiff{E}{t}|_{\rm{lin}}/2 E$ from the nonlinear simulation. The shaded region marks the $1 \sigma$ spread in the turbulent spectrum,
+obtained from the distribution of growth rates in the nonlinear simulation.}
 \label{spec_nl_gamma}
@@
 
-To obtain the turbulent growth rate spectrum from the nonlinear simulation, we begin with the expression $\diff{E(m,n)}{t}\big|_{\rm{lin}} = Q(m,n) + D(m,n)$, 
+Since the transient growth, which is behind subcritical turbulence, is a purely linear phenomenon, it has long been a goal of researchers to predict the onset of subcritical turbulence using
+only linear, non-modal calculations. It is our goal, however, to use linear calculations to understand and even predict turbulent properties.
+Perhaps one of the most useful statistical properties that we can predict is the turbulent growth rate spectrum.
+To obtain a growth rate from the nonlinear simulation, we begin with the expression $\diff{E(m,n)}{t}\big|_{\rm{lin}} = Q(m,n) + D(m,n)$, 
 which accounts for the total net energy injection at each $m,n$ into the fluctuations. 
-An effective instantaneous growth rate $\gamma_{\rm{eff}}(m,n,t) = \diff{E(m,n,t)}{t}\big|_{\rm{lin}}/ 2 E(m,n,t)$ can then be defined. 
+An effective instantaneous growth rate $\gamma_t(m,n) = \diff{E(m,n)}{t}\big|_{\rm{lin}}/ 2 E(m,n)$ can then be defined. 
 This effective growth rate is calculated from the spatial structure of the fluctuations, so it is well-defined for both turbulent fluctuations and eigenmode fluctuations. If calculated
-from a linear simulation run for a sufficiently long time, $\gamma_{\rm{eff}}(m,n)$ is the same as the eigenmode growth rate of the least stable linear eigenmode for each $m,n$.
-For turbulence, $\gamma_{\rm{eff}}(m,n)$ is generally not equal to an eigenmode growth rate, especially in the case where the system is non-normal.
+from a linear simulation run for a sufficiently long time, $\gamma_t(m,n)$ is the same as the eigenmode growth rate of the least stable linear eigenmode for each $m,n$.
+For turbulence, $\gamma_t(m,n)$ is generally not equal to an eigenmode growth rate, especially in the case where the system is non-normal.
 We plot this growth rate in Fig.~\ref{spec_nl_gamma} for $n=0$ and $n=1$ components. The turbulent growth rate curves are calculated from this formula, 
-averaged over a time of about $3 \times 10^{4} \ \omega_{ci}^{-1}$ during the saturated turbulent phase of the nonlinear simulation: $\gamma_{\rm{Turb}}(m,n) = \int \gamma_{\rm{eff}}(m,n,t) dt$. 
-We show this turbulent growth rate along with the $1 \sigma$ contour, which indicates how greatly $\gamma_{\rm{eff}}(m,n,t)$ varies over time.
-Along with these, the growth rates of the least stable (or most unstable) linear eigenmodes are shown.
-
-As already stated, one sees from Fig.~\ref{spec_nl_gamma} that the linear eigenmodes are unstable for $n=1$ and stable for $n=0$ as expected for drift waves. 
-However, the turbulent growth rates are positive for $n=0, m<50$ (subcritical-like behavior), but negative for $n=1$ at all $m$. 
-There was already a hint of this from the initial behavior of the
-curves in Fig.~\ref{transient_n0_n1_growth}, but this figure represents
-quantitative confirmation.
-Again, the $n=0$ energy injection is a manifestation of the convective filaments drawing energy from the equilibrium density gradient and depositing this energy into the density fluctuations.
+averaged over a time of about $3 \times 10^{4} \ \omega_{ci}^{-1}$ during the saturated turbulent phase of the nonlinear simulation.
+Also, we show the $1 \sigma$ contour, which indicates how greatly $\gamma_t$ varies over time.
+Fig.~\ref{spec_nl_gamma} indicates that while the linear eigenmodes are unstable for $n=1$ and stable for $n=0$,
+the turbulent growth rates are positive for $n=0, m<50$ (subcritical-like behavior), but negative elsewhere. 
+Fig.~\ref{transient_n0_n1_growth} hinted at these characteristics, but this figure represents quantitative confirmation.
 
@@
 structures, which for non-normal systems, may be very different from eigenmode structures, and which change in time due to transient evolution.
-It is clear, however, that the turbulent structures at any given time are a result of both linear evolution and nonlinear energy transfer. 
+It is clear, though, that the turbulent structures are a result of both linear evolution and nonlinear energy transfer. 
 If we imagine that the nonlinearities randomize the turbulent structures at each wavenumber while the linearities evolve the structures
-deterministically, we can develop a procedure for choosing proper spatial structures that may be used in calculating effective growth rates.
-
+deterministically, we can develop a procedure for choosing proper spatial structures to use in calculating effective growth rates.
 To elucidate this procedure, we begin by taking Eqs.~\ref{ni_eq}-\ref{te_eq} and Fourier decomposing in the azimuthal and axial directions.
-Then, we discritize in the radial direction ($r \rightarrow r_0, r_1, \ldots, r_N $), and approximate radial derivatives with finite differences. 
+Then, we discritize in the radial direction and approximate radial derivatives with finite differences. 
 The resulting system of equations may be written in matrix form:
@@
 In order to use non-modal analysis to calculate growth rates and other measures, one must choose a norm and inner product with which to work. While any choice of
-inner product is possible, including that which orthogonalizes all eigenvectors, a physically relevant one is generally preferred~\cite{camargo1998,schmid2007,camporeale2010}. The most commonly used
-is an energy inner product. Recall that the inner product of two vectors may be written $\left< \mathbf{x},\mathbf{y} \right> = \mathbf{y}^{*} \mathbf{M} \mathbf{x}$,
-where $^*$ stands for the conjugate transpose, and $\mathbf{M}$ is a positive-definite matrix. $\mathbf{M}$ should be chosen such that 
-$||\mathbf{v}|| = \sqrt{\left< \mathbf{v},\mathbf{v} \right>} = \sqrt{E}$. That is, the norm of the state vector is the square root of the energy. Furthermore, it is often convenient in
+inner product is possible, including that which orthogonalizes all eigenvectors, a physically relevant one such as an energy inner product 
+is generally preferred~\cite{camargo1998,schmid2007,camporeale2010}. 
+Recall that the inner product of two vectors may be written $\left< \mathbf{x},\mathbf{y} \right> = \mathbf{y}^{*} \mathbf{M} \mathbf{x}$,
+where $^*$ stands for the conjugate transpose, and $\mathbf{M}$ is a positive-definite matrix. $\mathbf{M}$ is chosen so that 
+$||\mathbf{v}|| = \sqrt{\left< \mathbf{v},\mathbf{v} \right>} = \sqrt{E}$, where $E$ is the energy. Furthermore, it is convenient in
 computations to use the $L_2$-norm, $||\mathbf{u}||_2 = \sqrt{\sum_i |u_i|^2}$.
@@
 \eeq
-The eigenvalues of $\mathbf{A}$ are the same as those of $\mathbf{B}^{-1} \mathbf{C}$ because $\mathbf{M}^{\frac{1}{2}} \mathbf{B}^{-1} \mathbf{C} \mathbf{M}^{-\frac{1}{2}}$ is a similarity transformation.
-
-The solution of Eq.~\ref{lin_eq_A} is
-
-\beq
-\label{lin_soln}
-\mathbf{u}(t) = e^{\mathbf{A} t} \mathbf{u}(0).
-\eeq
-This solution depends on the initial condition $\mathbf{u}(0)$ in addition to the spectral properties of $\mathbf{A}$. For purposes of turbulent growth rate prediction, we are interested in
+%The eigenvalues of $\mathbf{A}$ are the same as those of $\mathbf{B}^{-1} \mathbf{C}$ because $\mathbf{M}^{\frac{1}{2}} \mathbf{B}^{-1} \mathbf{C} \mathbf{M}^{-\frac{1}{2}}$ is a similarity transformation.
+The solution of Eq.~\ref{lin_eq_A} is $\mathbf{u}(t) = e^{\mathbf{A} t} \mathbf{u}(0)$, which
+depends on the initial condition $\mathbf{u}(0)$ in addition to the spectral properties of $\mathbf{A}$. For purposes of turbulent growth rate prediction, we are interested in
 the behavior of $||\mathbf{u}(t)||$. Therefore, we introduce the growth ratio $G(t)$, which measures the amplification or reduction in the square root of the energy from an initial state:
@@
 \eeq
-It can be shown that this growth ratio is bounded from above by $G_{\rm{max}}(t) = ||e^{\mathbf{A} t}||$. Furthermore, $G_{\rm{max}}(t) = e^{\gamma_s t}$, where $\gamma_s$ is the spectral growth rate
-(the real part of the eigenvalue with largest real part), if and only if $\mathbf{A}$ is normal~\cite{schmid2007}. Otherwise, $G_{\rm{max}}(t) > e^{\gamma_s t}$.
+This ratio is bounded from above by $G_{\rm{max}}(t) = ||e^{\mathbf{A} t}||$. $G_{\rm{max}}(t) = e^{\gamma_s t}$, where $\gamma_s$ is the spectral growth rate
+if and only if $\mathbf{A}$ is normal~\cite{schmid2007}. Otherwise, $G_{\rm{max}}(t) > e^{\gamma_s t}$.
 
@@
 
-We illustrate $G_{\rm{max}}(t)$ along with $G(t)$ for several random initial conditions in Fig.~\ref{m20n0_time_evs} a), all at $n=0$ and $m=20$. Despite the fact that all eigenmodes at this
+We illustrate $G_{\rm{max}}(t)$ along with $G(t)$ at $m=20, n=0$ for several random initial conditions in Fig.~\ref{m20n0_time_evs} a).. Despite the fact that all eigenmodes at this
 wavenumber are stable, there is the possibility for transient amplification by a factor of $20$. Furthermore, a sample of randomly initialized vectors are all transiently amplified, but none display
-optimal growth as compared to $G_{\rm{max}}(t)$. We show a zoomed-in view of these randomly initialized curves in Fig.~\ref{m20n0_time_evs} b).
-It appears extremely unlikely to choose a random initial vector that either optimally amplifies the energy at any time or one that monotonically decays, which would be the case
+optimal growth as compared to $G_{\rm{max}}(t)$.
+It is extremely unlikely to choose a random initial vector that either optimally amplifies the energy at any time or one that monotonically decays, which would be the case
 if the initial vector were one of the linear eigenvectors.
@@
 \caption{a) The maximum growth ratio curve $G_{\rm{max}}(t)$ (dashed line) for $n=0,m=20$, and an ensemble of growth ratio curves (solid gray lines, with the solid black line the ensemble average)
-that start with a random initial condition $u(0)$ and evolve under the linear operator. b) The ensemble of growth ratio curves alone, and c) cut off at $t=\omega_s^{-1}$, the drift wave eigenmode
-time scale. d) The instantaneous growth rate of the ensemble averaged growth ratio along with the average value.}
+that start with a random initial condition $u(0)$ and evolve under the linear operator. b) The ensemble curves with $t=\omega_s^{-1}$ indicated, at which we cut off the evolution.}
 \label{m20n0_time_evs}
@@
 The key to quantifying non-modal analysis and making it predictive is to successfully model the effect that the nonlinearities have on the transient linear processes. 
-To this effect, we first note that the advective nonlinearity in Eq.~\ref{mat_eq} has the form of the state vector divided by a time $\tau_{\rm{nl}} \sim (v_E k_\perp)^{-1}$. This nonlinear
-time scale is generally associated with the eddy turnover time or the time in which eddies become decorrelated. In light of this, we present a heuristical model of the nonlinearities 
+To this effect, we note that the advective nonlinearity in Eq.~\ref{mat_eq} has the form of the state vector divided by a time $\tau_{\rm{nl}} \sim (v_E k_\perp)^{-1}$. This nonlinear
+time scale is generally associated with the eddy turnover time or the time in which eddies become decorrelated. We therefore present a heuristical model of the nonlinearities 
 as a randomizing force that acts on this characteristic nonlinear time scale.
-Specifically, we start with a random initial condition, let it evolve linearly for a time $\tau_{\rm{nl}}$, and then re-randomized the system.
+Specifically, we start with a random initial condition, evolve it linearly for a time $\tau_{\rm{nl}}$, and then re-randomized the system.
 While the nonlinearities obviously do not affect the system in discrete time steps, this simplification provides a way to use linear non-modal analysis in a predictive manner.
 In other words, there are certain spatial structures associated with the curves in Fig.~\ref{m20n0_time_evs} b) that change as a function of time. We consider only those structures from $t=0$ (where
-the structures are completely random) to $t = \tau_{\rm{nl}}$, and associate those structures with turbulent structures. These provide us with all we need to characterize the turbulence.
+the structures are completely random) to $t = \tau_{\rm{nl}}$, and associate those structures with turbulent structures.
 
@@
 unjustified, however, due to the inherent nonlinear coupling of different modes, which is evident in Eq.~\ref{mat_eq}.
+From Fig.~\ref{m20n0_time_evs} b), it is clear that the transient growth is cut off well before optimal amplification occurs. 
+But as reflected by our procedure, there is no guarantee that the nonlinearities will allow complete non-modal amplification.
 
-To illustrate our procedure further, we show in Fig.~\ref{m20n0_time_evs} c) the randomly initialized linear evolution curves from start up until a time of $\omega_s^{-1}$. 
-The transient growth is cut off well before optimal amplification occurs, but as reflected by our procedure, there is no guarantee that the nonlinearities will allow complete non-modal amplification.
-Then, in Fig.~\ref{m20n0_time_evs} d), we calculate the instantaneous growth rate as a function of time of the ensemble averaged curve over that time period.
-The average of this instantaneous growth rate curve (the dashed line) gives a ``Transient Growth Rate.''
+We go on to calculate the ``Transient Growth Rate'', which is the average from $t=0$ to $t=\omega_s^{-1}$ of the instantaneous growth rate of the ensemble averaged curve.
 In Fig.~\ref{spec_nl_transient_gamma}, we compare the Transient Growth Rate Spectra (labeled ``Trans. $\omega_s^{-1}$'') to the Turbulent Growth Rate Spectra.
@@
 
-There is one more growth rate curve in Fig.~\ref{spec_nl_transient_gamma}, labelled ``Trans $\omega_s^{-1}$ cascade.'' We include this to address the apparent downshift (in $m$) of
-the ``Trans $\omega_s^{-1}$'' curve in comparison to the turbulent curve. The downshift is another consequence of our oversimplification of the nonlinearity. Specifically, we
-used $\omega_s^{-1}$ at each $m$ for the nonlinear time scale at that same $m$. As seen in Eq.~\ref{mat_eq}, $v_{m,n}$ is affected by every $\mathbf{v}_{E,m-m',n-n'} k^{'}_\perp$, not by only
-$\mathbf{v}_{E,m,n} k_\perp$. Since we do not want to overcomplicate matters, we address this by posing a simple forward cascade model for the nonlinearity. That is, $v_{m,n}$ should be
+We include one more growth rate curve in Fig.~\ref{spec_nl_transient_gamma}, labelled ``Trans $\omega_s^{-1}$ cascade,'' to address the apparent downshift (in $m$) of
+the ``Trans $\omega_s^{-1}$'' curve in comparison to the turbulent curve. The downshift is a consequence of using $\omega_s^{-1}$ at each $m$ for the nonlinear time scale at that same $m$. 
+As seen in Eq.~\ref{mat_eq}, though, $v_{m,n}$ is affected by every $\mathbf{v}_{E,m-m',n-n'} k^{'}_\perp$, not by only $\mathbf{v}_{E,m,n} k_\perp$. 
+To address this, we pose a simple forward cascade model for the nonlinearity. That is, $v_{m,n}$ should be
 most affected by the term with $m' = m/2 \rightarrow m-m' = m/2$, corresponding to eddy mitosis.
@@
 The $n=0$ curve matches well with the $n=0$ Turbulent Growth Rate curve at $m>20$, but poorly for $m<20$, which is expected because the cascade applies only at high $m$. \emph{A priori}, we
-cannot determine the minimum $m$ for which the cascade should apply, so this method is simply explanatory.
+cannot determine the minimum $m$ for which the cascade should apply, so this method is not predictive.
 
@@
 \begin{acknowledgments}
-
+This work was supported by the National Science Foundation (Grant PHY-1202007)
 \end{acknowledgments}
@@
 
-%\bibliographystyle{unsrt}
+\bibliographystyle{phaip}
 \bibliography{refs}

### 2013-12-20 Transient_growth_prl.tex
Dec 20 2

@@
 \begin{figure}
-\centerline{\includegraphics[width=0.4\textwidth]{transient_n0_n1_growth}}
-\caption{Linear evolution of energy starting from a turbulent initial state. The $n=0$ curves have an initial period of transient growth before exponentially decaying.}
-\label{transient_n0_n1_growth}
+\centerline{\includegraphics[width=0.55\textwidth]{n0_n1_growth}}
+\caption{a) Linear evolution of energy starting from a turbulent initial state. The $n=0$ curves have an initial period of transient growth before exponentially decaying.
+b) Linear and turbulent growth rate spectra for $n=0$ (solid lines) and $n=1$ (dashed lines) Fourier components. The linear growth rates are those of the least stable eigenmodes,
+while the turbulent growth rates represent $\pdiff{E}{t}|_{\rm{lin}}/2 E$ from the nonlinear simulation. The shaded region marks the $1 \sigma$ spread in the turbulent spectrum,
+obtained from the distribution of growth rates in the nonlinear simulation.}
+\label{n0_n1_growth}
 \end{figure}
@@
 Both are transient growth processes. We see this in our simulations by following the evolution of the energy of several $m,n$ modes after turning off the nonlinearities in an already
-turbulent simulation. A few representative modes are shown in Fig.~\ref{transient_n0_n1_growth}. The linear transient growth of the filamentary $n=0$ structures is evident as their modes 
+turbulent simulation. A few representative modes are shown in Fig.~\ref{n0_n1_growth} a). The linear transient growth of the filamentary $n=0$ structures is evident as their modes 
 grow transiently before decaying exponentially at the rate of their least stable eigenmode. Such behavior is rather unituitive to some because all $n=0$ linear eigenmodes are stable, 
@@
 can occur in normal systems, but it nevertheless explains later results.
-
-\begin{figure}
-\centerline{\includegraphics[width=0.45\textwidth]{spec_nl_gamma}}
-\caption{Linear and turbulent growth rate spectra for $n=0$ (solid lines) and $n=1$ (dashed lines) Fourier components. The linear growth rates are those of the least stable eigenmodes,
-while the turbulent growth rates represent $\pdiff{E}{t}|_{\rm{lin}}/2 E$ from the nonlinear simulation. The shaded region marks the $1 \sigma$ spread in the turbulent spectrum,
-obtained from the distribution of growth rates in the nonlinear simulation.}
-\label{spec_nl_gamma}
-\end{figure}
 
@@
 For turbulence, $\gamma_t(m,n)$ is generally not equal to an eigenmode growth rate, especially in the case where the system is non-normal.
-We plot this growth rate in Fig.~\ref{spec_nl_gamma} for $n=0$ and $n=1$ components. The turbulent growth rate curves are calculated from this formula, 
+We plot this growth rate in Fig.~\ref{n0_n1_growth} b) for $n=0$ and $n=1$ components. The turbulent growth rate curves are calculated from this formula, 
 averaged over a time of about $3 \times 10^{4} \ \omega_{ci}^{-1}$ during the saturated turbulent phase of the nonlinear simulation.
 Also, we show the $1 \sigma$ contour, which indicates how greatly $\gamma_t$ varies over time.
-Fig.~\ref{spec_nl_gamma} indicates that while the linear eigenmodes are unstable for $n=1$ and stable for $n=0$,
+Fig.~\ref{n0_n1_growth} b) indicates that while the linear eigenmodes are unstable for $n=1$ and stable for $n=0$,
 the turbulent growth rates are positive for $n=0, m<50$ (subcritical-like behavior), but negative elsewhere. 
-Fig.~\ref{transient_n0_n1_growth} hinted at these characteristics, but this figure represents quantitative confirmation.
+Fig.~\ref{n0_n1_growth} a) hinted at these characteristics, but this figure represents quantitative confirmation.
 
-Our goal now is to reproduce the turbulent growth rate curves in Fig.~\ref{spec_nl_gamma} without performing nonlinear simulations. This is difficult because the growth rates depend on turbulent
+Our goal now is to reproduce the turbulent growth rate curves in Fig.~\ref{n0_n1_growth} b) without performing nonlinear simulations. This is difficult because the growth rates depend on turbulent
 structures, which for non-normal systems, may be very different from eigenmode structures, and which change in time due to transient evolution.
@@
 \begin{figure}
-\centerline{\includegraphics[width=0.5\textwidth]{m20n0_time_evs}}
+\centerline{\includegraphics[width=0.55\textwidth]{m20n0_time_evs}}
 \caption{a) The maximum growth ratio curve $G_{\rm{max}}(t)$ (dashed line) for $n=0,m=20$, and an ensemble of growth ratio curves (solid gray lines, with the solid black line the ensemble average)

### 2013-12-23 Transient_growth_prl.tex
Dec 23

@@
 
-Normal mode analysis -- the calculation of eigenvalues and eigenvectors of a linearized dynamical system -- has long been used to solve a number of problems ranging from
+Normal mode analysis -- the calculation of eigenvalues and eigenvectors of a linearized dynamical system -- has long been used to solve problems ranging from
 heat conduction along a solid bar to quantum mechanical energy states of atoms. Despite its wide-ranging success, normal mode analysis has
 failed in important instances, particularly in predicting the onset of subcritical turbulence in hydrodynamic flows. 
-The reason for this failure was not explained until the early 1990's when Trefethen and others attributed the pitfalls of normal mode analysis to the non-normality of linear operators of
+The reason for this failure was explained in the early 1990's when Trefethen and others attributed the pitfalls of normal mode analysis to the non-normality of linear operators of
 dynamical systems~\cite{trefethen1993,schmid2007}. A non-normal operator has 
@@
 
-The nonlinear simulation reveals a fascinating property of the turbulence -- it is dominated by a nonlinear instability process despite being linearly unstable to drift waves~\cite{friedman2012b,friedman2013}.
+The nonlinear simulation reveals a fascinating property of the turbulence -- 
+it is dominated by a nonlinear instability process despite being linearly unstable to drift waves~\cite{friedman2012b,friedman2013}.
 This nonlinear instability process, which was discovered by Drake et al.~\cite{drake1995} in a similar plasma model, works as follows: 
-magnetic-field-aligned ($k_\para=0$) convective filaments transport density across the equilibrium density gradient, setting up $k_\para=0$ density filaments (the same applies to the temperature). 
-These filaments are unstable to a secondary drift wave instability,
-which grows on the periphery of the filaments. 
-These drift waves, which have finite $k_\para$ and propagate primarily
-radially, nonlinearly couple to one another and reinforce the original
-convective filaments.
+magnetic-field-aligned ($k_\para=0$) convective filaments transport density across the equilibrium density gradient, setting up $k_\para=0$ density filaments. 
+These filaments are unstable to a secondary drift wave instability, which grows on the periphery of the filaments. 
+These drift waves, which have finite $k_\para$, nonlinearly couple to one another and reinforce the original convective filaments.
 
@@
 
-This convective transport of density filaments is akin to the paradigmatic ``lift-up'' mechanism in hydrodynamic shear flows whereby streamwise vortices drive streamwise streaks~\cite{krommes1999}.
+This convective transport of density filaments is akin to the paradigmatic ``lift-up'' mechanism in hydrodynamic shear flows whereby streamwise vortices drive streamwise 
+streaks~\cite{trefethen1993,krommes1999}.
 Both are transient growth processes. We see this in our simulations by following the evolution of the energy of several $m,n$ modes after turning off the nonlinearities in an already
@@
 superpositions of nonorthogonal vectors may behave unintuitively.
-Notice now that the $n=1,m=20$ mode decays transiently before growing exponentially with growth rate of the most unstable eigenmode. This isn't that interesting since this behavior
-can occur in normal systems, but it nevertheless explains later results.
+Notice now that the $n=1,m=20$ mode decays transiently before growing exponentially with growth rate of the most unstable eigenmode. This isn't necessarily due to eigenvector nonorthogonality,
+but is nevertheless unintuitive.
 
-Since the transient growth, which is behind subcritical turbulence, is a purely linear phenomenon, it has long been a goal of researchers to predict the onset of subcritical turbulence using
-only linear, non-modal calculations. It is our goal, however, to use linear calculations to understand and even predict turbulent properties.
+Since transient growth, which is responsible for subcritical turbulence, is a purely linear phenomenon, it has long been a goal of researchers to predict the onset of subcritical turbulence using
+only linear, non-modal calculations.
+It is our goal, however, to use linear calculations to go beyond predicting onset and predict fluctuation amplitudes and transport associated with saturated turbulence.
 Perhaps one of the most useful statistical properties that we can predict is the turbulent growth rate spectrum.
 To obtain a growth rate from the nonlinear simulation, we begin with the expression $\diff{E(m,n)}{t}\big|_{\rm{lin}} = Q(m,n) + D(m,n)$, 
-which accounts for the total net energy injection at each $m,n$ into the fluctuations. 
-An effective instantaneous growth rate $\gamma_t(m,n) = \diff{E(m,n)}{t}\big|_{\rm{lin}}/ 2 E(m,n)$ can then be defined. 
-This effective growth rate is calculated from the spatial structure of the fluctuations, so it is well-defined for both turbulent fluctuations and eigenmode fluctuations. If calculated
-from a linear simulation run for a sufficiently long time, $\gamma_t(m,n)$ is the same as the eigenmode growth rate of the least stable linear eigenmode for each $m,n$.
-For turbulence, $\gamma_t(m,n)$ is generally not equal to an eigenmode growth rate, especially in the case where the system is non-normal.
+which accounts for the total net energy injection into the fluctuations at each $m,n$. 
+An effective growth rate $\gamma_e(m,n) = \diff{E(m,n)}{t}\big|_{\rm{lin}}/ 2 E(m,n)$ can then be defined. 
+This growth rate is calculated from the spatial structure of the fluctuations, so it is well-defined for both turbulent fluctuations and eigenmode fluctuations. If calculated
+from a linear simulation run for a sufficiently long time, $\gamma_e(m,n)$ is the same as the spectral eigenmode growth rate $\gamma_s(m,n)$ of the least stable linear eigenmode.
+For turbulence, $\gamma_e(m,n)$ is generally not equal to an eigenmode growth rate, especially when the system is non-normal.
 We plot this growth rate in Fig.~\ref{n0_n1_growth} b) for $n=0$ and $n=1$ components. The turbulent growth rate curves are calculated from this formula, 
 averaged over a time of about $3 \times 10^{4} \ \omega_{ci}^{-1}$ during the saturated turbulent phase of the nonlinear simulation.
-Also, we show the $1 \sigma$ contour, which indicates how greatly $\gamma_t$ varies over time.
+Also, we show the $1 \sigma$ contour, which indicates how greatly $\gamma_e$ varies over time.
 Fig.~\ref{n0_n1_growth} b) indicates that while the linear eigenmodes are unstable for $n=1$ and stable for $n=0$,
@@
 
-Our goal now is to reproduce the turbulent growth rate curves in Fig.~\ref{n0_n1_growth} b) without performing nonlinear simulations. This is difficult because the growth rates depend on turbulent
-structures, which for non-normal systems, may be very different from eigenmode structures, and which change in time due to transient evolution.
-It is clear, though, that the turbulent structures are a result of both linear evolution and nonlinear energy transfer. 
-If we imagine that the nonlinearities randomize the turbulent structures at each wavenumber while the linearities evolve the structures
-deterministically, we can develop a procedure for choosing proper spatial structures to use in calculating effective growth rates.
-To elucidate this procedure, we begin by taking Eqs.~\ref{ni_eq}-\ref{te_eq} and Fourier decomposing in the azimuthal and axial directions.
-Then, we discritize in the radial direction and approximate radial derivatives with finite differences. 
-The resulting system of equations may be written in matrix form:
+Our goal now is to reproduce the turbulent growth rate curves in Fig.~\ref{n0_n1_growth} b) without performing nonlinear simulations. 
+The difficulty is in determining what kinds of spatial structures will dominate the turbulence, or alternatively, how will nonlinear energy transfer affect transient growth properties? 
+We use the anzatz that nonlinearities randomize the turbulent spatial structure at each wavenumber with a period of 
+one eddy decorrelation time to develop a procedure for calculating effective growth rates. 
+We begin by taking Eqs.~\ref{ni_eq}-\ref{te_eq} and Fourier decomposing in the azimuthal and axial directions.  Then, we discritize in the radial direction and
+approximate radial derivatives with finite differences.  The resulting system of equations may be written in matrix form:
 
@@
 where $\mathbf{v}_{m,n} = \left( N(r), \vpe(r), \phi(r), T_e(r) \right)_{m,n}^{T}$ is the state of the system,
-and $\mathbf{B}_{m,n}$ and $\mathbf{C}_{m,n}$ are coefficient matrices that include the equilibrium information and finite difference coefficients.
-The first term on the RHS represents the linearities,
-and the second term the nonlinearities.
-The equations for different $m$ and $n$ are separated, although the nonlinearity couples the different $m$'s and $n$'s. 
-In other words, eigenmodes with different $m$ and $n$ are orthogonal and linearly non-interacting. Note that for each $m,n$, there exist $4 \times N_r$ linearly
-independent, but nonorthogonal eigenvectors. Hence forth, we drop the $m,n$ subscripts.
+and $\mathbf{B}_{m,n}$ and $\mathbf{C}_{m,n}$ are coefficient matrices that include the equilibrium information and finite difference coefficients. The first term on the RHS represents the linearities
+and the second term the nonlinearities. Note that for each $m,n$, there exist $4 \times N_r$ linearly independent, but nonorthogonal eigenvectors. Hence forth, we drop the $m,n$ subscripts.
 
 In order to use non-modal analysis to calculate growth rates and other measures, one must choose a norm and inner product with which to work. While any choice of
-inner product is possible, including that which orthogonalizes all eigenvectors, a physically relevant one such as an energy inner product 
-is generally preferred~\cite{camargo1998,schmid2007,camporeale2010}. 
+inner product is possible, a physically relevant one such as an energy inner product is generally preferred~\cite{camargo1998,schmid2007,camporeale2010}. 
 Recall that the inner product of two vectors may be written $\left< \mathbf{x},\mathbf{y} \right> = \mathbf{y}^{*} \mathbf{M} \mathbf{x}$,
@@
 \eeq
-This ratio is bounded from above by $G_{\rm{max}}(t) = ||e^{\mathbf{A} t}||$. $G_{\rm{max}}(t) = e^{\gamma_s t}$, where $\gamma_s$ is the spectral growth rate
-if and only if $\mathbf{A}$ is normal~\cite{schmid2007}. Otherwise, $G_{\rm{max}}(t) > e^{\gamma_s t}$.
+This ratio is bounded from above by $G_{\rm{max}}(t) = ||e^{\mathbf{A} t}||$. $G_{\rm{max}}(t) = e^{\gamma_s t}$ if and only if $\mathbf{A}$ is normal~\cite{schmid2007}. 
+Otherwise, $G_{\rm{max}}(t) > e^{\gamma_s t}$.
 
-It is common practice in normal mode analysis to look for the least stable eigenmode, the idea being that eventually the fastest growing structure will overcome the others.
+It is common practice in normal mode analysis to look for the least stable eigenmode because the fastest growing structure will eventually overcome the others.
 For the non-modal case, it is common to study the properties of $G_{\rm{max}}(t)$, because if $G_{\rm{max}}(t) > 1$ at any time, there is a possibility of amplification, even if it is only transient.
-This allows for the possibility of exciting subcritical turbulence, which is where non-modal analysis is most often applied. 
+Subcritical turbulence may then be possible.
 However, it can be misleading to study only $G_{\rm{max}}(t)$ when predicting specific properties of turbulence because
@@
 In non-normal systems, unlike in the normal systems, optimal structures don't amplify themselves, rather, they evolve while increasing the total fluctuating energy.
+For our purposes, it is more useful to study $G(t)$ curves that use initially random starting conditions.
 
-We illustrate $G_{\rm{max}}(t)$ along with $G(t)$ at $m=20, n=0$ for several random initial conditions in Fig.~\ref{m20n0_time_evs} a).. Despite the fact that all eigenmodes at this
+We illustrate $G_{\rm{max}}(t)$ along with $G(t)$ at $m=20, n=0$ for several random initial conditions in Fig.~\ref{m20n0_time_evs} a). Despite the fact that all eigenmodes at this
 wavenumber are stable, there is the possibility for transient amplification by a factor of $20$. Furthermore, a sample of randomly initialized vectors are all transiently amplified, but none display
@@
 To this effect, we note that the advective nonlinearity in Eq.~\ref{mat_eq} has the form of the state vector divided by a time $\tau_{\rm{nl}} \sim (v_E k_\perp)^{-1}$. This nonlinear
-time scale is generally associated with the eddy turnover time or the time in which eddies become decorrelated. We therefore present a heuristical model of the nonlinearities 
+time scale is generally associated with the eddy turnover or decorrelation time. We therefore present a heuristic model of the nonlinearities 
 as a randomizing force that acts on this characteristic nonlinear time scale.
@@
 While the nonlinearities obviously do not affect the system in discrete time steps, this simplification provides a way to use linear non-modal analysis in a predictive manner.
-In other words, there are certain spatial structures associated with the curves in Fig.~\ref{m20n0_time_evs} b) that change as a function of time. We consider only those structures from $t=0$ (where
-the structures are completely random) to $t = \tau_{\rm{nl}}$, and associate those structures with turbulent structures.
+Under this assumption, transient growth occurs only for one decorrelation time and the turbulent growth rate is calculated as the average growth rate experienced over one decorrelation time.
 
-Now, since we do not know \emph{a priori} at what level the turbulence will saturate, we cannot calculate the relative magnitude of $\tau_{\rm{nl}}$ from $(v_E k_\perp)^{-1}$ directly.
+Since we do not know \emph{a priori} at what level the turbulence will saturate, we cannot calculate the relative magnitude of $\tau_{\rm{nl}}$ from $(v_E k_\perp)^{-1}$ directly.
 Thus, we invoke the conjecture of \emph{critical balance}, which posits that the nonlinear time scale equals the linear time scale at all spatial scales~\cite{schekochihin2012}. 
 We can then estimate the nonlinear time scale as $\tau_{\rm{nl}} = \omega_s^{-1}$ -- the inverse of the linear frequency (at each $m,n$).
-However, $\omega_s = 0$ for $n=0$ eigenmodes, so we are forced to use $\omega_s$ for the fastest growing $n=1$ eigenmode at each $m$ to get a meaningful time scale. This is not wholly
+However, $\omega_s = 0$ for $n=0$ linear eigenmodes, so we are forced to use $\omega_s$ for the fastest growing $n=1$ eigenmode at each $m$ to get a meaningful time scale. This is not wholly
 unjustified, however, due to the inherent nonlinear coupling of different modes, which is evident in Eq.~\ref{mat_eq}.
-From Fig.~\ref{m20n0_time_evs} b), it is clear that the transient growth is cut off well before optimal amplification occurs. 
-But as reflected by our procedure, there is no guarantee that the nonlinearities will allow complete non-modal amplification.
+In Fig.~\ref{m20n0_time_evs} b), this time scale is indicated -- from this it is clear that the transient growth is cut off well before optimal amplification occurs.
 
-We go on to calculate the ``Transient Growth Rate'', which is the average from $t=0$ to $t=\omega_s^{-1}$ of the instantaneous growth rate of the ensemble averaged curve.
+We define the ``Transient Growth Rate'' as the average from $t=0$ to $t=\omega_s^{-1}$ of the instantaneous growth rate of the ensemble averaged growth curve.
 In Fig.~\ref{spec_nl_transient_gamma}, we compare the Transient Growth Rate Spectra (labeled ``Trans. $\omega_s^{-1}$'') to the Turbulent Growth Rate Spectra.
-The $n=0$ curves in Fig.~\ref{spec_nl_transient_gamma} a) are fairly similar, especially when compared to the eigenmode spectra.
-The $n=1$ curves in Fig.~\ref{spec_nl_transient_gamma} b), on the other hand, are not very similar, although they trend with $m$ in the same way, which is very different from the linear eigenmode trend.
+Although the Transient Growth Rate Spectrum for $n=0$ is downshifted in $m$ number relative to the spectrum derived from the nonlinear simulation, the two curves in Fig.~\ref{spec_nl_transient_gamma} a)
+are fairly similar in shape and magnitude.
+The $n=1$ curves in Fig.~\ref{spec_nl_transient_gamma} b), on the other hand, are not very similar, although the Transient Growth Rate correctly predicts energy damping in the region where
+eigenmode growth is predicted.
 The difference between the $n=1$ transient and turbulent growth rate spectra is due to nonlinear effects not captured by the simple transient calculation procedure. In particular, the nonlinear
-simulation has a strong energy transfer from $n=0$ to $n=1$ Fourier components which is then dissipated by electron-ion collisions as the electrons stream along the field lines~\cite{friedman2012b}.
+simulation has a strong energy transfer from $n=0$ to $n=1$ which is then dissipated by electron-ion collisions as the electrons stream along the field lines~\cite{friedman2012b}.
 
@@
 Then, $\mathbf{v}_{E,m-m',n-n'} k^{'}_\perp \sim \mathbf{v}_{E,m/2} k^{'}_\perp/2 \sim \omega_{s,m/2}^{-1}$,
-so we use $\omega_s^{-1}$ at $m/2$ for the randomizing time scale at $m$ in deriving the ``Trans $\omega_s^{-1}$ cascade'' curves in Fig.~\ref{spec_nl_transient_gamma}. 
+so we use $\omega_s^{-1}$ at $m/2$ for the randomizing time scale at $m$ in deriving the ``Trans $\omega_s^{-1}$ cascade'' curves. 
 The $n=0$ curve matches well with the $n=0$ Turbulent Growth Rate curve at $m>20$, but poorly for $m<20$, which is expected because the cascade applies only at high $m$. \emph{A priori}, we
@@
 Nevertheless, our randomizing transient method, which requires only linear calculations, reproduces energetic growth properties of the turbulence that are not captured by eigenmode analysis.
-Clearly, when the linear systems are highly non-normal, transient growth rate analyses should be performed in place of (or in addition to) linear eigenmode analyses. 
+Clearly, when the linear systems are highly non-normal, transient growth rate analyses should be preferred over linear eigenmode analyses. 
 Furthermore, the transient growth rates should be used in the kinds of predictions that are currently made with eigenmode growth rates. 
-For instance, quasilinear theory uses the most unstable eigenmode growth rate $\gamma_{s,max}$ to predict the turbulent saturation levels through the mixing length formula $\gamma/k_\perp^2$. 
-Using this formula with $\gamma_{s,\rm{max}} = 0.002$ and $m=60$, one would predict a turbulent saturation level on the order of $0.1 \%$. 
+For instance, quasilinear theory uses the most unstable eigenmode growth rate $\gamma_{s,max}$ to predict the turbulent saturation levels using a mixing-length argument 
+(saturated amplitude $\sim \gamma/k_\perp^2$). 
+Using this formula with $\gamma_{s,\rm{max}} = 0.002$ and $m=60$, one would predict a turbulent saturation level on the order of $0.1 \%$ which is wholly inconsistent with experiment and simulation. 
 On the other hand, using this formula with the transient values $\gamma_{t,\rm{max}} = 0.003, \ m=10$, we predict the saturation level on the order of $10 \%$,
-which matches the saturation level in the nonlinear simulation.
+which is consistent with the nonlinear simulation and with experimental measurements.
 
-In summary, non-modal linear calculations are more informative than normal mode calculations when the linear system under consideration is highly non-normal. In the specific case of a low-flow
-LAPD experiment, a nonlinear instability with linear $k_\para=0$ drive dominates the energy dynamics of the turbulence. While this is a surprising result in light of normal mode analysis,
-it is not at all surprising given the system's non-modal behavior, which we show with a technique that takes into account the effect of the nonlinearities. 
-In general, non-modal initial-value analysis is difficult to quantify and make predictive, but using some simple nonlinear modelling, we have shown that it is possible. Future studies will
-attempt to improve on the nonlinear modelling and expand the use of this non-modal technique to other non-normal turbulence models.
-
-\begin{acknowledgments}
-This work was supported by the National Science Foundation (Grant PHY-1202007)
-\end{acknowledgments}
+In summary, we present a procedure in which prediction of turbulent properties, in particular the turbulent growth rate spectrum, can be
+made using non-modal linear calculations. In the specific case of a low-flow LAPD experiment, this procedure captures the behavior of a
+nonlinear instability that is observed to dominate the dynamics of the turbulence in a well-validated nonlinear simulation.  In general,
+non-modal initial-value analysis is difficult to quantify and make predictive, but using some simple nonlinear modelling, we have shown
+that it is possible. Future studies will attempt to improve on the nonlinear modelling and expand the use of this non-modal technique to
+other non-normal turbulence models.  This work was supported by the National Science Foundation (Grant PHY-1202007)
 

### 2013-12-28 Transient_growth_prl.tex
Dec 27

@@
 
-Normal mode analysis -- the calculation of eigenvalues and eigenvectors of a linearized dynamical system -- has long been used to solve problems ranging from
-heat conduction along a solid bar to quantum mechanical energy states of atoms. Despite its wide-ranging success, normal mode analysis has
-failed in important instances, particularly in predicting the onset of subcritical turbulence in hydrodynamic flows. 
+Normal mode analysis -- the calculation of eigenvalues and eigenvectors of a linearized dynamical system -- has been used to solve many problems over the years.
+Despite its wide-ranging success, it has failed in important instances, particularly in predicting the onset of subcritical turbulence in hydrodynamic flows. 
 The reason for this failure was explained in the early 1990's when Trefethen and others attributed the pitfalls of normal mode analysis to the non-normality of linear operators of
@@
 While the concepts behind this technique are general enough to be applied to many nonlinear dynamical systems, the details vary for each case, 
-so we restrict our treatment to one particular turbulence model. The technique reproduces the turbulent growth rate spectrum of the direct nonlinear simulation fairly well, especially
-in comparison to the linear spectrum, which has little bearing on the turbulence.
+so we restrict our treatment to one particular turbulence model. For this model, the technique reproduces the turbulent growth rate spectrum of the direct nonlinear simulation fairly well, especially
+in comparison to the linear eigenmode spectrum.
 
@@
 normalized to the ion sound gyroradius $\rho_s$, times to the ion cyclotron time $\omega_{ci}^{-1}$, velocities to the sound speed $c_s$, densities to the equilibrium peak density, and electron
-temperatures and potentials to the equilibrium peak electron temperature. The profiles $N_0$ and $T_{e0}$ and other parameters are taken from experimental measurements. $\phi_0$ 
-is absent in the equations because we model one particular experiment in which the equilibrium radial electric field was reduced to near zero value using a
-limiter biasing technique~\cite{schaffner2012}.
+temperatures and potentials to the equilibrium peak electron temperature. The profiles $N_0$ and $T_{e0}$ and other parameters are taken from experimental 
+measurements~\cite{schaffner2012,friedman2012b,friedman2013}. 
+%$\phi_0$  is absent in the equations because we model one particular experiment in which the equilibrium radial electric field was reduced to near zero value using a limiter biasing technique~\cite{schaffner2012}.
 
@@
 magnetic-field-aligned ($k_\para=0$) convective filaments transport density across the equilibrium density gradient, setting up $k_\para=0$ density filaments. 
-These filaments are unstable to a secondary drift wave instability, which grows on the periphery of the filaments. 
+These filaments are unstable to a secondary drift waves, which grow on the periphery of the filaments. 
 These drift waves, which have finite $k_\para$, nonlinearly couple to one another and reinforce the original convective filaments.
@@
 where $m$ and $n$ represent the azimuthal and axial Fourier mode numbers. 
-Since Eqs.~\ref{ni_eq}-\ref{te_eq} are not explicitly dependent on $\theta$ and $z$, the energy equation is separable for every $m$ and $n$. 
+%Since Eqs.~\ref{ni_eq}-\ref{te_eq} are not explicitly dependent on $\theta$ and $z$, the energy equation is separable for every $m$ and $n$. 
 $Q$ represents the linear energy exchange with the equilibrium (convective transport), $D$ the dissipation from irreversible processes
@@
 turbulent simulation. A few representative modes are shown in Fig.~\ref{n0_n1_growth} a). The linear transient growth of the filamentary $n=0$ structures is evident as their modes 
-grow transiently before decaying exponentially at the rate of their least stable eigenmode. Such behavior is rather unituitive to some because all $n=0$ linear eigenmodes are stable, 
+grow transiently before decaying exponentially at the rate of their least stable eigenmode. Such behavior is rather unituitive because all $n=0$ linear eigenmodes are stable, 
 and the normal mode paradigm has conditioned us to believe that a linear superposition of such eigenmodes should decay. But as the hydrodynamics community discovered, linear
 superpositions of nonorthogonal vectors may behave unintuitively.
-Notice now that the $n=1,m=20$ mode decays transiently before growing exponentially with growth rate of the most unstable eigenmode. This isn't necessarily due to eigenvector nonorthogonality,
-but is nevertheless unintuitive.
+Notice finally that the $n=1,m=20$ mode decays transiently before growing exponentially with growth rate of the most unstable eigenmode.
+%This isn't necessarily due to eigenvector nonorthogonality, but is nevertheless unintuitive.
 
-Since transient growth, which is responsible for subcritical turbulence, is a purely linear phenomenon, it has long been a goal of researchers to predict the onset of subcritical turbulence using
+Since transient growth is a purely linear phenomenon, it has been a goal of researchers to predict the onset of subcritical turbulence using
 only linear, non-modal calculations.
@@
 This growth rate is calculated from the spatial structure of the fluctuations, so it is well-defined for both turbulent fluctuations and eigenmode fluctuations. If calculated
-from a linear simulation run for a sufficiently long time, $\gamma_e(m,n)$ is the same as the spectral eigenmode growth rate $\gamma_s(m,n)$ of the least stable linear eigenmode.
+from a linear simulation that is run for a sufficiently long time, $\gamma_e(m,n)$ is the same as the spectral eigenmode growth rate $\gamma_s(m,n)$ of the least stable linear eigenmode.
 For turbulence, $\gamma_e(m,n)$ is generally not equal to an eigenmode growth rate, especially when the system is non-normal.
@@
 Our goal now is to reproduce the turbulent growth rate curves in Fig.~\ref{n0_n1_growth} b) without performing nonlinear simulations. 
-The difficulty is in determining what kinds of spatial structures will dominate the turbulence, or alternatively, how will nonlinear energy transfer affect transient growth properties? 
-We use the anzatz that nonlinearities randomize the turbulent spatial structure at each wavenumber with a period of 
-one eddy decorrelation time to develop a procedure for calculating effective growth rates. 
-We begin by taking Eqs.~\ref{ni_eq}-\ref{te_eq} and Fourier decomposing in the azimuthal and axial directions.  Then, we discritize in the radial direction and
+To accomplish this, we use the anzatz that nonlinearities randomize the turbulent spatial structure at each wavenumber on a time scale of 
+one eddy decorrelation time, while the linearities evolve the spatial structures deterministically. From this we can calculate a ``Transient Growth Rate'' spectrum,
+which is our prediction of the turbulent growth rate spectrum. To illustrate,
+we begin by taking Eqs.~\ref{ni_eq}-\ref{te_eq} and Fourier decomposing in the azimuthal and axial directions.  Then, we discritize in the radial direction and
 approximate radial derivatives with finite differences.  The resulting system of equations may be written in matrix form:
@@
 Recall that the inner product of two vectors may be written $\left< \mathbf{x},\mathbf{y} \right> = \mathbf{y}^{*} \mathbf{M} \mathbf{x}$,
-where $^*$ stands for the conjugate transpose, and $\mathbf{M}$ is a positive-definite matrix. $\mathbf{M}$ is chosen so that 
+where $^*$ stands for the conjugate transpose, and $\mathbf{M}$ is a positive-definite matrix. We choose $\mathbf{M}$ so that 
 $||\mathbf{v}|| = \sqrt{\left< \mathbf{v},\mathbf{v} \right>} = \sqrt{E}$, where $E$ is the energy. Furthermore, it is convenient in
-computations to use the $L_2$-norm, $||\mathbf{u}||_2 = \sqrt{\sum_i |u_i|^2}$.
-These requirements can be accomplished through the change of variables $\mathbf{u} = \mathbf{M}^{\frac{1}{2}} \mathbf{v}$.
+computations to use the $L_2$-norm, $||\mathbf{u}||_2 = \sqrt{\sum_i |u_i|^2}$. This can be accomplished through the change of variables $\mathbf{u} = \mathbf{M}^{\frac{1}{2}} \mathbf{v}$.
 Then the linear portion of Eq.~\ref{mat_eq} becomes
@@
 
-It is common practice in normal mode analysis to look for the least stable eigenmode because the fastest growing structure will eventually overcome the others.
-For the non-modal case, it is common to study the properties of $G_{\rm{max}}(t)$, because if $G_{\rm{max}}(t) > 1$ at any time, there is a possibility of amplification, even if it is only transient.
-Subcritical turbulence may then be possible.
-However, it can be misleading to study only $G_{\rm{max}}(t)$ when predicting specific properties of turbulence because
-$G_{\rm{max}}(t)$ is only the upper envelope of all possible $G(t)$ curves. No one particular initial condition $\mathbf{u}(0)$ evolves along $G_{\rm{max}}(t)$. 
-Furthermore, it isn't obvious what kind of spatial structure or structures will come to dominate a turbulent system. 
-In non-normal systems, unlike in the normal systems, optimal structures don't amplify themselves, rather, they evolve while increasing the total fluctuating energy.
-For our purposes, it is more useful to study $G(t)$ curves that use initially random starting conditions.
-
-We illustrate $G_{\rm{max}}(t)$ along with $G(t)$ at $m=20, n=0$ for several random initial conditions in Fig.~\ref{m20n0_time_evs} a). Despite the fact that all eigenmodes at this
-wavenumber are stable, there is the possibility for transient amplification by a factor of $20$. Furthermore, a sample of randomly initialized vectors are all transiently amplified, but none display
-optimal growth as compared to $G_{\rm{max}}(t)$.
-It is extremely unlikely to choose a random initial vector that either optimally amplifies the energy at any time or one that monotonically decays, which would be the case
-if the initial vector were one of the linear eigenvectors.
-
 \begin{figure}
@@
 
-Non-modal analysis has likely remained a qualitative endeavor due to the time-dependent growth properties of initial-value analysis as well as its intrinsic dependence on initial conditions.
+It is common practice in normal mode analysis to look for the least stable eigenmode. 
+For the non-modal case, it is common to study the properties of $G_{\rm{max}}(t)$ because if $G_{\rm{max}}(t) > 1$ at any time, fluctuations may be amplified, leading to subcritical turbulence.
+However, it can be misleading to study only $G_{\rm{max}}(t)$ when predicting specific properties of turbulence because
+$G_{\rm{max}}(t)$ is only the upper envelope of all possible $G(t)$ curves. No one particular initial condition $\mathbf{u}(0)$ evolves along $G_{\rm{max}}(t)$. 
+Furthermore, it isn't obvious what kind of spatial structure or structures will come to dominate a turbulent system.
+In non-normal systems, unlike in the normal systems, optimal structures don't amplify themselves, rather, they evolve while increasing the total fluctuating energy.
+
+We illustrate $G_{\rm{max}}(t)$ along with $G(t)$ at $m=20, n=0$ for several random initial conditions in Fig.~\ref{m20n0_time_evs} a). Despite the fact that all eigenmodes at this
+wavenumber are stable, there is the possibility for transient amplification by a factor of $20$. Furthermore, a sample of randomly initialized vectors are all transiently amplified, but none display
+optimal growth. It is extremely unlikely to choose a random initial vector that either optimally amplifies the energy at any time or one that monotonically decays, which would be the case
+if the initial vector were one of the linear eigenvectors.
+
 The key to quantifying non-modal analysis and making it predictive is to successfully model the effect that the nonlinearities have on the transient linear processes. 
@@
 Specifically, we start with a random initial condition, evolve it linearly for a time $\tau_{\rm{nl}}$, and then re-randomized the system.
-While the nonlinearities obviously do not affect the system in discrete time steps, this simplification provides a way to use linear non-modal analysis in a predictive manner.
-Under this assumption, transient growth occurs only for one decorrelation time and the turbulent growth rate is calculated as the average growth rate experienced over one decorrelation time.
+To estimate $\tau_{\rm{nl}}$, we invoke the conjecture of \emph{critical balance}, which posits that the nonlinear time scale equals the linear time scale at all spatial scales~\cite{schekochihin2012}.
+For the linear time scale, we take the inverse of the linear frequency (at each $m,n$) -- $\tau_{\rm{nl}} = \omega_s^{-1}$.
+However, $\omega_s = 0$ for $n=0$ linear eigenmodes, so we are forced to use $\omega_s$ for the fastest growing $n=1$ eigenmode at each $m$ to get a meaningful time scale. This is not wholly
+unjustified, however, due to the inherent nonlinear coupling of different modes, evident in Eq.~\ref{mat_eq}.
+In Fig.~\ref{m20n0_time_evs} b), this time scale is indicated. Note that we cut off the transient growth well before optimal amplification occurs, at least for this wavenumber.
 
-Since we do not know \emph{a priori} at what level the turbulence will saturate, we cannot calculate the relative magnitude of $\tau_{\rm{nl}}$ from $(v_E k_\perp)^{-1}$ directly.
-Thus, we invoke the conjecture of \emph{critical balance}, which posits that the nonlinear time scale equals the linear time scale at all spatial scales~\cite{schekochihin2012}. 
-We can then estimate the nonlinear time scale as $\tau_{\rm{nl}} = \omega_s^{-1}$ -- the inverse of the linear frequency (at each $m,n$).
-However, $\omega_s = 0$ for $n=0$ linear eigenmodes, so we are forced to use $\omega_s$ for the fastest growing $n=1$ eigenmode at each $m$ to get a meaningful time scale. This is not wholly
-unjustified, however, due to the inherent nonlinear coupling of different modes, which is evident in Eq.~\ref{mat_eq}.
-In Fig.~\ref{m20n0_time_evs} b), this time scale is indicated -- from this it is clear that the transient growth is cut off well before optimal amplification occurs.
-
-We define the ``Transient Growth Rate'' as the average from $t=0$ to $t=\omega_s^{-1}$ of the instantaneous growth rate of the ensemble averaged growth curve.
-In Fig.~\ref{spec_nl_transient_gamma}, we compare the Transient Growth Rate Spectra (labeled ``Trans. $\omega_s^{-1}$'') to the Turbulent Growth Rate Spectra.
-Although the Transient Growth Rate Spectrum for $n=0$ is downshifted in $m$ number relative to the spectrum derived from the nonlinear simulation, the two curves in Fig.~\ref{spec_nl_transient_gamma} a)
+We define the ``Transient Growth Rate'' $\gamma_{\rm{TR}}$ as the average from $t=0$ to $t=\omega_s^{-1}$ of the instantaneous growth rate of the ensemble averaged growth curve.
+In Fig.~\ref{spec_nl_transient_gamma}, we compare the $\gamma_{\rm{TR}}$ spectrum (labeled ``Trans. $\omega_s^{-1}$'') to the turbulent spectrum.
+Although the $\gamma_{\rm{TR}}$ spectrum for $n=0$ is downshifted in $m$ number relative to the spectrum derived from the nonlinear simulation, the two curves in Fig.~\ref{spec_nl_transient_gamma} a)
 are fairly similar in shape and magnitude.
-The $n=1$ curves in Fig.~\ref{spec_nl_transient_gamma} b), on the other hand, are not very similar, although the Transient Growth Rate correctly predicts energy damping in the region where
+The $n=1$ curves in Fig.~\ref{spec_nl_transient_gamma} b), on the other hand, are not very similar, but $\gamma_{\rm{TR}}$ correctly predicts energy damping in the region where
 eigenmode growth is predicted.
 The difference between the $n=1$ transient and turbulent growth rate spectra is due to nonlinear effects not captured by the simple transient calculation procedure. In particular, the nonlinear
-simulation has a strong energy transfer from $n=0$ to $n=1$ which is then dissipated by electron-ion collisions as the electrons stream along the field lines~\cite{friedman2012b}.
+simulation has strong energy transfer from $n=0$ to $n=1$ which dissipates by electron-ion collisions~\cite{friedman2012b}.
 
 We include one more growth rate curve in Fig.~\ref{spec_nl_transient_gamma}, labelled ``Trans $\omega_s^{-1}$ cascade,'' to address the apparent downshift (in $m$) of
-the ``Trans $\omega_s^{-1}$'' curve in comparison to the turbulent curve. The downshift is a consequence of using $\omega_s^{-1}$ at each $m$ for the nonlinear time scale at that same $m$. 
-As seen in Eq.~\ref{mat_eq}, though, $v_{m,n}$ is affected by every $\mathbf{v}_{E,m-m',n-n'} k^{'}_\perp$, not by only $\mathbf{v}_{E,m,n} k_\perp$. 
-To address this, we pose a simple forward cascade model for the nonlinearity. That is, $v_{m,n}$ should be
-most affected by the term with $m' = m/2 \rightarrow m-m' = m/2$, corresponding to eddy mitosis.
-Then, $\mathbf{v}_{E,m-m',n-n'} k^{'}_\perp \sim \mathbf{v}_{E,m/2} k^{'}_\perp/2 \sim \omega_{s,m/2}^{-1}$,
-so we use $\omega_s^{-1}$ at $m/2$ for the randomizing time scale at $m$ in deriving the ``Trans $\omega_s^{-1}$ cascade'' curves. 
-The $n=0$ curve matches well with the $n=0$ Turbulent Growth Rate curve at $m>20$, but poorly for $m<20$, which is expected because the cascade applies only at high $m$. \emph{A priori}, we
-cannot determine the minimum $m$ for which the cascade should apply, so this method is not predictive.
+the ``Trans $\omega_s^{-1}$'' curve. As seen in Eq.~\ref{mat_eq}, $v_{m,n}$ is affected by every $\mathbf{v}_{E,m-m',n-n'} k^{'}_\perp$, not by only $\mathbf{v}_{E,m,n} k_\perp$ as we
+assumed when using $\omega_s^{-1}$. Thus, we pose a simple forward cascade (eddy mitosis) model for the nonlinearity.
+We use $\omega_s^{-1}$ at $m/2$ rather than $m$ for the randomizing time scale at each $m$ in deriving the ``Trans $\omega_s^{-1}$ cascade'' curves. 
+The $n=0$ curve matches well with the $n=0$ Turbulent Growth Rate curve at $m>20$, but poorly for $m<20$ as expected since the cascade applies only at high $m$. 
+We cannot determine \emph{a priori} the minimum $m$ for which the cascade should apply, so this method is not predictive.
 
@@
 \centerline{\includegraphics[width=0.5\textwidth]{spec_nl_transient_gamma}}
-\caption{The linear, turbulent, and transient growth rate spectra.
-The transient growth rate spectra are calculated by the average growth rate over a time $\omega_s^{-1}(m,n=1)$ for the ensemble average
+\caption{The linear, turbulent, and transient growth rate spectra. The transient spectra are calculated by the average growth rate over a time $\omega_s^{-1}(m,n=1)$ for the ensemble average
 growth ratio curves with the cascade curve using $\omega_s^{-1}$ at $m/2$ rather than $m$.}
@@
 
-Nevertheless, our randomizing transient method, which requires only linear calculations, reproduces energetic growth properties of the turbulence that are not captured by eigenmode analysis.
-Clearly, when the linear systems are highly non-normal, transient growth rate analyses should be preferred over linear eigenmode analyses. 
-Furthermore, the transient growth rates should be used in the kinds of predictions that are currently made with eigenmode growth rates. 
+Nevertheless, our non-modal method, which requires only linear calculations, reproduces energetic growth properties of the turbulence not captured by eigenmode analysis.
+Furthermore, when linear systems are highly non-normal, transient growth rates should be used in the kinds of predictions that are currently made with eigenmode growth rates. 
 For instance, quasilinear theory uses the most unstable eigenmode growth rate $\gamma_{s,max}$ to predict the turbulent saturation levels using a mixing-length argument 
 (saturated amplitude $\sim \gamma/k_\perp^2$). 
-Using this formula with $\gamma_{s,\rm{max}} = 0.002$ and $m=60$, one would predict a turbulent saturation level on the order of $0.1 \%$ which is wholly inconsistent with experiment and simulation. 
-On the other hand, using this formula with the transient values $\gamma_{t,\rm{max}} = 0.003, \ m=10$, we predict the saturation level on the order of $10 \%$,
-which is consistent with the nonlinear simulation and with experimental measurements.
+Using this formula with $\gamma_{s,\rm{max}} = 0.002$ and $m=60$, one predicts a turbulent saturation level of $\sim 0.1 \%$ which is wholly inconsistent with experiment and simulation. 
+On the other hand, using this formula with the transient values $\gamma_{\rm{TR, max}} = 0.003, \ m=10$, we predict the saturation level of $\sim 10 \%$,
+which is consistent with the nonlinear simulation and experimental measurements.
 
-In summary, we present a procedure in which prediction of turbulent properties, in particular the turbulent growth rate spectrum, can be
-made using non-modal linear calculations. In the specific case of a low-flow LAPD experiment, this procedure captures the behavior of a
-nonlinear instability that is observed to dominate the dynamics of the turbulence in a well-validated nonlinear simulation.  In general,
+In summary, we present a procedure for predicting the turbulent growth rate spectrum using non-modal linear calculations. 
+In the case of an LAPD experiment, this procedure captures the behavior of a nonlinear instability that dominates the dynamics of the turbulence.  In general,
 non-modal initial-value analysis is difficult to quantify and make predictive, but using some simple nonlinear modelling, we have shown
-that it is possible. Future studies will attempt to improve on the nonlinear modelling and expand the use of this non-modal technique to
-other non-normal turbulence models.  This work was supported by the National Science Foundation (Grant PHY-1202007)
+that it is possible. Future studies will attempt to improve on the nonlinear modelling and expand the use of this technique to other non-normal turbulence models.
+
+This work was supported by the National Science Foundation (Grant PHY-1202007)
 

### 2014-02-23 Referee_Response1/referee_response.tex
Feb 22

\documentclass[12pt]{article}
\usepackage{graphicx}
\setlength{\parindent}{0in}

% General Latex  --------------------------------------------------
\def\beq{\begin{equation}}
\def\eeq{\end{equation}}
\def\beqar{\begin{eqnarray}}
\def\eeqar{\end{eqnarray}}
\def\nn{\nonumber}
\def\ol{\overline}
\def\para{\parallel}

% Operators  ------------------------------------------------------
\newcommand{\diff}[2]{\frac{d#1}{d#2}}
\newcommand{\diffs}[2]{\frac{d^2#1}{d#2^2}}
\newcommand{\pdiff}[2]{\frac{\partial#1}{\partial#2}}
\newcommand{\pdiffs}[2]{\frac{\partial^2#1}{\partial#2^2}}
\newcommand{\pdiffxy}[3]{\frac{\partial^2#1}{\partial#2 \partial#3}}
\newcommand{\pdt}{\partial_t}
\newcommand{\pdr}{\partial_r}
\newcommand{\pdth}{\partial_\theta}
\newcommand{\pdrr}{\partial^2_r}

\newcommand{\enum}[2]{{#1}\times10^{#2}} % 4.2x10^{3} = \enum{4.2}{3}

\newcommand{\vect}[1]{{\bf #1}}
%\newcommand{\vect}{\overrightarrow}
%\newcommand{\vect}{\vec}
\def\div{\nabla\cdot}
\def\grad{\nabla}
\def\curl{\nabla\times}
\newcommand{\gradpar}{\grad_\parallel}
\newcommand{\gradperp}{\grad_\perp}
\newcommand{\gradr}{\grad_r}
\newcommand{\defeq}{\ensuremath{\stackrel{\text{\tiny def}}{=}}}

\newcommand{\savg}[1]{\left<{#1}\right>}
\newcommand{\vavg}[1]{\left<{#1}\right>_V}
\newcommand{\thavg}[1]{\left<{#1}\right>_\theta}

% Variable names  -------------------------------------------------
\newcommand{\vpar} {v_\parallel}
\newcommand{\Apar} {A_\parallel}
\newcommand{\jpar} {j_\parallel}
\newcommand{\kpar} {k_\parallel}
\newcommand{\kperp} {k_\perp }
\newcommand{\vperp} {v_\perp }
\newcommand{\kthe}{k_\theta}

\newcommand{\Evec}{\ensuremath{\boldsymbol{{\rm E}}}}
\newcommand{\Bvec}{\ensuremath{\boldsymbol{{\rm B}}}}
\newcommand{\Jvec}{\ensuremath{\boldsymbol{{\rm J}}}}
\newcommand{\Fvec}{\ensuremath{\boldsymbol{{\rm F}}}}
\newcommand{\fvec}{\ensuremath{\boldsymbol{{\rm f}}}}
\newcommand{\vE}{\ensuremath{\boldsymbol{{\rm v}_{E}}}}
\newcommand{\bo}{\ensuremath{\boldsymbol{{\rm b}_0}}}
\newcommand{\bvec}{\ensuremath{\boldsymbol{{\rm b}}}}
\newcommand{\xvec}{\ensuremath{\boldsymbol{{\rm x}}}}
\newcommand{\yvec}{\ensuremath{\boldsymbol{{\rm y}}}}
\newcommand{\zvec}{\ensuremath{\boldsymbol{{\rm z}}}}
\newcommand{\vvec}{\ensuremath{\boldsymbol{{\rm v}}}}
\newcommand{\jvec}{\ensuremath{\boldsymbol{{\rm j}}}}

\newcommand{\bxgp}{\bvec\times\gradperp}

\newcommand{\vve}{\ensuremath{\boldsymbol{{\rm v}}_{e}}}
\newcommand{\vvi}{\ensuremath{\boldsymbol{{\rm v}}_{i}}}
\newcommand{\vpe}{v_{\parallel e}}
\newcommand{\vpi}{v_{\parallel i}}
\newcommand{\vvE}{\ensuremath{\boldsymbol{{\rm v}}_{E}}}
\newcommand{\vvD}{\ensuremath{\boldsymbol{{\rm v}}_{D}}}

\newcommand{\nuei}{\nu_{ei}}
\newcommand{\nuii}{\nu_{ii}}
\newcommand{\nue}{\nu_{e}}
\newcommand{\nuen}{\nu_{en}}
\newcommand{\nuin}{\nu_{in}}
\newcommand{\kpe}{\kappa_{\parallel e}}

\newcommand{\rs}{\rho_{s}}
\newcommand{\ri}{\rho_{i}}
\newcommand{\wci}{\Omega_{i}}
\newcommand{\wcix}{\Omega_{ix}}
\newcommand{\wce}{\Omega_{e}}
\newcommand{\tomega}{\tilde\omega}
\newcommand{\Isat}{I_{\rm sat}}
\newcommand{\fmie}{\frac{m_i}{m_e}}
\newcommand{\fmei}{\frac{m_e}{m_i}}

% Often used dimensions
\newcommand{\cm}{\rm cm}
\newcommand{\mm}{\rm mm}
\newcommand{\cmn}{{\rm cm}^{-3}}
\newcommand{\mn}{{\rm m}^{-3}}
\newcommand{\eV}{\rm eV}
\newcommand{\G}{\rm G}
\newcommand{\T}{\rm T}

\begin{document}

{\bf Response to Referees \\
 \\
 \\} 
Brett Friedman \\
Updated: \today \\

\hrulefill

\section{Referee 1}

\emph{1) After advocating the use of non-modal structures rather than linear eigenmodes for the
entirety of the paper, the author use a time-scale derived from linear eigenmodes (see
page 4, paragraph 2). Furthermore, they use the frequency for the n=1 mode instead of
that for the n=0 mode(s), with no apparent justification (other than "nonlinear coupling”).
It would seem much more consistent to use the time-scale for the nonlinear instability
they talk about earlier (although this would mean perhaps a loss of predictability that they
are very keen on). To us, this appears to be the weakest link in the paper. Perhaps, if they
present this paper as a study of the turbulence properties instead, they could give
information on how the curves in Figure 3 change with the choice of the linear time-
scale. It would seem that this would be much more useful to the reader in assessing the
power of the non-modal analysis, rather than presenting just two choices, and claiming
predictability. } \\

Response: \\

\emph{2) What closure assumption is used to write equation (5)?} \\

Response: We do not really understand this question. We do not use any closure to write equation 5. It is an exact equation that can be derived from the model equations~\cite{friedman2012b,friedman2013}. \\

\emph{3) On page 2, last paragraph, there is a puzzling statement: "Perhaps one of the most
useful statistical properties that we can predict is the turbulent growth rate spectrum.” Do
they actually have a measurement of the turbulent growth rate spectrum? Since
predictability is one of their prime concerns, what measurable transport property does the
growth rate spectrum relate to?}

Response: We acknowledge that this was too cryptic a statement. In it's place, we have inserted the following, ``These properties can be calculated through mixing length formulas if the turbulent growth rate spectrum is known.'' Then, as before, we proceed to calculate the turbulent growth rate spectrum from the nonlinear simulation (shown in Figure 1b) and also calculate it from our non-modal linear procedure (labelled $\gamma_{TR}$, the results of which are shown in Figure 3. At the end of the paper, we use the $\gamma_{TR}$ spectrum through the mixing length formula $\gamma/k_\perp^2$ to calculate the turbulent saturation level. \\

\emph{4) We think a log scale should be used for Figure 2. On the scale used in the paper, it is
impossible to determine if the growth rate at short times is somewhat similar to $G_{max} (t)$
or quite different. If similar, the potential for formulating fully analytic models is more
promising.} \\

Response: We have changed Figure 2 to a log scale. The growth rate of the ensemble of curves is not close to the growth rate of $G_{max}(t)$, unfortunately, which is why we did not use it in the calculations of the transient growth rate spectrum. \\

\emph{5) Minor point: Figure 3 is confusing simply because the important curves are not
sufficiently well-delineated. For example, the turbulent growth rate spectrum should be
bold (or something equivalent) to distinguish it from the others.} \\

Response: The curves are different colors, so they should be distinguishable, but not in black and white. Thus, we have put symbols on them. \\

\section{Referee 2}

\emph{1) I find this manuscript somewhat confusing, especially when the authors try to cast it as a ‘non-modal analysis’. For non-modal stability analysis of a linear system,}

\beq
\diff{\mathbf{u}}{t} = \mathbf{A u} \nonumber
\eeq

\emph{I would expect the prediction of the short time behavior of some norm of $\mathbf{u}$ through the mathematical analysis of $\mathbf{A}$
by methods other than the standard eigenmode decomposition, e.g., by methods based on the $\epsilon$-pseudospectrum and/or numerical range.}

\emph{If I understand correctly, the method presented here starts by dropping the nonlinear terms in
the equations of motion of the reduced Braginskii 2-fluid model. The resulting linear system is
numerically simulated up to a time $\tau_{nl}$ at which the system is then ‘re-randomized’ — whatever
it means. A growth rate $\gamma_{TR}$ defined and computed using data between $t = 0$ and $t = \tau_{nl}$ then
constitutes the prediction by this method. So it seems that the ‘non-modal method’ described in
this manuscript is basically a numerical simulation of the linearized equations of motion. While
prediction of turbulent properties by numerical simulations can be useful, it probably has less
impact than a theoretical prediction (which is admittedly much more difficult).} \\

Response: We contend that our method is a form of 'non-modal analysis'. That is, we use only linear calculations to find growth rates without using normal modes (eigenvalues and eigenvectors). While most non-modal analysis in the literature seems to center around the calculation of the $epsilon$-pseudospectrum, much also includes the calculation of $||e^{\mathbf{A} t}||$ and $||e^{\mathbf{A} t} \mathbf{u}(0)||$ with various optimal $\mathbf{u}(0)$ vectors. \\

\emph{2) For the procedure described above, there are several things I do not fully understand. First of
all, if $\gamma_{TR}$ is only computed using data for $0 < t < \tau_{nl}$, how does the ‘re-randomization’ at $t = \tau_{nl}$
come into play? I am probably missing something here.} \\

Response: We agree that this was a confusing way to word our procedure. We therefore have replaced our previous wording with the following: ``we start with an ensemble of random initial conditions, evolve them linearly for a time $\tau_{\rm{nl}}$, and then take the time and ensemble average growth rate of these curves. 
This procedure mimicks what would happen if the turbulence evolved linearly for a time $\tau_{\rm{nl}}$, randomized, and then repeated.''
\\

\emph{3) Secondly, the various growth rates used here are defined as}

\beq
\gamma(m,n) = \frac{Q(m,n) + D(m,n)}{E(m,n)} \nonumber
\eeq
\emph{whereas the evolution of the energy for mode $(m, n)$ is given by}

\beq
\diff{E(m,n)}{t} = Q(m,n) + D(m,n) + \sum_{m',n'} T(m,m',n,n') \nonumber
\eeq
\emph{What is the reason of ignoring $T$ in the definition of $\gamma$? Similarly, what is the meaning/significance
of plotting $E(m, n, t)$ with nonlinearities turned off?  $\sum_{m',n'} T(m,m',n,n')$ is non-zero and contributes to 
$\diff{E(m, n)}{t}$ even though $\sum_{m,m',n,n'} T(m,m',n,n') = 0$. Would it be more informative
to compare $\gamma_{TR}$ to $\diff{E(m, n)}{t}$ instead of the $\gamma_e (m, n)$ defined in the manuscript (which ignores
contribution from $T$ )?} \\

Response: The reason why $T$ is missing from the definition of $\gamma$ is because $\diff{E(m,n)}{t} = Q(m,n) + D(m,n) + \sum_{m',n'} T(m,m',n,n')$ is virtually zero when the turbulence is in steady state (by definition of steady state). It may not be zero at a given time, but it is zero on average, and thus not suitable for use in defining the growth rate spectrum, which is a time-averaged property (see Figure 1b for the average and the spread in the growth rate). If we were to include $T$ in the definition, the turbulent growth rate curves in Figure 1b would be zero with some finite spread. So we can't use $T$ in the definition.

More informatively, the growth rate we use (We didn't invent it. It has been used for many years in numerous papers.) is a measure, by wavenumber, of the rate at which the fluctuations take energy from the equilibrium gradients (or return energy if the growth rate is negative). This comes purely from the linear terms. The nonlinear terms transfer energy between different fluctuation 'modes', but do not affect the overall energy budget.
So the picture is that certain modes or structures (those with positive $\gamma$) inject energy from the equilibrium gradients into the fluctuations through the linear terms. Those modes then non-linearly transfer energy to other modes in an overall conservative way (like a cascade). When energy is nonlinearly transferred into modes with negative growth rate, that energy is dissipated. In the case of normal linear operators, the unstable linear eigenvectors inject energy, transfer that energy to stable linear eigenvectors, which then dissipate the energy. For normal operators, one would expect this growth rate to be pretty similar to the fastest growing eigenmode growth rate, though it could be smaller. But it certainly cannot be larger. We have added and modified quite a bit of text and added a reference~\cite{terry2006b}, which uses this growth rate so that it doesn't appear that we have invented it.

Finally, the reason we plot $E(m,n,t)$ with the nonlinearities turned off is to show that the linear terms alone cause (transient) growth of the fluctuations despite the fact that the $n=0$ mode numbers contain no unstable eigenmodes. It illustrates that the turbulent structures inject energy into the fluctuations in the short term. \\

\emph{4) I also feel that the clarity of the present manuscript suffers because of the page limitation of
PRL, a lot of the symbols are not defined (e.g. $v_E$, $N_r$ ) making it difficult for the general audience.
Together with the concerns raised above, I am afraid I cannot recommend the publication of the
present manuscript in the PRL.} \\

Response: \\

\bibliographystyle{unsrt}
\bibliography{refs}

%
%

\end{document}

### 2014-02-23 Transient_growth_prl.tex
Feb 22

@@
 \usepackage{graphicx}
+%\usepackage{amssymb}
 %\setlength{\parindent}{0in}
@@
 
-\title{A Linear Technique to Predict Non-Normal Turbulence Applied to a Magnetized Plasma}
+\title{A Linear Technique to Understand Non-Normal Turbulence Applied to a Magnetized Plasma}
 
@@
 [W. Gekelman \emph{et al.}, Rev. Sci. Inst. {\bf 62}, 2875 (1991)], where non-modal effects dominate the turbulence.
-We compare the resulting growth rate spectrum to that obtained from a nonlinear simulation, showing good quantitative agreement.
+We compare the resulting growth rate spectra to that obtained from a nonlinear simulation, showing good qualitative agreement.
 \end{abstract}
@@
 Such behavior is obscured by traditional normal mode analysis, which only effectively describes the long time asymptotic behavior of fluctuations under the 
-action of the linear operator. Transient growth events, which can dominate nonlinear evolution, require initial-value (non-modal) calculations rather than normal mode calculations.
+action of the linear operator. Transient growth behavior, which can dominate turbulent evolution, can be discovered only through non-modal calculations.
 
@@
 than on properties of fully-developed turbulence.
-This paper takes up the task of developing a quantitative approach to predicting turbulent properties using only non-modal linear calculations. 
+This paper takes up the task of developing an approach to understand turbulent properties using only non-modal linear calculations with the goal of 
+ultimately making quantitative predictions. 
 Our approach is to evolve the linear solution from an ensemble of random initial conditions 
@@
 nonlinear decorrelation time of the turbulent system, but in order to enable predictive capability, we employ critical balance
-arguments to use a characteristic \emph{linear} time.
-This procedure produces an effective growth rate spectrum that can be used to predict turbulent properties such as saturation levels.
+arguments to use a characteristic \emph{linear} time. Since several linear time scales exist, we test them all and compare the results to determine which works best.
+The procedure ultimately produces an effective growth rate spectrum that can be used to predict turbulent properties such as saturation levels through mixing length arguments.
 While the concepts behind this technique are general enough to be applied to many nonlinear dynamical systems, the details vary for each case, 
-so we restrict our treatment to one particular turbulence model. For this model, the technique reproduces the turbulent growth rate spectrum of the direct nonlinear simulation fairly well, especially
-in comparison to the linear eigenmode spectrum.
+so we restrict our treatment to one particular turbulence model. For this model, the technique qualitatively reproduces the turbulent growth rate spectrum of the direct 
+nonlinear simulation fairly well, especially in comparison to the linear eigenmode spectrum.
 
@@
 produced by the Large Plasma Device (LAPD)~\cite{gekelman1991}. 
-Due to LAPD's large size and high collisionality, a fluid model is
-appropriate to describe the turbulence~\cite{Popovich2010a,Popovich2010b,Umansky2011}. We use a reduced Braginskii 2-fluid model:
+We use a reduced Braginskii 2-fluid model~\cite{Popovich2010a,Popovich2010b,Umansky2011,friedman2012b,friedman2013}:
 
@@
 where $N$ is the density, $\vpe$ the parallel electron velocity, $\varpi \equiv \gradperp \cdot (N_0 \gradperp \phi)$ the potential vorticity,
-and $T_e$ the electron temperature. All lengths are
-normalized to the ion sound gyroradius $\rho_s$, times to the ion cyclotron time $\omega_{ci}^{-1}$, velocities to the sound speed $c_s$, densities to the equilibrium peak density, and electron
-temperatures and potentials to the equilibrium peak electron temperature. The profiles $N_0$ and $T_{e0}$ and other parameters are taken from experimental 
-measurements~\cite{schaffner2012,friedman2012b,friedman2013}. 
-%$\phi_0$  is absent in the equations because we model one particular experiment in which the equilibrium radial electric field was reduced to near zero value using a limiter biasing technique~\cite{schaffner2012}.
+$T_e$ the electron temperature, ${\mathbf v_E}$ the ${\mathbf E} \times {\mathbf B}$ velocity, and $S_N$ and $S_T$ density and temperature sources. All lengths are
+normalized to the ion sound gyroradius $\rho_s$, times to the ion cyclotron time $\omega_{ci}^{-1}$, velocities to the sound speed $c_s$, 
+densities to the equilibrium peak density, and electron temperatures and potentials to the equilibrium peak electron temperature. 
+The profiles $N_0$ and $T_{e0}$ and other parameters are taken from experimental measurements~\cite{schaffner2012,friedman2012b,friedman2013}.
 
-The equations are global and partially linearized. The only nonlinearities retained are the advective nonlinearities, which are written with Poisson brackets. 
+The equations are global and we retain only advective nonlinearities, which are written with Poisson brackets. 
 We add artificial diffusion and viscosity terms with small numerical coefficients ($10^{-3}$)
@@
 boundary conditions. Further details of the model, including validation studies, may be found in the references~\cite{Popovich2010a,Popovich2010b,Umansky2011,friedman2012b,friedman2013},
-though we mention here that the model does very well in reproducing
-the statistical properties of the experimentally-observed turbulence.
+though we mention here that the model does very well in reproducing the statistical properties of the experimentally-observed turbulence.
 
@@
 it is dominated by a nonlinear instability process despite being linearly unstable to drift waves~\cite{friedman2012b,friedman2013}.
-This nonlinear instability process, which was discovered by Drake et al.~\cite{drake1995} in a similar plasma model, works as follows: 
+The nonlinear instability, which was discovered by Drake et al.~\cite{drake1995}, works as follows: 
 magnetic-field-aligned ($k_\para=0$) convective filaments transport density across the equilibrium density gradient, setting up $k_\para=0$ density filaments. 
@@
 \label{dEdt_def}
-\diff{E(m,n)}{t} = Q(m,n) + D(m,n) + \sum_{m',n'} T(m,m',n,n')
+\diff{E(m,n)}{t} = \diff{E_{l}(m,n)}{t} + \diff{E_{nl}(m,n)}{t}
 \eeq
 where $m$ and $n$ represent the azimuthal and axial Fourier mode numbers. 
-%Since Eqs.~\ref{ni_eq}-\ref{te_eq} are not explicitly dependent on $\theta$ and $z$, the energy equation is separable for every $m$ and $n$. 
-$Q$ represents the linear energy exchange with the equilibrium (convective transport), $D$ the dissipation from irreversible processes
-like collisions, and $T$ the nonlinear triad wave transfer. Since $\int f \{g,f\} dV = 0$, or equivalently $\sum_{m,m',n,n'} T(m,m',n,n')=0$, 
-the Poisson bracketed advective nonlinearities in the model are energetically conservative and therefore do no inject or dissipate energy on the whole.
-Consequently, the fluctuation energy originates completely from $Q$, the linear convective transport mechanism.
+$\diff{E_{l}(m,n)}{t}$ comes from the linear terms in Eqs.~\ref{ni_eq}-\ref{te_eq}, $\diff{E_{nl}(m,n)}{t}$ comes from the nonlinear terms. 
+$\diff{E_{l}(m,n)}{t}$ describes the injection (or dissipation) of energy into the fluctuations from the free energy in the equilibrium gradients.
+$\diff{E_{nl}(m,n)}{t}$ describes the energy exchange between fluctuations with different $m,n$ and it is conservative: $\sum_{m,n} \diff{E_{nl}(m,n)}{t} = 0$.
+Moreover, in quasi-steady state turbulence, the rate of energy injection (or dissipation) into the fluctuations at each $m,n$ by the linear terms must be balanced by
+the rate of energy removal (or deposition) from the nonlinear terms:
+
+\beqar
+\label{steady_state}
+\gamma(m,n) \equiv  \displaystyle\lim_{T \to \infty} \frac{1}{T} \int_0^T \frac{dE(m,n)/dt}{2 E(m,n)} dt \nonumber \\
+= \displaystyle\lim_{T \to \infty} \frac{{\rm{Log}}\left[ E(T)/E(0) \right]}{T} = 0 \quad {\rm{in \ steady \ state}}.
+\eeqar
 
@@
 b) Linear and turbulent growth rate spectra for $n=0$ (solid lines) and $n=1$ (dashed lines) Fourier components. The linear growth rates are those of the least stable eigenmodes,
-while the turbulent growth rates represent $\pdiff{E}{t}|_{\rm{lin}}/2 E$ from the nonlinear simulation. The shaded region marks the $1 \sigma$ spread in the turbulent spectrum,
+while the turbulent growth rates represent $\pdiff{E_l}{t}/2 E$ from the nonlinear simulation. The shaded region marks the $1 \sigma$ spread in the turbulent spectrum,
 obtained from the distribution of growth rates in the nonlinear simulation.}
@@
 
+From, Eqs.~\ref{dEdt_def} and~\ref{steady_state}, it follows that $ \gamma(m,n) = \gamma_l(m,n) + \gamma_{nl}(m,n) = 0$.
+The rate of energy injection $\gamma_l(m,n)$ from the equilibrium gradient into the fluctuations is a quantity of great interest~\cite{friedman2012b,terry2006b}. 
+When positive for some wavenumber, turbulence can be sustained. Additionally, it may be used in the mixing length formula $\gamma/k_\perp^2$ to predict the turbulent saturation level, 
+and it is proportional to the transport rate~\cite{terry2006b}. 
+$\gamma_l(m,n)$ may be calculated from the spatial structures of of the plasma state variables, so it has meaning in turbulent plasma states. We plot
+$\gamma_l(m,n)$ in Fig.~\ref{n0_n1_growth} a) for both the linear and steady state turbulent stages of our nonlinear simulation. The linear stage, which occurs when fluctuations
+are small and exponentially growing has a time-independent $\gamma_l(m,n)$ equal to $\gamma_s(m,n)$ -- the ``spectral'' growth rate of the fastest growing eigenmode at each $m,n$.
+During the turbulent stage, $\gamma_l(m,n)$ is time-dependent (indicated by the $1 \sigma$ spread), and generally much different than $\gamma_s(m,n)$. Significantly,
+the turbulent $\gamma_l$ is positive at $n=0$ and low $m$ despite the fact that all linear eigenmodes have $\gamma_s < 0$ for $n=0$. This is a manifestation
+of non-normality, for in normal systems, $\gamma_l(m,n) \le \gamma_s(m,n)$. Physically, it is the manifestation of the nonlinear
+instability, specifically the linear part of the mechanism in which convective filaments drive density filaments from the equilibrium density gradient.
+
 This convective transport of density filaments is akin to the paradigmatic ``lift-up'' mechanism in hydrodynamic shear flows whereby streamwise vortices drive streamwise 
@@
 Both are transient growth processes. We see this in our simulations by following the evolution of the energy of several $m,n$ modes after turning off the nonlinearities in an already
-turbulent simulation. A few representative modes are shown in Fig.~\ref{n0_n1_growth} a). The linear transient growth of the filamentary $n=0$ structures is evident as their modes 
-grow transiently before decaying exponentially at the rate of their least stable eigenmode. Such behavior is rather unituitive because all $n=0$ linear eigenmodes are stable, 
-and the normal mode paradigm has conditioned us to believe that a linear superposition of such eigenmodes should decay. But as the hydrodynamics community discovered, linear
-superpositions of nonorthogonal vectors may behave unintuitively.
-Notice finally that the $n=1,m=20$ mode decays transiently before growing exponentially with growth rate of the most unstable eigenmode.
-%This isn't necessarily due to eigenvector nonorthogonality, but is nevertheless unintuitive.
+turbulent simulation. A few representative modes are shown in Fig.~\ref{n0_n1_growth} b). The linear transient growth of the filamentary $n=0$ structures is evident as their modes 
+grow transiently before decaying exponentially at the rate of their least stable eigenmode. Such behavior is indicative of non-modal behavior. It must be since 
+all $n=0$ linear eigenmodes are stable. Notice also that the $n=1,m=20$ mode decays transiently before growing exponentially with growth rate of the most unstable eigenmode. This transient
+is another non-modal result.
 
-Since transient growth is a purely linear phenomenon, it has been a goal of researchers to predict the onset of subcritical turbulence using
-only linear, non-modal calculations.
-It is our goal, however, to use linear calculations to go beyond predicting onset and predict fluctuation amplitudes and transport associated with saturated turbulence.
-Perhaps one of the most useful statistical properties that we can predict is the turbulent growth rate spectrum.
-To obtain a growth rate from the nonlinear simulation, we begin with the expression $\diff{E(m,n)}{t}\big|_{\rm{lin}} = Q(m,n) + D(m,n)$, 
-which accounts for the total net energy injection into the fluctuations at each $m,n$. 
-An effective growth rate $\gamma_e(m,n) = \diff{E(m,n)}{t}\big|_{\rm{lin}}/ 2 E(m,n)$ can then be defined. 
-This growth rate is calculated from the spatial structure of the fluctuations, so it is well-defined for both turbulent fluctuations and eigenmode fluctuations. If calculated
-from a linear simulation that is run for a sufficiently long time, $\gamma_e(m,n)$ is the same as the spectral eigenmode growth rate $\gamma_s(m,n)$ of the least stable linear eigenmode.
-For turbulence, $\gamma_e(m,n)$ is generally not equal to an eigenmode growth rate, especially when the system is non-normal.
-We plot this growth rate in Fig.~\ref{n0_n1_growth} b) for $n=0$ and $n=1$ components. The turbulent growth rate curves are calculated from this formula, 
-averaged over a time of about $3 \times 10^{4} \ \omega_{ci}^{-1}$ during the saturated turbulent phase of the nonlinear simulation.
-Also, we show the $1 \sigma$ contour, which indicates how greatly $\gamma_e$ varies over time.
-Fig.~\ref{n0_n1_growth} b) indicates that while the linear eigenmodes are unstable for $n=1$ and stable for $n=0$,
-the turbulent growth rates are positive for $n=0, m<50$ (subcritical-like behavior), but negative elsewhere. 
-Fig.~\ref{n0_n1_growth} a) hinted at these characteristics, but this figure represents quantitative confirmation.
-
-Our goal now is to reproduce the turbulent growth rate curves in Fig.~\ref{n0_n1_growth} b) without performing nonlinear simulations. 
-To accomplish this, we use the anzatz that nonlinearities randomize the turbulent spatial structure at each wavenumber on a time scale of 
+Since transient growth is a purely linear phenomenon, it has been a goal of researchers to understand and predict the onset of subcritical turbulence using only linear, non-modal calculations. 
+In our system, the turbulence is not subcritical in the traditional sense, but it has a subcritical component in the $n=0$ modes. They have no unstable eigenmodes, but still inject energy into the
+turbulence -- $\gamma_l(m<50,n=0) > 0$, yet $\gamma_s(m,n=0) < 0$. It is our goal, then, to use linear non-modal calculations to understand this behavior and to move toward predictive capability
+of $\gamma_l$.
+To accomplish this, we make the anzatz that nonlinearities randomize the turbulent spatial structure at each wavenumber on a time scale of 
 one eddy decorrelation time, while the linearities evolve the spatial structures deterministically. From this we can calculate a ``Transient Growth Rate'' spectrum,
@@
 and $\mathbf{B}_{m,n}$ and $\mathbf{C}_{m,n}$ are coefficient matrices that include the equilibrium information and finite difference coefficients. The first term on the RHS represents the linearities
-and the second term the nonlinearities. Note that for each $m,n$, there exist $4 \times N_r$ linearly independent, but nonorthogonal eigenvectors. Hence forth, we drop the $m,n$ subscripts.
+and the second term the nonlinearities. Note that for each $m,n$, there exist $4 N_r$ linearly independent, but nonorthogonal eigenvectors, where $N_r$ is the number of radial grid points. 
+Hence forth, we drop the $m,n$ subscripts.
 
@@
 inner product is possible, a physically relevant one such as an energy inner product is generally preferred~\cite{camargo1998,schmid2007,camporeale2010}. 
-Recall that the inner product of two vectors may be written $\left< \mathbf{x},\mathbf{y} \right> = \mathbf{y}^{*} \mathbf{M} \mathbf{x}$,
-where $^*$ stands for the conjugate transpose, and $\mathbf{M}$ is a positive-definite matrix. We choose $\mathbf{M}$ so that 
-$||\mathbf{v}|| = \sqrt{\left< \mathbf{v},\mathbf{v} \right>} = \sqrt{E}$, where $E$ is the energy. Furthermore, it is convenient in
-computations to use the $L_2$-norm, $||\mathbf{u}||_2 = \sqrt{\sum_i |u_i|^2}$. This can be accomplished through the change of variables $\mathbf{u} = \mathbf{M}^{\frac{1}{2}} \mathbf{v}$.
+Recall that the inner product of two vectors may be written $\left< \mathbf{x},\mathbf{y} \right> = \mathbf{y}^{\dagger} \mathbf{M} \mathbf{x}$,
+We choose $\mathbf{M}$ so that $||\mathbf{v}||^2 = \left< \mathbf{v},\mathbf{v} \right> = E$. Furthermore, it is convenient in
+computations to use the $L_2$-norm, $||\mathbf{u}||_2^2 = \sum_i |u_i|^2$. This can be accomplished through the change of variables $\mathbf{u} = \mathbf{M}^{\frac{1}{2}} \mathbf{v}$.
 Then the linear portion of Eq.~\ref{mat_eq} becomes
@@
 \eeq
-%The eigenvalues of $\mathbf{A}$ are the same as those of $\mathbf{B}^{-1} \mathbf{C}$ because $\mathbf{M}^{\frac{1}{2}} \mathbf{B}^{-1} \mathbf{C} \mathbf{M}^{-\frac{1}{2}}$ is a similarity transformation.
 The solution of Eq.~\ref{lin_eq_A} is $\mathbf{u}(t) = e^{\mathbf{A} t} \mathbf{u}(0)$, which
 depends on the initial condition $\mathbf{u}(0)$ in addition to the spectral properties of $\mathbf{A}$. For purposes of turbulent growth rate prediction, we are interested in
-the behavior of $||\mathbf{u}(t)||$. Therefore, we introduce the growth ratio $G(t)$, which measures the amplification or reduction in the square root of the energy from an initial state:
+the behavior of $G(t) = E(t)/E(0) = ||\mathbf{u}(t)||^2/||\mathbf{u}(0)||^2$ and its rate of growth.
 
-\beq
-\label{g_def}
-G(t) = \frac{||\mathbf{u}(t)||}{||\mathbf{u}(0)||} = \frac{||e^{\mathbf{A} t} \mathbf{u}(0)||}{||\mathbf{u}(0)||}.
-\eeq
-This ratio is bounded from above by $G_{\rm{max}}(t) = ||e^{\mathbf{A} t}||$. $G_{\rm{max}}(t) = e^{\gamma_s t}$ if and only if $\mathbf{A}$ is normal~\cite{schmid2007}. 
-Otherwise, $G_{\rm{max}}(t) > e^{\gamma_s t}$.
+%\label{g_def}
+%G(t) = \frac{||\mathbf{u}(t)||}{||\mathbf{u}(0)||} = \frac{||e^{\mathbf{A} t} \mathbf{u}(0)||}{||\mathbf{u}(0)||}.
+%\eeq
+%This ratio is bounded from above by $G_{\rm{max}}(t) = ||e^{\mathbf{A} t}||$. $G_{\rm{max}}(t) = e^{\gamma_s t}$ if and only if $\mathbf{A}$ is normal~\cite{schmid2007}. 
+%Otherwise, $G_{\rm{max}}(t) > e^{\gamma_s t}$.
 
 \begin{figure}
-\centerline{\includegraphics[width=0.55\textwidth]{m20n0_time_evs}}
-\caption{a) The maximum growth ratio curve $G_{\rm{max}}(t)$ (dashed line) for $n=0,m=20$, and an ensemble of growth ratio curves (solid gray lines, with the solid black line the ensemble average)
-that start with a random initial condition $u(0)$ and evolve under the linear operator. b) The ensemble curves with $t=\omega_s^{-1}$ indicated, at which we cut off the evolution.}
-\label{m20n0_time_evs}
+\centerline{\includegraphics[width=0.55\textwidth]{m20n0_E_ev}}
+\caption{An ensemble of growth ratio curves (solid gray lines, with the solid black line the ensemble average)
+that start with a random initial condition $u(0)$ and evolve under the linear operator.}
+\label{m20n0_E_ev}
 \end{figure}
@@
 It is common practice in normal mode analysis to look for the least stable eigenmode. 
-For the non-modal case, it is common to study the properties of $G_{\rm{max}}(t)$ because if $G_{\rm{max}}(t) > 1$ at any time, fluctuations may be amplified, leading to subcritical turbulence.
+For the non-modal case, it is common to study the properties of $G_{\rm{max}}(t) = ||e^{\mathbf{A} t}||$ because if this is greater at any time, fluctuations may be amplified, 
+leading to subcritical turbulence~\cite{trefethen2005,schmid2007}.
 However, it can be misleading to study only $G_{\rm{max}}(t)$ when predicting specific properties of turbulence because
@@
 
-We illustrate $G_{\rm{max}}(t)$ along with $G(t)$ at $m=20, n=0$ for several random initial conditions in Fig.~\ref{m20n0_time_evs} a). Despite the fact that all eigenmodes at this
-wavenumber are stable, there is the possibility for transient amplification by a factor of $20$. Furthermore, a sample of randomly initialized vectors are all transiently amplified, but none display
-optimal growth. It is extremely unlikely to choose a random initial vector that either optimally amplifies the energy at any time or one that monotonically decays, which would be the case
-if the initial vector were one of the linear eigenvectors.
-
-The key to quantifying non-modal analysis and making it predictive is to successfully model the effect that the nonlinearities have on the transient linear processes. 
+We contend that the key to understanding and prediction turbulent properties through non-modal analysis is to successfully model the effect that the nonlinearities have on the transient linear processes.
 To this effect, we note that the advective nonlinearity in Eq.~\ref{mat_eq} has the form of the state vector divided by a time $\tau_{\rm{nl}} \sim (v_E k_\perp)^{-1}$. This nonlinear
@@
 as a randomizing force that acts on this characteristic nonlinear time scale.
-Specifically, we start with a random initial condition, evolve it linearly for a time $\tau_{\rm{nl}}$, and then re-randomized the system.
-To estimate $\tau_{\rm{nl}}$, we invoke the conjecture of \emph{critical balance}, which posits that the nonlinear time scale equals the linear time scale at all spatial scales~\cite{schekochihin2012}.
-For the linear time scale, we take the inverse of the linear frequency (at each $m,n$) -- $\tau_{\rm{nl}} = \omega_s^{-1}$.
-However, $\omega_s = 0$ for $n=0$ linear eigenmodes, so we are forced to use $\omega_s$ for the fastest growing $n=1$ eigenmode at each $m$ to get a meaningful time scale. This is not wholly
-unjustified, however, due to the inherent nonlinear coupling of different modes, evident in Eq.~\ref{mat_eq}.
-In Fig.~\ref{m20n0_time_evs} b), this time scale is indicated. Note that we cut off the transient growth well before optimal amplification occurs, at least for this wavenumber.
+Specifically, we start with an ensemble of random initial conditions, evolve them linearly for a time $\tau_{\rm{nl}}$, and then take the time and ensemble average growth rate of these curves.
+This procedure mimicks what would happen if the turbulence evolved linearly for a time $\tau_{\rm{nl}}$, randomized, and then repeated. Now, this procedure does not require actual linear
+simulations from an ensemble of random initial conditions. Recall that the time evolution of the energy from an initial condition is
 
-We define the ``Transient Growth Rate'' $\gamma_{\rm{TR}}$ as the average from $t=0$ to $t=\omega_s^{-1}$ of the instantaneous growth rate of the ensemble averaged growth curve.
-In Fig.~\ref{spec_nl_transient_gamma}, we compare the $\gamma_{\rm{TR}}$ spectrum (labeled ``Trans. $\omega_s^{-1}$'') to the turbulent spectrum.
-Although the $\gamma_{\rm{TR}}$ spectrum for $n=0$ is downshifted in $m$ number relative to the spectrum derived from the nonlinear simulation, the two curves in Fig.~\ref{spec_nl_transient_gamma} a)
-are fairly similar in shape and magnitude.
-The $n=1$ curves in Fig.~\ref{spec_nl_transient_gamma} b), on the other hand, are not very similar, but $\gamma_{\rm{TR}}$ correctly predicts energy damping in the region where
-eigenmode growth is predicted.
-The difference between the $n=1$ transient and turbulent growth rate spectra is due to nonlinear effects not captured by the simple transient calculation procedure. In particular, the nonlinear
-simulation has strong energy transfer from $n=0$ to $n=1$ which dissipates by electron-ion collisions~\cite{friedman2012b}.
+\beq
+\label{E_t_from_u0}
+E(t) = ||e^{\mathbf{A} t} \mathbf{u}(0)||^2 = e^{\mathbf{A} t} \mathbf{u}(0) \mathbf{u}^{\dagger}(0) e^{\mathbf{A}^{\dagger}t}.
+\eeq
+If the $\mathbf{u}(0)$ in the ensemble are random with uncorrelated components and normalized to unity, it follows that~\cite{camargo1998}
 
-We include one more growth rate curve in Fig.~\ref{spec_nl_transient_gamma}, labelled ``Trans $\omega_s^{-1}$ cascade,'' to address the apparent downshift (in $m$) of
-the ``Trans $\omega_s^{-1}$'' curve. As seen in Eq.~\ref{mat_eq}, $v_{m,n}$ is affected by every $\mathbf{v}_{E,m-m',n-n'} k^{'}_\perp$, not by only $\mathbf{v}_{E,m,n} k_\perp$ as we
-assumed when using $\omega_s^{-1}$. Thus, we pose a simple forward cascade (eddy mitosis) model for the nonlinearity.
-We use $\omega_s^{-1}$ at $m/2$ rather than $m$ for the randomizing time scale at each $m$ in deriving the ``Trans $\omega_s^{-1}$ cascade'' curves. 
-The $n=0$ curve matches well with the $n=0$ Turbulent Growth Rate curve at $m>20$, but poorly for $m<20$ as expected since the cascade applies only at high $m$. 
-We cannot determine \emph{a priori} the minimum $m$ for which the cascade should apply, so this method is not predictive.
+\beq
+\label{E_t_ensemble_avg}
+\left< E(t)/E(0) \right>_{{\rm{ens}}} = \frac{1}{4 N_r} {\rm{trace}} e^{\mathbf{A} t} e^{\mathbf{A}^{\dagger}t}.
+\eeq
+We show the validity of this statistical averaging by plotting $1000$ ensemble curves in Fig.~\ref{m20n0_E_ev} along with their expected average from Eq.~\ref{E_t_ensemble_avg} and their actual average,
+which line up perfectly.
+One point to note is that in global equation sets like ours, where we discritize and use finite differences in the radial direction rather than a Fourier decomposition, randomizing $\mathbf{u}(0)$
+amounts to setting the initial $k_r$ spectrum to a step function that goes to zero at the Nyquist wavenumber.
+This means that $\left< E(t)/E(0) \right>_{{\rm{ens}}}$ may depend on $N_r$, so we must be careful in
+choosing $N_r$ for this analysis. We choose $N_r$ such that the grid spacing equals $\rho_s$, the general Nyquist wavenumber of drift wave simulations~\cite{scott1992}.
+
+Mathematically, our procedure is to calculate $\gamma_l$ by the following formula:
+
+\beq
+\label{gamma_l_calc}
+\gamma_l = \frac{1}{\tau_{nl}} \int_0^{\tau_{nl}} \pdiff{E(t)}{t}/ 2 E(t) dt = \frac{1}{2 \tau_{nl}} {\rm{Log}} \left[ \frac{E(\tau_{nl})}{E(0)}\right]
+\eeq
+where $E(t)$ is the ensemble averaged energy calculated from Eq.~\ref{E_t_ensemble_avg}, and $E(0) = 1$ by our normalization.
+In order to move toward predictive capability, we must estimate $\tau_{\rm{nl}}$ with only knowledge of linear (modal or non-modal) information. 
+We thus invoke the conjecture of \emph{critical balance}, which posits that the nonlinear time scale equals the linear time scale at all spatial scales~\cite{schekochihin2012}, which is clear from 
+Eqs.~\ref{dEdt_def} and~\ref{steady_state}, which lead to $\gamma_l = - \gamma_{nl}$. Now there are several linear time scales that we may choose to test. We label these times in Fig.~\ref{m20n0_E_ev}
+for the case of $n=0, m=20$. The first linear time is the linear eigenmode frequency, labeled
+$\tau_{nl} = \omega_s^{-1}$. However, $\omega_s = 0$ for $n=0$ linear eigenmodes, so we are forced to use $\omega_s$ for the fastest growing $n=1$ eigenmode at each $m$ to get a meaningful time scale.
+Second is the parallel free-streaming time of the electrons $\tau_{nl} = t_{\rm{st}} = L_\para/v_{t e}$ often cited in critical balance arguments~\cite{barnes}. Third is the time before modal effects
+take over, which can be approximated as the time when $E(t)$ turns over as seen in Fig.~\ref{n0_n1_growth}, where $E(t)$ can be either a maximum or minimum, labeled $\tau_{nl} = E_{\rm{max}}$. 
+Fourth, we take the $\tau_{nl} = 1/\gamma_{nl} = - 1/\gamma_l$. Inserting this into Eq.~\ref{gamma_l_calc} gives
+
+\beq
+\label{fourth_gamma_cond}
+ {\rm{Log}} \left[ E(1/\gamma_l)\right] = \pm 2 \ \rightarrow E(1/\gamma_l) = e^{\pm 2}.
+\eeq
+In other words, $\gamma_l$ is found by the inverse of the time when $E(t)$ grows to the value of $e^2$ or decays to the value of $e^{-2}$, labeled $\tau_{nl} = e^2$. We also indicate two times
+labeled $\gamma_{\rm{sim}}$, which when inserted into Eq.~\ref{gamma_l_calc} give the ``correct'' $\gamma_l$ ($\gamma_l(m=20,n=0)$ calculated directly from the nonlinear simulation).
+
+In Fig.~\ref{gamma_comparisons}, we compare the $\gamma_{l}$ spectrum from the nonlinear simulation to $\gamma_l$ from the non-modal procedure using the four different linear time scales. We also show
+$\gamma_s$ for reference.
 
 \begin{figure}
-\centerline{\includegraphics[width=0.5\textwidth]{spec_nl_transient_gamma}}
+\centerline{\includegraphics[width=0.5\textwidth]{gamma_comparisons}}
 \caption{The linear, turbulent, and transient growth rate spectra. The transient spectra are calculated by the average growth rate over a time $\omega_s^{-1}(m,n=1)$ for the ensemble average
 growth ratio curves with the cascade curve using $\omega_s^{-1}$ at $m/2$ rather than $m$.}
-\label{spec_nl_transient_gamma}
+\label{gamma_comparisons}
 \end{figure}
 
-Nevertheless, our non-modal method, which requires only linear calculations, reproduces energetic growth properties of the turbulence not captured by eigenmode analysis.
-Furthermore, when linear systems are highly non-normal, transient growth rates should be used in the kinds of predictions that are currently made with eigenmode growth rates. 
-For instance, quasilinear theory uses the most unstable eigenmode growth rate $\gamma_{s,max}$ to predict the turbulent saturation levels using a mixing-length argument 
-(saturated amplitude $\sim \gamma/k_\perp^2$). 
-Using this formula with $\gamma_{s,\rm{max}} = 0.002$ and $m=60$, one predicts a turbulent saturation level of $\sim 0.1 \%$ which is wholly inconsistent with experiment and simulation. 
-On the other hand, using this formula with the transient values $\gamma_{\rm{TR, max}} = 0.003, \ m=10$, we predict the saturation level of $\sim 10 \%$,
-which is consistent with the nonlinear simulation and experimental measurements.
-
-In summary, we present a procedure for predicting the turbulent growth rate spectrum using non-modal linear calculations. 
+In summary, we present a procedure for calculating the turbulent growth rate spectrum using non-modal linear calculations. 
 In the case of an LAPD experiment, this procedure captures the behavior of a nonlinear instability that dominates the dynamics of the turbulence.  In general,
-non-modal initial-value analysis is difficult to quantify and make predictive, but using some simple nonlinear modelling, we have shown
-that it is possible. Future studies will attempt to improve on the nonlinear modelling and expand the use of this technique to other non-normal turbulence models.
+non-modal analysis is difficult to quantify and make predictive, but using some simple nonlinear modelling, we have shown
+that it may be possible. Future studies will attempt to test this procedure on other non-normal turbulence models and to see if it can predict critical parameters for subcritical turbulent onset.
 

### 2014-02-24 Referee_Response1/referee_response.tex
Feb 23

@@
 
+\section{Overview of Changes}
+
+Before responding to specific comments, questions, and requests from the referees, we would like to note that we have significantly changed parts of the manuscript. First, we have
+changed the way we calculate the $|e^{A t}u(0)|$ curves where we used an ensemble of initial conditions. Before we made many initial conditions, evolved them by this formula and then
+calculated the ensemble average. This way seemed a bit like performing numerical simulations. Now we have an analytic method of calculation which doesn't involve setting up any initial
+conditions, so it is not at all like simulation (see Eq. 10) and is much more elegant.
+
+Second, recall that our method involved choosing a linear time scale as a proxy for the eddy turnover time, for which we chose the inverse of a linear eigenvalue frequency. 
+As the first referee pointed out, we didn't properly justify this choice. Since we have no theoretical justification, 
+we have now included a few more linear time scales and studied the result for each of these. Furthermore, we no longer claim that our method
+is predictive because we have no way to know \emph{a priori} which linear time scale will work best. However, all of the linear time scales do work well to a certain degree, so
+we do obtain a degree of predictivity.
+
+Otherwise, we have made various minor changes and changed two of the figures to make sense with the changes mentioned above. 
+Overall the manuscript is significantly different than before, and we feel it is much stronger.
+
 \section{Referee 1}
@@
 
-Response: \\
+Response: This is a valid criticism. We thought a lot about this and tried a few different linear time scales. The one we chose worked best, but it is difficult to justify it otherwise.
+We agree that this is the weakest link of the paper. We also think that the suggestion of presenting several choices of linear time scales is a good one, so this is what we've done.
+We've presented four different choices of linear time scale and compared the results to each other. We have also softened our language on predictive capability and presented our results
+as a way of understanding the turbulent growth rate through linear non-modal calculations with the ultimate goal of achieving prediction. \\
 
@@
 It is an exact equation that can be derived from the model equations~\cite{friedman2012b,friedman2013}.
-\\
+We have also now written it differently to make it simpler and more obvious. \\
 
@@
 predictability is one of their prime concerns, what measurable transport property does the
-growth rate spectrum relate to?}
+growth rate spectrum relate to?} \\
 
 Response: We acknowledge that this was too cryptic a statement.
-In it's place, we have inserted the following, ``These properties can be calculated through mixing length formulas if the turbulent growth rate spectrum is known.'' Then, as before, we proceed to calculate the turbulent growth rate spectrum from the nonlinear simulation (shown in Figure 1b) and also calculate it from our non-modal linear procedure (labelled $\gamma_{TR}$, the results of which are shown in Figure 3.
-At the end of the paper, we use the $\gamma_{TR}$ spectrum through the mixing length formula $\gamma/k_\perp^2$ to calculate the turbulent saturation level.
+In it's place, we have inserted the following, ``These properties can be calculated through mixing length formulas if the turbulent growth rate spectrum is known.'' Then, as before, we proceed to calculate the turbulent growth rate spectrum from the nonlinear simulation (shown in Figure 1a) and also calculate it from our non-modal linear procedure.
 \\
@@
 
-Response: We have changed Figure 2 to a log scale.
-The growth rate of the ensemble of curves is not close to the growth rate of $G_{max}(t)$, unfortunately, which is why we did not use it in the calculations of the transient growth rate spectrum.
+Response: We have changed Figure 2 so that we do not show $G_{max}(t)$ any longer.
+The growth rate of the ensemble of curves is not close to the growth rate of $G_{max}(t)$.
+$G_{max}(t)$  grows so quickly,
+and it isn't clear how one would use it in a procedure like ours because the points along it require linear evolution from very specific initial conditions, and every point comes from a different initial condition.
 \\
@@
 
-Response: The curves are different colors, so they should be distinguishable, but not in black and white. Thus, we have put symbols on them. \\
+Response: This figure is also much different now, but we have tried to make all of the curves distinguishable by using different colors along with symbols so that they are even distinguishable in black and white.\\
 
@@
 That is, we use only linear calculations to find growth rates without using normal modes (eigenvalues and eigenvectors).
-While most non-modal analysis in the literature seems to center around the calculation of the $epsilon$-pseudospectrum, much also includes the calculation of $||e^{\mathbf{A} t}||$ and $||e^{\mathbf{A} t} \mathbf{u}(0)||$ with various optimal $\mathbf{u}(0)$ vectors.
+While most non-modal analysis in the literature seems to center around the calculation of the $epsilon$-pseudospectrum, much also includes the calculation of $||e^{\mathbf{A} t}||$ and $||e^{\mathbf{A} t} \mathbf{u}(0)||$ with various optimal $\mathbf{u}(0)$ vectors, i.e.
+to get the numerical range.
+More importantly, however, it would be helpful to have a more theoretical prediction that doesn't require simulation.
+In that regard, we have come across a method~\cite{camargo1998} that can replicate our numerical simulations with a simple calculation that looks much more like a typical non-modal calculation.
+See Eqs.
+9 and 10.
 \\
@@
 We therefore have replaced our previous wording with the following: ``we start with an ensemble of random initial conditions, evolve them linearly for a time $\tau_{\rm{nl}}$, and then take the time and ensemble average growth rate of these curves.
-This procedure mimicks what would happen if the turbulence evolved linearly for a time $\tau_{\rm{nl}}$, randomized, and then repeated.''
+This procedure mimicks what would happen if the turbulence evolved linearly for a time $\tau_{\rm{nl}}$, randomized, and then repeated.'' Also, we use the non-modal calculation discussed above, added more equations, and replaced Figure 2 to make our procedure more clear.
 \\
@@
 
-Response: \\
+Response: We have now tried to define all symbols and have also removed a lot of unnecessary parts of the paper and replaced them with better text.
+We hope the major changes have made the manuscript significantly better and more valuable.
+\\
 

### 2014-02-24 Transient_growth_prl.tex
Feb 23

@@
 In nonlinear dynamical systems with highly nonorthogonal linear eigenvectors, linear non-modal analysis is more useful than normal mode analysis in predicting turbulent properties. 
-However, the non-trivial time evolution of non-modal structures makes quantitative prediction difficult.
+However, the non-trivial time evolution of non-modal structures makes quantitative understanding and prediction difficult. 
 We present a technique to overcome this difficulty by modelling the effect that the advective nonlinearities have on spatial turbulent structures.
 The nonlinearities are taken as a periodic randomizing force with
-period consistent with critical balance arguments. We apply this technique to a model of drift wave turbulence in the Large Plasma Device (LAPD) 
+time scale consistent with critical balance arguments. We apply this technique to a model of drift wave turbulence in the Large Plasma Device (LAPD) 
 [W. Gekelman \emph{et al.}, Rev. Sci. Inst. {\bf 62}, 2875 (1991)], where non-modal effects dominate the turbulence.
@@
 arguments to use a characteristic \emph{linear} time. Since several linear time scales exist, we test them all and compare the results to determine which works best.
-The procedure ultimately produces an effective growth rate spectrum that can be used to predict turbulent properties such as saturation levels through mixing length arguments.
+The procedure ultimately produces growth rate spectra that can be used to predict turbulent properties such as saturation levels and transport rates through mixing length arguments.
 While the concepts behind this technique are general enough to be applied to many nonlinear dynamical systems, the details vary for each case, 
@@
 
-The equations are global and we retain only advective nonlinearities, which are written with Poisson brackets. 
+The equations are global and we retain advective nonlinearities, which are written with Poisson brackets, but we neglect other nonlinear terms. 
 We add artificial diffusion and viscosity terms with small numerical coefficients ($10^{-3}$)
@@
 where $m$ and $n$ represent the azimuthal and axial Fourier mode numbers. 
-$\diff{E_{l}(m,n)}{t}$ comes from the linear terms in Eqs.~\ref{ni_eq}-\ref{te_eq}, $\diff{E_{nl}(m,n)}{t}$ comes from the nonlinear terms. 
-$\diff{E_{l}(m,n)}{t}$ describes the injection (or dissipation) of energy into the fluctuations from the free energy in the equilibrium gradients.
-$\diff{E_{nl}(m,n)}{t}$ describes the energy exchange between fluctuations with different $m,n$ and it is conservative: $\sum_{m,n} \diff{E_{nl}(m,n)}{t} = 0$.
+$\diff{E_{l}(m,n)}{t}$ comes from the linear terms in Eqs.~\ref{ni_eq}-\ref{te_eq}. $\diff{E_{nl}(m,n)}{t}$ comes from the nonlinear terms. 
+$\diff{E_{l}(m,n)}{t}$ represents the injection (or dissipation) of energy into the fluctuations from the free energy in the equilibrium gradients.
+$\diff{E_{nl}(m,n)}{t}$ accounts for the energy exchange between fluctuations with different $m,n$ and it is conservative: $\sum_{m,n} \diff{E_{nl}(m,n)}{t} = 0$.
 Moreover, in quasi-steady state turbulence, the rate of energy injection (or dissipation) into the fluctuations at each $m,n$ by the linear terms must be balanced by
@@
 and it is proportional to the transport rate~\cite{terry2006b}. 
-$\gamma_l(m,n)$ may be calculated from the spatial structures of of the plasma state variables, so it has meaning in turbulent plasma states. We plot
+$\gamma_l(m,n)$ may be calculated from the spatial structures of the plasma state variables, so it is always well-defined, even in a turbulent plasma. We plot
 $\gamma_l(m,n)$ in Fig.~\ref{n0_n1_growth} a) for both the linear and steady state turbulent stages of our nonlinear simulation. The linear stage, which occurs when fluctuations
@@
 all $n=0$ linear eigenmodes are stable. Notice also that the $n=1,m=20$ mode decays transiently before growing exponentially with growth rate of the most unstable eigenmode. This transient
-is another non-modal result.
+decay is also a non-modal result since $\gamma_s(n=1,m=20) > 0$.
 
@@
 In our system, the turbulence is not subcritical in the traditional sense, but it has a subcritical component in the $n=0$ modes. They have no unstable eigenmodes, but still inject energy into the
-turbulence -- $\gamma_l(m<50,n=0) > 0$, yet $\gamma_s(m,n=0) < 0$. It is our goal, then, to use linear non-modal calculations to understand this behavior and to move toward predictive capability
+turbulence; $\gamma_l(m<50,n=0) > 0$, yet $\gamma_s(m,n=0) < 0$. It is our goal, then, to use linear non-modal calculations to understand this behavior and to move toward predictive capability
 of $\gamma_l$.
 To accomplish this, we make the anzatz that nonlinearities randomize the turbulent spatial structure at each wavenumber on a time scale of 
-one eddy decorrelation time, while the linearities evolve the spatial structures deterministically. From this we can calculate a ``Transient Growth Rate'' spectrum,
-which is our prediction of the turbulent growth rate spectrum. To illustrate,
+one eddy decorrelation time, while the linearities evolve the spatial structures deterministically. From this we can calculate a non-modal $\gamma_l$ spectrum,
+which is our prediction of the turbulent $\gamma_l$ spectrum. To illustrate,
 we begin by taking Eqs.~\ref{ni_eq}-\ref{te_eq} and Fourier decomposing in the azimuthal and axial directions.  Then, we discritize in the radial direction and
@@
 inner product is possible, a physically relevant one such as an energy inner product is generally preferred~\cite{camargo1998,schmid2007,camporeale2010}. 
-Recall that the inner product of two vectors may be written $\left< \mathbf{x},\mathbf{y} \right> = \mathbf{y}^{\dagger} \mathbf{M} \mathbf{x}$,
+Recall that the inner product of two vectors may be written $\left< \mathbf{x},\mathbf{y} \right> = \mathbf{y}^{\dagger} \mathbf{M} \mathbf{x}$.
 We choose $\mathbf{M}$ so that $||\mathbf{v}||^2 = \left< \mathbf{v},\mathbf{v} \right> = E$. Furthermore, it is convenient in
@@
 The solution of Eq.~\ref{lin_eq_A} is $\mathbf{u}(t) = e^{\mathbf{A} t} \mathbf{u}(0)$, which
-depends on the initial condition $\mathbf{u}(0)$ in addition to the spectral properties of $\mathbf{A}$. For purposes of turbulent growth rate prediction, we are interested in
+depends on the initial condition $\mathbf{u}(0)$. For purposes of turbulent growth rate prediction, we are interested in
 the behavior of $G(t) = E(t)/E(0) = ||\mathbf{u}(t)||^2/||\mathbf{u}(0)||^2$ and its rate of growth.
@@
 
-\begin{figure}
-\centerline{\includegraphics[width=0.55\textwidth]{m20n0_E_ev}}
-\caption{An ensemble of growth ratio curves (solid gray lines, with the solid black line the ensemble average)
-that start with a random initial condition $u(0)$ and evolve under the linear operator.}
-\label{m20n0_E_ev}
-\end{figure}
-
 It is common practice in normal mode analysis to look for the least stable eigenmode. 
-For the non-modal case, it is common to study the properties of $G_{\rm{max}}(t) = ||e^{\mathbf{A} t}||$ because if this is greater at any time, fluctuations may be amplified, 
+For the non-modal case, it is common to study the properties of $G_{\rm{max}}(t) = ||e^{\mathbf{A} t}||$ because if this is greater than unity at any time, fluctuations may be amplified, 
 leading to subcritical turbulence~\cite{trefethen2005,schmid2007}.
@@
 We contend that the key to understanding and prediction turbulent properties through non-modal analysis is to successfully model the effect that the nonlinearities have on the transient linear processes.
-To this effect, we note that the advective nonlinearity in Eq.~\ref{mat_eq} has the form of the state vector divided by a time $\tau_{\rm{nl}} \sim (v_E k_\perp)^{-1}$. This nonlinear
+To this effect, we note that the advective nonlinearity in Eq.~\ref{mat_eq} has the form of the state vector divided by a time $\tau_{nl} \sim (v_E k_\perp)^{-1}$. This nonlinear
 time scale is generally associated with the eddy turnover or decorrelation time. We therefore present a heuristic model of the nonlinearities 
 as a randomizing force that acts on this characteristic nonlinear time scale.
-Specifically, we start with an ensemble of random initial conditions, evolve them linearly for a time $\tau_{\rm{nl}}$, and then take the time and ensemble average growth rate of these curves.
-This procedure mimicks what would happen if the turbulence evolved linearly for a time $\tau_{\rm{nl}}$, randomized, and then repeated. Now, this procedure does not require actual linear
+Specifically, we start with an ensemble of random initial conditions, evolve them linearly for a time $\tau_{nl}$, and then take the time and ensemble average growth rate of these curves.
+This procedure mimicks what would happen if the turbulence evolved linearly for a time $\tau_{nl}$, randomized, and then repeated. Now, this procedure does not require actual linear
 simulations from an ensemble of random initial conditions. Recall that the time evolution of the energy from an initial condition is
@@
 \label{E_t_ensemble_avg}
-\left< E(t)/E(0) \right>_{{\rm{ens}}} = \frac{1}{4 N_r} {\rm{trace}} e^{\mathbf{A} t} e^{\mathbf{A}^{\dagger}t}.
+\left< E(t)/E(0) \right>_{{\rm{ens}}} = \frac{1}{4 N_r} {\rm{tr}} \{ e^{\mathbf{A} t} e^{\mathbf{A}^{\dagger}t} \}.
 \eeq
@@
 
+\begin{figure}
+\centerline{\includegraphics[width=0.55\textwidth]{m20n0_E_ev}}
+\caption{An ensemble of initially randomized growth ratio curves (solid gray lines, with the solid black line their average, and the dashed orange line the average as calculated from Eq.~\ref{E_t_ensemble_avg}).
+The vertical lines indicate various linear time scales.}
+\label{m20n0_E_ev}
+\end{figure}
+
 Mathematically, our procedure is to calculate $\gamma_l$ by the following formula:
@@
 \label{gamma_l_calc}
-\gamma_l = \frac{1}{\tau_{nl}} \int_0^{\tau_{nl}} \pdiff{E(t)}{t}/ 2 E(t) dt = \frac{1}{2 \tau_{nl}} {\rm{Log}} \left[ \frac{E(\tau_{nl})}{E(0)}\right]
+\gamma_l = \frac{1}{\tau_{nl}} \int_0^{\tau_{nl}} \frac{\pdiff{E(t)}{t}}{2 E(t)} dt = \frac{1}{2 \tau_{nl}} {\rm{Log}} \left[ \frac{E(\tau_{nl})}{E(0)}\right]
 \eeq
 where $E(t)$ is the ensemble averaged energy calculated from Eq.~\ref{E_t_ensemble_avg}, and $E(0) = 1$ by our normalization.
-In order to move toward predictive capability, we must estimate $\tau_{\rm{nl}}$ with only knowledge of linear (modal or non-modal) information. 
-We thus invoke the conjecture of \emph{critical balance}, which posits that the nonlinear time scale equals the linear time scale at all spatial scales~\cite{schekochihin2012}, which is clear from 
+In order to move toward predictive capability, we must estimate $\tau_{nl}$ with only knowledge of linear (modal or non-modal) information. 
+We thus invoke the conjecture of \emph{critical balance}, which posits that the nonlinear time scale equals the linear time scale at all spatial scales~\cite{schekochihin2012}, which follows from 
 Eqs.~\ref{dEdt_def} and~\ref{steady_state}, which lead to $\gamma_l = - \gamma_{nl}$. Now there are several linear time scales that we may choose to test. We label these times in Fig.~\ref{m20n0_E_ev}
 for the case of $n=0, m=20$. The first linear time is the linear eigenmode frequency, labeled
-$\tau_{nl} = \omega_s^{-1}$. However, $\omega_s = 0$ for $n=0$ linear eigenmodes, so we are forced to use $\omega_s$ for the fastest growing $n=1$ eigenmode at each $m$ to get a meaningful time scale.
-Second is the parallel free-streaming time of the electrons $\tau_{nl} = t_{\rm{st}} = L_\para/v_{t e}$ often cited in critical balance arguments~\cite{barnes}. Third is the time before modal effects
-take over, which can be approximated as the time when $E(t)$ turns over as seen in Fig.~\ref{n0_n1_growth}, where $E(t)$ can be either a maximum or minimum, labeled $\tau_{nl} = E_{\rm{max}}$. 
-Fourth, we take the $\tau_{nl} = 1/\gamma_{nl} = - 1/\gamma_l$. Inserting this into Eq.~\ref{gamma_l_calc} gives
+$\omega_s^{-1}$. However, $\omega_s = 0$ for $n=0$ linear eigenmodes, so we are forced to use $\omega_s$ for the fastest growing $n=1$ eigenmode at each $m$ to get a meaningful time scale.
+Second is the parallel free-streaming time of the electrons $t_{\rm{st}} = L_\para/v_{t e}$ often cited in critical balance arguments~\cite{barnes2011}. Third is the time before modal effects
+take over, which can be approximated as the time when $E(t)$ turns over.
+As seen in Fig.~\ref{n0_n1_growth}, $E(t)$ can be either a maximum or minimum before turning over.
+We label this time $E_{\rm{max}}$.
+Fourth, we use the steady-state condition $\gamma_l = - \gamma_{nl}$ and the approximation $\tau_{nl} \sim 1/|\gamma_{nl}|$ to get $\tau_{nl} = 1/|\gamma_l|$.
+Inserting this into Eq.~\ref{gamma_l_calc} gives
 
@@
 \eeq
-In other words, $\gamma_l$ is found by the inverse of the time when $E(t)$ grows to the value of $e^2$ or decays to the value of $e^{-2}$, labeled $\tau_{nl} = e^2$. We also indicate two times
-labeled $\gamma_{\rm{sim}}$, which when inserted into Eq.~\ref{gamma_l_calc} give the ``correct'' $\gamma_l$ ($\gamma_l(m=20,n=0)$ calculated directly from the nonlinear simulation).
+In other words, we find the time at which $E(t)$ grows to the value of $e^2$ or decays to the value of $e^{-2}$, and then use this time to get $\gamma_l$.
+We label this time $e^2$.
+In Fig.~\ref{m20n0_E_ev},
+we also indicate two times labeled $\gamma_{\rm{sim}}$, which when inserted into Eq.~\ref{gamma_l_calc} give the ``correct'' $\gamma_l(m=20,n=0)$ calculated directly from the nonlinear simulation.
 
@@
 $\gamma_s$ for reference.
+All of the non-modal growth rates are positive at $n=0$ for low $m$, like $\gamma_l$ from the simulation and unlike $\gamma_s$, indicating that the non-modal analysis can reveal
+what normal mode analysis cannot in this turbulent system. Furthermore, the choice of linear time scale does not significantly affect the qualitative picture. The non-modal analysis does not always
+predict complete stability at $n=1$ for all choices of linear time scale, but it does indicate that $n=1$ modes can dissipate rather than inject fluctuation energy despite the presence of unstable
+linear eigenmodes. Interestingly, the linear eigenmode time scale $\omega_s^{-1}$ gives the best match to $\gamma_{\rm{sim}}$, however, this may not be a general result.
 
@@
 \centerline{\includegraphics[width=0.5\textwidth]{gamma_comparisons}}
-\caption{The linear, turbulent, and transient growth rate spectra. The transient spectra are calculated by the average growth rate over a time $\omega_s^{-1}(m,n=1)$ for the ensemble average
-growth ratio curves with the cascade curve using $\omega_s^{-1}$ at $m/2$ rather than $m$.}
+\caption{The growth rate spectra of the spectral growth rate $\gamma_s$, turbulent simulation growth rate $\gamma_{\rm{sim}}$, and growth rates calculated from the non-modal procedure in Eq.~\ref{gamma_l_calc}.
+The growth rates as a function of $m$ for a) $n=0$ and b) $n=1$.}
 \label{gamma_comparisons}

### 2014-02-25 Referee_Response1/referee_response.tex
Feb 24

@@
 Before responding to specific comments, questions, and requests from the referees, we would like to note that we have significantly changed parts of the manuscript. First, we have
-changed the way we calculate the $|e^{A t}u(0)|$ curves where we used an ensemble of initial conditions. Before we made many initial conditions, evolved them by this formula and then
-calculated the ensemble average. This way seemed a bit like performing numerical simulations. Now we have an analytic method of calculation which doesn't involve setting up any initial
+changed the way we calculate the $|e^{A t}u(0)|$ curves, in which we used an ensemble of initial conditions for the $u(0)$. Before we explicitely randomized a thousand or so initial conditions, 
+evolved them by this formula and then calculated the ensemble average. This way seemed a bit like performing numerical simulations as pointed out by referee \#2. 
+Now we have an analytic method of calculation which doesn't involve setting up any initial
 conditions, so it is not at all like simulation (see Eq. 10) and is much more elegant.
@@
 Second, recall that our method involved choosing a linear time scale as a proxy for the eddy turnover time, for which we chose the inverse of a linear eigenvalue frequency. 
-As the first referee pointed out, we didn't properly justify this choice. Since we have no theoretical justification, 
+As referee \#1 pointed out, we didn't properly justify this choice. Since we have no theoretical justification for this choice over others, 
 we have now included a few more linear time scales and studied the result for each of these. Furthermore, we no longer claim that our method
@@
 
-Otherwise, we have made various minor changes and changed two of the figures to make sense with the changes mentioned above. 
-Overall the manuscript is significantly different than before, and we feel it is much stronger.
+Otherwise, we have made various minor changes and changed two of the figures to fit in with the above changes to the manuscript. 
+Overall the paper is significantly different than before, and we feel it is much stronger, thanks in part to the comments from the referees.
+Below we respond specifically to the points brought forth by the referees.
 
-\section{Referee 1}
+\section{Referee \#1}
 
@@
 Response: This is a valid criticism. We thought a lot about this and tried a few different linear time scales. The one we chose worked best, but it is difficult to justify it otherwise.
-We agree that this is the weakest link of the paper. We also think that the suggestion of presenting several choices of linear time scales is a good one, so this is what we've done.
+We agree that this was the weakest link of the paper. We also thought that the suggestion of presenting several choices of linear time scales was a good one, so this is what we've done.
 We've presented four different choices of linear time scale and compared the results to each other. We have also softened our language on predictive capability and presented our results
@@
 
-Response: We do not really understand this question.
-We do not use any closure to write equation 5.
+Response: We do not really understand this question. If by closure, you mean an approximation of the nonlinearity, then we do not use any closure to write equation 5. 
 It is an exact equation that can be derived from the model equations~\cite{friedman2012b,friedman2013}.
-We have also now written it differently to make it simpler and more obvious. \\
+We have also now written it differently to make it simpler and more transparent we hope. \\
 
@@
 
-Response: We acknowledge that this was too cryptic a statement.
-In it's place, we have inserted the following, ``These properties can be calculated through mixing length formulas if the turbulent growth rate spectrum is known.'' Then, as before, we proceed to calculate the turbulent growth rate spectrum from the nonlinear simulation (shown in Figure 1a) and also calculate it from our non-modal linear procedure.
-\\
+Response: We acknowledge that this was too cryptic a statement. In it's place, we have inserted the following, 
+``The procedure ultimately produces growth rate spectra that can be used to predict turbulent properties such as saturation levels and transport rates through mixing length arguments.''
+Then, as before, we proceed to calculate the growth rate spectrum from the nonlinear simulation (shown in Figure 1a) and also calculate it from our non-modal linear procedure. \\
 
@@
 
-\section{Referee 2}
+\section{Referee \#2}
 
@@
 
-Response: We contend that our method is a form of 'non-modal analysis'.
+Response: We contend that our method (especially our new one) is a form of 'non-modal analysis'.
 That is, we use only linear calculations to find growth rates without using normal modes (eigenvalues and eigenvectors).
 While most non-modal analysis in the literature seems to center around the calculation of the $epsilon$-pseudospectrum, much also includes the calculation of $||e^{\mathbf{A} t}||$ and $||e^{\mathbf{A} t} \mathbf{u}(0)||$ with various optimal $\mathbf{u}(0)$ vectors, i.e.
-to get the numerical range.
+to get the numerical and numerical abscissa.
 More importantly, however, it would be helpful to have a more theoretical prediction that doesn't require simulation.
@@
 Response: We agree that this was a confusing way to word our procedure.
-We therefore have replaced our previous wording with the following: ``we start with an ensemble of random initial conditions, evolve them linearly for a time $\tau_{\rm{nl}}$, and then take the time and ensemble average growth rate of these curves.
-This procedure mimicks what would happen if the turbulence evolved linearly for a time $\tau_{\rm{nl}}$, randomized, and then repeated.'' Also, we use the non-modal calculation discussed above, added more equations, and replaced Figure 2 to make our procedure more clear.
+We therefore have replaced our previous wording with the following: ``we start with an ensemble of random initial conditions, evolve them linearly for a time $\tau_{nl}$, then take the time and ensemble averaged growth rate of these curves.
+This procedure calculates the average growth rate of the turbulence as if the turbulence evolved linearly for a time $\tau_{nl}$, randomized, and then repeated.''
+Also, we use the non-modal calculation discussed above, added more equations, and replaced Figure 2 to make our procedure more clear.
 \\
@@
 
-Response: The reason why $T$ is missing from the definition of $\gamma$ is because $\diff{E(m,n)}{t} = Q(m,n) + D(m,n) + \sum_{m',n'} T(m,m',n,n')$ is virtually zero when the turbulence is in steady state (by definition of steady state).
-It may not be zero at a given time, but it is zero on average, and thus not suitable for use in defining the growth rate spectrum, which is a time-averaged property (see Figure 1b for the average and the spread in the growth rate).
+Response: Note that we have changed our notation in the text, and added some more equations and discussions of this so that this issue should now be clear. But we will explain here anyway.
+
+The reason why $T$ is missing from the definition of $\gamma$ is because $\diff{E(m,n)}{t} = Q(m,n) + D(m,n) + \sum_{m',n'} T(m,m',n,n')$ is virtually zero when the turbulence is in steady state (by definition of steady state).
+It may not be zero at a given time, but it is zero on average, and thus not suitable for use in defining the growth rate spectrum, which is a time-averaged property (see Figure 1a for the average and the spread in the growth rate).
 If we were to include $T$ in the definition, the turbulent growth rate curves in Figure 1b would be zero with some finite spread.
@@
 When energy is nonlinearly transferred into modes with negative growth rate, that energy is dissipated.
-In the case of normal linear operators, the unstable linear eigenvectors inject energy, transfer that energy to stable linear eigenvectors, which then dissipate the energy.
+In the case of normal linear operators, the unstable linear eigenvectors inject energy, the nonlinearities transfer that energy to stable linear eigenvectors which then dissipate the energy.
 For normal operators, one would expect this growth rate to be pretty similar to the fastest growing eigenmode growth rate, though it could be smaller.

### 2014-02-25 Transient_growth_prl.tex
Feb 24

@@
 nonlinear decorrelation time of the turbulent system, but in order to enable predictive capability, we employ critical balance
-arguments to use a characteristic \emph{linear} time. Since several linear time scales exist, we test them all and compare the results to determine which works best.
+arguments to use a characteristic \emph{linear} time. Since there is no obvious single linear time scale, we test several and compare the results to determine which works best.
 The procedure ultimately produces growth rate spectra that can be used to predict turbulent properties such as saturation levels and transport rates through mixing length arguments.
@@
 
-The model describes pressure-gradient-driven turbulence in the uniformly magnetized, cylindrical plasma
+The model we use describes pressure-gradient-driven turbulence in the uniformly magnetized, cylindrical plasma
 produced by the Large Plasma Device (LAPD)~\cite{gekelman1991}. 
@@
 magnetic-field-aligned ($k_\para=0$) convective filaments transport density across the equilibrium density gradient, setting up $k_\para=0$ density filaments. 
-These filaments are unstable to a secondary drift waves, which grow on the periphery of the filaments. 
+These filaments are unstable to secondary drift waves, which grow on the periphery of the filaments. 
 These drift waves, which have finite $k_\para$, nonlinearly couple to one another and reinforce the original convective filaments.
@@
 meaning that the convective transport is the only step responsible for energy injection into the fluctuations.
+
 Deriving an equation for the evolution of the energy from Eqs.~\ref{ni_eq}-\ref{te_eq}~\cite{friedman2012b,friedman2013}, we may symbolically write
@@
 Since transient growth is a purely linear phenomenon, it has been a goal of researchers to understand and predict the onset of subcritical turbulence using only linear, non-modal calculations. 
-In our system, the turbulence is not subcritical in the traditional sense, but it has a subcritical component in the $n=0$ modes. They have no unstable eigenmodes, but still inject energy into the
-turbulence; $\gamma_l(m<50,n=0) > 0$, yet $\gamma_s(m,n=0) < 0$. It is our goal, then, to use linear non-modal calculations to understand this behavior and to move toward predictive capability
-of $\gamma_l$.
+In our system, the turbulence is not subcritical in the traditional sense, but it has a subcritical component because $\gamma_l(m<50,n=0) > 0$, yet $\gamma_s(m,n=0) < 0$. 
+It is our goal, then, to use linear non-modal calculations to understand this behavior and to move toward predictive capability of $\gamma_l$.
 To accomplish this, we make the anzatz that nonlinearities randomize the turbulent spatial structure at each wavenumber on a time scale of 
-one eddy decorrelation time, while the linearities evolve the spatial structures deterministically. From this we can calculate a non-modal $\gamma_l$ spectrum,
+one eddy decorrelation time, while the linearities evolve the spatial structures deterministically. From this we can calculate a non-modal $\gamma_{\rm{n-m}}$ spectrum,
 which is our prediction of the turbulent $\gamma_l$ spectrum. To illustrate,
@@
 and the second term the nonlinearities. Note that for each $m,n$, there exist $4 N_r$ linearly independent, but nonorthogonal eigenvectors, where $N_r$ is the number of radial grid points. 
-Hence forth, we drop the $m,n$ subscripts.
+Hence forth, we drop the $m,n$ Fourier subscripts.
 
@@
 Furthermore, it isn't obvious what kind of spatial structure or structures will come to dominate a turbulent system.
-In non-normal systems, unlike in the normal systems, optimal structures don't amplify themselves, rather, they evolve while increasing the total fluctuating energy.
+In non-normal systems, unlike in normal systems, optimal structures don't amplify themselves, rather, they evolve while increasing the total fluctuating energy.
 
-We contend that the key to understanding and prediction turbulent properties through non-modal analysis is to successfully model the effect that the nonlinearities have on the transient linear processes.
+We contend that the key to understanding and predicting turbulent properties through non-modal analysis is 
+to successfully model the effect that the nonlinearities have on the transient linear processes. 
 To this effect, we note that the advective nonlinearity in Eq.~\ref{mat_eq} has the form of the state vector divided by a time $\tau_{nl} \sim (v_E k_\perp)^{-1}$. This nonlinear
@@
 as a randomizing force that acts on this characteristic nonlinear time scale.
-Specifically, we start with an ensemble of random initial conditions, evolve them linearly for a time $\tau_{nl}$, and then take the time and ensemble average growth rate of these curves.
-This procedure mimicks what would happen if the turbulence evolved linearly for a time $\tau_{nl}$, randomized, and then repeated. Now, this procedure does not require actual linear
-simulations from an ensemble of random initial conditions. Recall that the time evolution of the energy from an initial condition is
+Specifically, we start with an ensemble of random initial conditions, evolve them linearly for a time $\tau_{nl}$, then take the time and ensemble averaged growth rate of these curves.
+This procedure calculates the average growth rate of the turbulence as if the turbulence evolved linearly for a time $\tau_{nl}$, randomized, and then repeated. 
+Now, this procedure does not require actual linear simulations from an ensemble of random initial conditions. To see this, recall that the time evolution of the energy from an initial condition is
 
@@
 \begin{figure}
-\centerline{\includegraphics[width=0.55\textwidth]{m20n0_E_ev}}
+\centerline{\includegraphics[width=0.45\textwidth]{m20n0_E_ev}}
 \caption{An ensemble of initially randomized growth ratio curves (solid gray lines, with the solid black line their average, and the dashed orange line the average as calculated from Eq.~\ref{E_t_ensemble_avg}).
@@
 
-Mathematically, our procedure is to calculate $\gamma_l$ by the following formula:
+Mathematically, our procedure is to calculate $\gamma_{\rm{n-m}}$ by the following formula:
 
@@
 \label{gamma_l_calc}
-\gamma_l = \frac{1}{\tau_{nl}} \int_0^{\tau_{nl}} \frac{\pdiff{E(t)}{t}}{2 E(t)} dt = \frac{1}{2 \tau_{nl}} {\rm{Log}} \left[ \frac{E(\tau_{nl})}{E(0)}\right]
+\gamma_{\rm{n-m}} = \frac{1}{\tau_{nl}} \int_0^{\tau_{nl}} \frac{\pdiff{E(t)}{t}}{2 E(t)} dt = \frac{1}{2 \tau_{nl}} {\rm{Log}} \left[ \frac{E(\tau_{nl})}{E(0)}\right]
 \eeq
@@
 In order to move toward predictive capability, we must estimate $\tau_{nl}$ with only knowledge of linear (modal or non-modal) information. 
-We thus invoke the conjecture of \emph{critical balance}, which posits that the nonlinear time scale equals the linear time scale at all spatial scales~\cite{schekochihin2012}, which follows from 
-Eqs.~\ref{dEdt_def} and~\ref{steady_state}, which lead to $\gamma_l = - \gamma_{nl}$. Now there are several linear time scales that we may choose to test. We label these times in Fig.~\ref{m20n0_E_ev}
+We thus invoke the conjecture of \emph{critical balance}, which posits that the nonlinear time scale equals the linear time scale at all spatial scales~\cite{schekochihin2012}. This follows from 
+the previously derived steady-state result: $\gamma_l = - \gamma_{nl}$.
+
+Now there are several linear time scales that we may choose to test. We label these times in Fig.~\ref{m20n0_E_ev}
 for the case of $n=0, m=20$. The first linear time is the linear eigenmode frequency, labeled
@@
 Second is the parallel free-streaming time of the electrons $t_{\rm{st}} = L_\para/v_{t e}$ often cited in critical balance arguments~\cite{barnes2011}. Third is the time before modal effects
-take over, which can be approximated as the time when $E(t)$ turns over.
-As seen in Fig.~\ref{n0_n1_growth}, $E(t)$ can be either a maximum or minimum before turning over.
+take over, which can be approximated as the time when $E(t)$ turns over. As seen in Fig.~\ref{n0_n1_growth}, $E(t)$ can be either a maximum or minimum before turning over. 
 We label this time $E_{\rm{max}}$. 
-Fourth, we use the steady-state condition $\gamma_l = - \gamma_{nl}$ and the approximation $\tau_{nl} \sim 1/|\gamma_{nl}|$ to get $\tau_{nl} = 1/|\gamma_l|$.
+Fourth, we use the steady-state condition $\gamma_l = - \gamma_{nl}$ and the approximation $\tau_{nl} \sim 1/|\gamma_{nl}|$ to get $\tau_{nl} = 1/|\gamma_{\rm{n-m}}|$. 
 Inserting this into Eq.~\ref{gamma_l_calc} gives
@@
 \label{fourth_gamma_cond}
- {\rm{Log}} \left[ E(1/\gamma_l)\right] = \pm 2 \ \rightarrow E(1/\gamma_l) = e^{\pm 2}.
+ {\rm{Log}} \left[ E(1/\gamma_{\rm{n-m}})\right] = \pm 2 \ \rightarrow E(1/\gamma_{\rm{n-m}}) = e^{\pm 2}.
 \eeq
-In other words, we find the time at which $E(t)$ grows to the value of $e^2$ or decays to the value of $e^{-2}$, and then use this time to get $\gamma_l$.
-We label this time $e^2$.
+In other words, we find the time at which $E(t)$ grows to the value of $e^2$ or decays to the value of $e^{-2}$, and then use this time to get $\gamma_{\rm{n-m}}$. We label this time $e^2$. 
 In Fig.~\ref{m20n0_E_ev},
-we also indicate two times labeled $\gamma_{\rm{sim}}$, which when inserted into Eq.~\ref{gamma_l_calc} give the ``correct'' $\gamma_l(m=20,n=0)$ calculated directly from the nonlinear simulation.
+we also indicate two times labeled $t_{\rm{cor}}$, which when inserted into Eq.~\ref{gamma_l_calc} give the ``correct'' $\gamma_l(m=20,n=0)$ calculated directly from the nonlinear simulation.
 
-In Fig.~\ref{gamma_comparisons}, we compare the $\gamma_{l}$ spectrum from the nonlinear simulation to $\gamma_l$ from the non-modal procedure using the four different linear time scales. We also show
+In Fig.~\ref{gamma_comparisons}, we compare the $\gamma_{l}$ spectrum from the nonlinear simulation to $\gamma_{n-m}$ from the non-modal procedure using the four different linear time scales.
+We also show
 $\gamma_s$ for reference.
 All of the non-modal growth rates are positive at $n=0$ for low $m$, like $\gamma_l$ from the simulation and unlike $\gamma_s$, indicating that the non-modal analysis can reveal
-what normal mode analysis cannot in this turbulent system. Furthermore, the choice of linear time scale does not significantly affect the qualitative picture. The non-modal analysis does not always
+what normal mode analysis cannot in this turbulent system. Furthermore, the choice of linear time scale does not significantly affect the qualitative picture.
+
+On the other hand, the non-modal analysis does not always
 predict complete stability at $n=1$ for all choices of linear time scale, but it does indicate that $n=1$ modes can dissipate rather than inject fluctuation energy despite the presence of unstable
-linear eigenmodes. Interestingly, the linear eigenmode time scale $\omega_s^{-1}$ gives the best match to $\gamma_{\rm{sim}}$, however, this may not be a general result.
+linear eigenmodes. Interestingly, the linear eigenmode time scale $\omega_s^{-1}$ gives the best match to $\gamma_l$, however, this may not be a general result.
 
 \begin{figure}
-\centerline{\includegraphics[width=0.5\textwidth]{gamma_comparisons}}
-\caption{The growth rate spectra of the spectral growth rate $\gamma_s$, turbulent simulation growth rate $\gamma_{\rm{sim}}$, and growth rates calculated from the non-modal procedure in Eq.~\ref{gamma_l_calc}.
+\centerline{\includegraphics[width=0.55\textwidth]{gamma_comparisons}}
+\caption{The growth rate spectra of the spectral growth rate $\gamma_s$, turbulent simulation growth rate $\gamma_l$, and growth rates calculated from the non-modal procedure in Eq.~\ref{gamma_l_calc}.
 The growth rates as a function of $m$ for a) $n=0$ and b) $n=1$.}

### 2014-03-04 Referee_Response1/referee_response.tex
March 4

@@
 {\bf Response to Referees \\
- \\
  \\} 
-Brett Friedman \\
-Updated: \today \\
+Brett Friedman and Troy Carter
+%Updated: \today \\
 
@@
 
-\section{Overview of Changes}
+\section{Referee \#1}
 
-Before responding to specific comments, questions, and requests from the referees, we would like to note that we have significantly changed parts of the manuscript. First, we have
-changed the way we calculate the $|e^{A t}u(0)|$ curves, in which we used an ensemble of initial conditions for the $u(0)$. Before we explicitely randomized a thousand or so initial conditions, 
-evolved them by this formula and then calculated the ensemble average. This way seemed a bit like performing numerical simulations as pointed out by referee \#2. 
-Now we have an analytic method of calculation which doesn't involve setting up any initial
-conditions, so it is not at all like simulation (see Eq. 10) and is much more elegant.
-
-Second, recall that our method involved choosing a linear time scale as a proxy for the eddy turnover time, for which we chose the inverse of a linear eigenvalue frequency. 
-As referee \#1 pointed out, we didn't properly justify this choice. Since we have no theoretical justification for this choice over others, 
-we have now included a few more linear time scales and studied the result for each of these. Furthermore, we no longer claim that our method
-is predictive because we have no way to know \emph{a priori} which linear time scale will work best. However, all of the linear time scales do work well to a certain degree, so
-we do obtain a degree of predictivity.
-
-Otherwise, we have made various minor changes and changed two of the figures to fit in with the above changes to the manuscript. 
-Overall the paper is significantly different than before, and we feel it is much stronger, thanks in part to the comments from the referees.
-Below we respond specifically to the points brought forth by the referees.
-
-\section{Referee \#1}
+\emph{This is an interesting paper, bringing non-modal analysis to bear upon a problem of
+plasma turbulence. Non-modal analysis has been used very effectively in hydrodynamics.
+Its use in plasma physics is a relatively recent trend, and holds the promise of providing
+new perspectives on old problems. This paper, which includes theory as well as a strong
+experimental connection, is consistent with the criteria for publication in PRL. Having
+said that, we would like to request the authors to address the following questions:} \\
 
@@
 
-Response: This is a valid criticism. We thought a lot about this and tried a few different linear time scales. The one we chose worked best, but it is difficult to justify it otherwise.
+Response: First we would like to thank the referee for the careful reading of the paper and the nice review. As for the first comment, we believe that this is a valid criticism. 
+We thought a lot about this before submitting the first manuscript and tried a few different linear time scales. The one we chose worked best, but it is difficult to justify it otherwise.
 We agree that this was the weakest link of the paper. We also thought that the suggestion of presenting several choices of linear time scales was a good one, so this is what we've done.
 We've presented four different choices of linear time scale and compared the results to each other. We have also softened our language on predictive capability and presented our results
-as a way of understanding the turbulent growth rate through linear non-modal calculations with the ultimate goal of achieving prediction. \\
+as a way of understanding the turbulent growth rate through linear non-modal calculations with the ultimate goal of achieving prediction. Overall, we have modified the manuscript quite
+a bit to make these changes, including changing the last two figures. \\
 
@@
 
-Response: We do not really understand this question. If by closure, you mean an approximation of the nonlinearity, then we do not use any closure to write equation 5. 
-It is an exact equation that can be derived from the model equations~\cite{friedman2012b,friedman2013}.
+Response: If by closure, you mean an approximation of the
+nonlinearity, then we do not use any closure to write equation 5.  It is an exact equation that can be derived from the model equations~\cite{friedman2012b,friedman2013}.
 We have also now written it differently to make it simpler and more transparent we hope. \\
@@
 
-Response: This figure is also much different now, but we have tried to make all of the curves distinguishable by using different colors along with symbols so that they are even distinguishable in black and white.\\
+Response: This figure is also much different now, but we have tried to make all of the curves distinguishable by using different colors along with symbols so that they are even distinguishable in black and white.
 
@@
 
-Response: We contend that our method (especially our new one) is a form of 'non-modal analysis'.
+Response: First, we would like to thank this referee for the careful reading and insights into the paper. As for the question over 'non-modal analysis', 
+we contend that our method (especially with revisions in the new manuscript) is a form of 'non-modal analysis'.
 That is, we use only linear calculations to find growth rates without using normal modes (eigenvalues and eigenvectors).
@@
 See Eqs.
-9 and 10.
+9 and 10 and Figure 2.
+With this new change, we believe it is very clear that what we are doing is non-modal analysis and that we are not simply using linear simulations.
 \\
@@
 
-Response: We agree that this was a confusing way to word our procedure.
-We therefore have replaced our previous wording with the following: ``we start with an ensemble of random initial conditions, evolve them linearly for a time $\tau_{nl}$, then take the time and ensemble averaged growth rate of these curves.
-This procedure calculates the average growth rate of the turbulence as if the turbulence evolved linearly for a time $\tau_{nl}$, randomized, and then repeated.''
-Also, we use the non-modal calculation discussed above, added more equations, and replaced Figure 2 to make our procedure more clear.
+Response: We agree that this was a confusing way to word our
+procedure. We therefore have removed this statement. At the beginning of the manuscript we offer this description of our model:
+`` We model the turbulent steady state as a series of processes:  (1) the turbulence starts as a spatially random state, (2) linear transient growth deterministically amplifies the turbulent energy (or
+decrease it in wavenumber ranges where linear damping dominates), (3) nonlinear transfer sets in at a specified timescale, terminating the transient growth process and re-randomizing the turbulent state (at which point the cycle repeats).`` This is not the procedure we use to do the actual calculation, but just our simplified description of the turbulent model.
+
+As for the actual procedure, we state: ``In practice, we implement this model by starting with an ensemble of random initial conditions, which we evolve linearly for a time $\tau_{nl}$, 
+and then take the time and ensemble averaged growth rate of these curves.''
+
+Also, we use the non-modal calculation discussed above, added more equations, and replaced Figure 2 to make our procedure more clear. Please read the manuscript and see if this is now clear.
 \\
@@
 It may not be zero at a given time, but it is zero on average, and thus not suitable for use in defining the growth rate spectrum, which is a time-averaged property (see Figure 1a for the average and the spread in the growth rate).
-If we were to include $T$ in the definition, the turbulent growth rate curves in Figure 1b would be zero with some finite spread.
-So we can't use $T$ in the definition.
+If we were to include $T$ in the definition, the turbulent growth rate curves in Figure 1a would be zero with some finite spread.
 
-More informatively, the growth rate we use (We didn't invent it.
-It has been used for many years in numerous papers.) is a measure, by wavenumber, of the rate at which the fluctuations take energy from the equilibrium gradients (or return energy if the growth rate is negative).
+More informatively, the growth rate we use is a measure, by wavenumber, of the rate at which the fluctuations take energy from the equilibrium gradients (or return energy if the growth rate is negative).
 This comes purely from the linear terms.
 The nonlinear terms transfer energy between different fluctuation 'modes', but do not affect the overall energy budget.
-So the picture is that certain modes or structures (those with positive $\gamma$) inject energy from the equilibrium gradients into the fluctuations through the linear terms.
-Those modes then non-linearly transfer energy to other modes in an overall conservative way (like a cascade).
-When energy is nonlinearly transferred into modes with negative growth rate, that energy is dissipated.
-In the case of normal linear operators, the unstable linear eigenvectors inject energy, the nonlinearities transfer that energy to stable linear eigenvectors which then dissipate the energy.
-For normal operators, one would expect this growth rate to be pretty similar to the fastest growing eigenmode growth rate, though it could be smaller.
-But it certainly cannot be larger.
-We have added and modified quite a bit of text and added a reference~\cite{terry2006b}, which uses this growth rate so that it doesn't appear that we have invented it.
+So the picture is that certain modes or structures (those with
+positive $\gamma$) inject energy from the equilibrium gradients into
+the fluctuations through the linear terms. Those modes then
+non-linearly transfer energy to other modes in an overall conservative
+way (like a cascade). When energy is nonlinearly transferred into
+modes with negative growth rate, that energy is dissipated. In the
+case of normal linear operators, the unstable linear eigenvectors
+inject energy, the nonlinearities transfer that energy to stable
+linear eigenvectors which then dissipate the energy. For normal
+operators, one would expect this growth rate to be pretty similar to
+the fastest growing eigenmode growth rate, though it could be
+smaller. But it certainly cannot be larger. We have added and modified
+quite a bit of text and added a reference~\cite{terry2006b}, which
+uses this growth rate definition.
 
@@
 
-Response: We have now tried to define all symbols and have also removed a lot of unnecessary parts of the paper and replaced them with better text.
-We hope the major changes have made the manuscript significantly better and more valuable.
-\\
+Response: We have now defined all symbols and removed unnecessary parts of the paper and replaced them with better text.
+We hope the major changes have made the manuscript significantly better and easier to understand.
 

### 2014-03-04 Transient_growth_prl.tex
March 4

@@
 [W. Gekelman \emph{et al.}, Rev. Sci. Inst. {\bf 62}, 2875 (1991)], where non-modal effects dominate the turbulence.
-We compare the resulting growth rate spectra to that obtained from a nonlinear simulation, showing good qualitative agreement.
+We compare the resulting growth rate spectra to that obtained from a nonlinear simulation, showing good qualitative agreement, especially in comparison to the eigenmode growth rate spectra.
 \end{abstract}
@@
 dynamical systems~\cite{trefethen1993,schmid2007}. A non-normal operator has 
-eigenvectors that are nonorthogonal to one another. One consequence of eigenvector nonorthogonality is that even when all eigenvectors decay exponentially under linear evolution, 
+eigenvectors that are not orthogonal to one another. One consequence of eigenvector nonorthogonality is that even when all eigenvectors decay exponentially under linear evolution, 
 superpositions of eigenvectors can grow, albeit transiently.
@@
 When combined with nonlinear effects, this allows for sustained subcritical turbulence.
-Such behavior is obscured by traditional normal mode analysis, which only effectively describes the long time asymptotic behavior of fluctuations under the 
-action of the linear operator. Transient growth behavior, which can dominate turbulent evolution, can be discovered only through non-modal calculations.
+Such behavior is obscured by traditional normal mode analysis, which only effectively describes the long time asymptotic behavior of fluctuations under  
+action of the linear operator. Transient growth, which can dominate turbulent evolution, can be discovered only through non-modal calculations.
 
 Non-modal analysis has been embraced by the hydrodynamics community over the past two decades in the attempt to explain and predict subcritical turbulence. But the plasma community
-still heavily relies on normal mode analysis to inform turbulent predictions and observations, with a few notable exceptions~\cite{camargo1998,camporeale2010,schekochihin2012}. 
+generally relies on normal mode analysis to inform turbulent predictions and observations, with a few exceptions~\cite{camargo1998,camporeale2010,schekochihin2012}. 
 Furthermore, non-modal treatments have generally been explanatory rather than predictive and have centered around the transition to turbulence in subcritical systems rather 
@@
 This paper takes up the task of developing an approach to understand turbulent properties using only non-modal linear calculations with the goal of 
-ultimately making quantitative predictions. 
-Our approach is to evolve the linear solution from an ensemble of random initial conditions 
-and calculate the average growth rate of the solutions over a specified timescale.  Optimally this time scale would be the
+ultimately making quantitative predictions.  Our approach is to calculate
+an average growth rate of turbulent fluctuations due to linear
+processes, specifically due to transient growth.  We model the
+turbulent steady state as a series of processes:  (1) the turbulence starts as a spatially random state, (2) linear transient growth deterministically amplifies the turbulent energy (or
+decrease it in wavenumber ranges where linear damping dominates), (3) nonlinear transfer sets in at a specified timescale, terminating the
+transient growth process and re-randomizing the turbulent state (at which point the cycle repeats).    Optimally, the timescale for the final step would be the
 nonlinear decorrelation time of the turbulent system, but in order to enable predictive capability, we employ critical balance
 arguments to use a characteristic \emph{linear} time. Since there is no obvious single linear time scale, we test several and compare the results to determine which works best.
-The procedure ultimately produces growth rate spectra that can be used to predict turbulent properties such as saturation levels and transport rates through mixing length arguments.
+The procedure ultimately produces growth rate spectra that can be used
+to predict turbulent properties such as saturation levels and transport rates through mixing length arguments.
 While the concepts behind this technique are general enough to be applied to many nonlinear dynamical systems, the details vary for each case, 
-so we restrict our treatment to one particular turbulence model. For this model, the technique qualitatively reproduces the turbulent growth rate spectrum of the direct 
-nonlinear simulation fairly well, especially in comparison to the linear eigenmode spectrum.
+so we restrict our treatment to one particular turbulence model. For
+this model, the technique does reasonably well in reproducing the turbulent growth rate spectrum of the direct 
+nonlinear simulation, especially in comparison to the linear eigenmode spectrum.
 
@@
 These filaments are unstable to secondary drift waves, which grow on the periphery of the filaments. 
-These drift waves, which have finite $k_\para$, nonlinearly couple to one another and reinforce the original convective filaments.
+These drift waves, which have finite $k_\para$, nonlinearly couple to one another and generate new convective filaments.
 
@@
 The rate of energy injection $\gamma_l(m,n)$ from the equilibrium gradient into the fluctuations is a quantity of great interest~\cite{friedman2012b,terry2006b}. 
-When positive for some wavenumber, turbulence can be sustained. Additionally, it may be used in the mixing length formula $\gamma/k_\perp^2$ to predict the turbulent saturation level, 
-and it is proportional to the transport rate~\cite{terry2006b}. 
+When positive for some wavenumber, turbulence can be
+sustained. Additionally, it may be used in a mixing length argument
+to predict the turbulent saturation level, 
+which can in turn be related to the rate of cross-field transport~\cite{terry2006b}. 
 $\gamma_l(m,n)$ may be calculated from the spatial structures of the plasma state variables, so it is always well-defined, even in a turbulent plasma. We plot
@@
 To accomplish this, we make the anzatz that nonlinearities randomize the turbulent spatial structure at each wavenumber on a time scale of 
-one eddy decorrelation time, while the linearities evolve the spatial structures deterministically. From this we can calculate a non-modal $\gamma_{\rm{n-m}}$ spectrum,
+one eddy decorrelation time, while the linearities evolve the spatial structures deterministically. From this we can calculate a non-modal $\gamma_{\rm{n-m}}$ spectrum
 which is our prediction of the turbulent $\gamma_l$ spectrum. To illustrate,
@@
 depends on the initial condition $\mathbf{u}(0)$. For purposes of turbulent growth rate prediction, we are interested in
-the behavior of $G(t) = E(t)/E(0) = ||\mathbf{u}(t)||^2/||\mathbf{u}(0)||^2$ and its rate of growth.
+the behavior of $G(t) = E(t)/E(0) = ||\mathbf{u}(t)||^2/||\mathbf{u}(0)||^2$.
 
@@
 $G_{\rm{max}}(t)$ is only the upper envelope of all possible $G(t)$ curves. No one particular initial condition $\mathbf{u}(0)$ evolves along $G_{\rm{max}}(t)$. 
-Furthermore, it isn't obvious what kind of spatial structure or structures will come to dominate a turbulent system.
+Furthermore, it isn't obvious what kind of spatial structures will come to dominate a turbulent system.
 In non-normal systems, unlike in normal systems, optimal structures don't amplify themselves, rather, they evolve while increasing the total fluctuating energy.
@@
 time scale is generally associated with the eddy turnover or decorrelation time. We therefore present a heuristic model of the nonlinearities 
-as a randomizing force that acts on this characteristic nonlinear time scale.
-Specifically, we start with an ensemble of random initial conditions, evolve them linearly for a time $\tau_{nl}$, then take the time and ensemble averaged growth rate of these curves.
-This procedure calculates the average growth rate of the turbulence as if the turbulence evolved linearly for a time $\tau_{nl}$, randomized, and then repeated. 
-Now, this procedure does not require actual linear simulations from an ensemble of random initial conditions. To see this, recall that the time evolution of the energy from an initial condition is
+as a randomizing force that acts on this characteristic nonlinear time scale. Again, this model is one in which 1) the turbulence begins as a random state, 
+2) evolves linearly for a time $\tau_{nl}$, and 3) randomizes by nonlinear energy transfer, at which point the steps repeat.
+In practice, we implement this model by starting with an ensemble of random initial conditions, which we evolve linearly for a time $\tau_{nl}$, 
+and then take the time and ensemble averaged growth rate of these curves.
+This procedure does not require actual linear simulations from an ensemble of random initial conditions. To see this, recall that the time evolution of the energy from an initial condition is
 
@@
 \eeq
-We show the validity of this statistical averaging by plotting $1000$ ensemble curves in Fig.~\ref{m20n0_E_ev} along with their expected average from Eq.~\ref{E_t_ensemble_avg} and their actual average,
-which line up perfectly.
-One point to note is that in global equation sets like ours, where we discritize and use finite differences in the radial direction rather than a Fourier decomposition, randomizing $\mathbf{u}(0)$
+We show the validity of this statistical averaging by plotting an
+ensemble of $1000$ curves generated with different random initial conditions in Fig.~\ref{m20n0_E_ev} along with their expected average from Eq.~\ref{E_t_ensemble_avg} and their actual average,
+which agree well.
+One point to note is that in global equation sets like ours -- where we discritize and use finite differences in the radial direction rather than a Fourier decomposition -- randomizing $\mathbf{u}(0)$
 amounts to setting the initial $k_r$ spectrum to a step function that goes to zero at the Nyquist wavenumber.
@@
 Second is the parallel free-streaming time of the electrons $t_{\rm{st}} = L_\para/v_{t e}$ often cited in critical balance arguments~\cite{barnes2011}. Third is the time before modal effects
-take over, which can be approximated as the time when $E(t)$ turns over. As seen in Fig.~\ref{n0_n1_growth}, $E(t)$ can be either a maximum or minimum before turning over. 
+take over, which can be approximated as the time when $E(t)$ turns over. As seen in Fig.~\ref{n0_n1_growth} b), $E(t)$ can be either a maximum or minimum before turning over. 
 We label this time $E_{\rm{max}}$. 
@@
 In summary, we present a procedure for calculating the turbulent growth rate spectrum using non-modal linear calculations. 
-In the case of an LAPD experiment, this procedure captures the behavior of a nonlinear instability that dominates the dynamics of the turbulence.  In general,
+In the case of a simulation of an LAPD experiment, this procedure captures the behavior of a nonlinear instability that dominates the dynamics of the turbulence.  In general,
 non-modal analysis is difficult to quantify and make predictive, but using some simple nonlinear modelling, we have shown
-that it may be possible. Future studies will attempt to test this procedure on other non-normal turbulence models and to see if it can predict critical parameters for subcritical turbulent onset.
+that it may be possible. Future studies will attempt to test this procedure on other turbulence models, and see if it can predict critical parameters for subcritical turbulent onset.
 
